---
title: "Taiwanese Mandarin / Korean / Kapampangan speech rate vs. rhythm"
author: "Márton Sóskuthy"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(ggbeeswarm)

twm <- read_csv("../data/taiwanese_mandarin_durations.csv") %>%
  arrange(speaker, file, phone_start) %>%
  mutate(recording_type = str_extract(file, "[rs]s$")) %>%
  filter(speaker != "mn_tw_66")
```

How much do speech rates vary?

```{r}
twm %>%
  group_by(speaker, recording_type, uttr_id) %>%
  summarise(speech_rate=speech_rate[1]) %>%
  ungroup() %>%
ggplot(aes(x=recording_type, col=recording_type, y=speech_rate)) +
  facet_wrap(~ speaker) +
  geom_beeswarm() +
  scale_y_log10()

twm %>%
  group_by(speaker, recording_type, uttr_id) %>%
  summarise(speech_rate=speech_rate[1]) %>%
  ungroup() %>%
ggplot(aes(x=recording_type, fill=recording_type, y=speech_rate)) +
  geom_violin() +
  stat_summary(fun=mean, geom="point", shape=16) +
  scale_y_log10()
```

Some data filtering.

```{r}
# utterance level filtering
# - at least 5 syllables
# - log sd based filtering for top end of speech rate

twm_uttr <- twm %>% 
  group_by(uttr_id, speaker) %>%
  summarise(speech_rate=speech_rate[1]) %>%
  ungroup() %>%
  group_by(speaker) %>%
  filter(log(speech_rate) < mean(log(speech_rate)) + 3*sd(log(speech_rate))) %>%
  ungroup()

twm <- twm %>%
  filter(
    num_syllables_uttr >= 5,
    uttr_id %in% twm_uttr$uttr_id
  )
```

Plot speech rates again.

```{r}
twm %>%
  group_by(speaker, recording_type, uttr_id) %>%
  summarise(speech_rate=speech_rate[1]) %>%
  ungroup() %>%
ggplot(aes(x=recording_type, col=recording_type, y=speech_rate)) +
  facet_wrap(~ speaker) +
  geom_beeswarm() +
  scale_y_log10()

twm %>%
  group_by(speaker, recording_type, uttr_id) %>%
  summarise(speech_rate=speech_rate[1]) %>%
  ungroup() %>%
ggplot(aes(x=recording_type, fill=recording_type, y=speech_rate)) +
  geom_violin() +
  stat_summary(fun=mean, geom="point", shape=16) +
  scale_y_log10()
```

Right, calculate %V!

```{r}
twm_rhythm <- twm %>%
  group_by(
    speaker, 
    sex,
    age,
    recording_type,
    uttr_id
  ) %>%
  summarise(
    speech_rate=num_syllables_uttr[1]/sum(phone_dur[segment_type %in% c("vowel", "consonant")]),
    speech_rate_phone=length(phone_dur[segment_type %in% c("vowel", "consonant")])/sum(phone_dur[segment_type %in% c("vowel", "consonant")]),
    num_words_uttr=num_words_uttr[1],
    num_syllables_uttr=num_syllables_uttr[1],
    uttr_start=uttr_start[1],
    uttr_end=uttr_end[1],
    uttr_dur=uttr_dur[1],
    v_prop=sum(phone_dur[segment_type=="vowel"])/sum(phone_dur[segment_type %in% c("vowel", "consonant")]),
    v_dur=mean(phone_dur[segment_type=="vowel"]),
    c_dur=mean(phone_dur[segment_type=="consonant"])
  ) %>%
  ungroup() %>%
  filter((1/speech_rate_phone) < 0.2)
```

Exploratory plots!

```{r}
ggplot(twm_rhythm, aes(x=1/speech_rate, y=v_prop*100)) +
  geom_point() +
  geom_smooth(method="gam") +
  ylab("V%") +
  xlab("average segment duration (1 / speech rate)")
ggsave("graphs/twm_avg_seg_dur-v_perc.png", width=6, height=4.5, dpi=300)

ggplot(twm_rhythm, aes(x=1/speech_rate_phone, y=v_prop*100)) +
  facet_wrap(~speaker) +
  geom_point() +
  geom_smooth(method="gam") +
  ylab("V%") +
  xlab("average segment duration (1 / speech rate)")
ggsave("graphs/twm_avg_seg_dur-v_perc-speaker.png", width=10, height=7.5, dpi=300)

ggplot(twm_rhythm, aes(x=1/speech_rate_phone)) +
  geom_point(aes(y=v_dur), col="orange", alpha=0.3) +
  geom_point(aes(y=c_dur), col="purple", alpha=0.3) +
  geom_smooth(aes(y=v_dur), col="darkorange2", method="gam", se=F) +
  geom_smooth(aes(y=c_dur), col="purple4", method="gam", se=F) +
  ylab("average duration in s")
ggsave("graphs/twm_avg_seg_dur-avg_C+V_dur.png", width=6, height=4.5, dpi=300)

# ggplot(twm_rhythm, aes(x=c_dur)) +
#   geom_point(aes(y=v_dur), col="orange", alpha=0.3) +
#   #geom_point(aes(y=c_dur), col="purple", alpha=0.3) +
#   geom_smooth(aes(y=v_dur), col="darkorange2", method="gam", se=T) +
#   geom_abline(slope=1, intercept=0, col="black") +
#   coord_equal() +
#   #geom_smooth(aes(y=c_dur), col="purple4", method="gam", se=F) +
#   ylab("average duration in s")

ggplot(twm_rhythm, aes(x=1/speech_rate)) +
  facet_wrap(~speaker) +
  geom_point(aes(y=v_dur), col="orange", alpha=0.3) +
  geom_point(aes(y=c_dur), col="purple", alpha=0.3) +
  geom_smooth(aes(y=v_dur), col="darkorange2", method="gam", se=F) +
  geom_smooth(aes(y=c_dur), col="purple4", method="gam", se=F) +
  ylab("average duration in s")
ggsave("graphs/twm_avg_seg_dur-avg_C+V_dur-speaker.png", width=10, height=7.5, dpi=300)
```

## Korean

```{r}
library(tidyverse)
library(ggbeeswarm)

k <- read_csv("../data/korean_durations.csv") %>%
  arrange(speaker, file, phone_start) %>%
  mutate(recording_type = str_extract(file, "[rs]s$"))
```

How much do speech rates vary?

```{r}
k %>%
  group_by(speaker, recording_type, uttr_id) %>%
  summarise(speech_rate=speech_rate[1]) %>%
  ungroup() %>%
ggplot(aes(x=recording_type, col=recording_type, y=speech_rate)) +
  facet_wrap(~ speaker) +
  geom_beeswarm() +
  scale_y_log10()

k %>%
  group_by(speaker, recording_type, uttr_id) %>%
  summarise(speech_rate=speech_rate[1]) %>%
  ungroup() %>%
ggplot(aes(x=recording_type, fill=recording_type, y=speech_rate)) +
  geom_violin() +
  stat_summary(fun=mean, geom="point", shape=16) +
  scale_y_log10()
```

Some data filtering.

```{r}
# utterance level filtering
# - at least 5 syllables
# - log sd based filtering for top end of speech rate

k_uttr <- k %>% 
  group_by(uttr_id, speaker) %>%
  summarise(speech_rate=speech_rate[1]) %>%
  ungroup() %>%
  group_by(speaker) %>%
  filter(log(speech_rate) < mean(log(speech_rate)) + 3*sd(log(speech_rate))) %>%
  ungroup()

k <- k %>%
  filter(
    num_syllables_uttr >= 5,
    uttr_id %in% k_uttr$uttr_id
  )
```

Plot speech rates again.

```{r}
k %>%
  group_by(speaker, recording_type, uttr_id) %>%
  summarise(speech_rate=speech_rate[1]) %>%
  ungroup() %>%
ggplot(aes(x=recording_type, col=recording_type, y=speech_rate)) +
  facet_wrap(~ speaker) +
  geom_beeswarm() +
  scale_y_log10()

k %>%
  group_by(speaker, recording_type, uttr_id) %>%
  summarise(speech_rate=speech_rate[1]) %>%
  ungroup() %>%
ggplot(aes(x=recording_type, fill=recording_type, y=speech_rate)) +
  geom_violin() +
  stat_summary(fun=mean, geom="point", shape=16) +
  scale_y_log10()
```

Right, calculate %V!

```{r}
k_rhythm <- k %>%
  group_by(
    speaker, 
    sex,
    age,
    recording_type,
    uttr_id
  ) %>%
  summarise(
    speech_rate=num_syllables_uttr[1]/sum(phone_dur[segment_type %in% c("vowel", "consonant")]),
    speech_rate_phone=length(phone_dur[segment_type %in% c("vowel", "consonant")])/sum(phone_dur[segment_type %in% c("vowel", "consonant")]),
    num_words_uttr=num_words_uttr[1],
    num_syllables_uttr=num_syllables_uttr[1],
    uttr_start=uttr_start[1],
    uttr_end=uttr_end[1],
    uttr_dur=uttr_dur[1],
    v_prop=sum(phone_dur[segment_type=="vowel"])/sum(phone_dur[segment_type %in% c("vowel", "consonant")]),
    v_dur=mean(phone_dur[segment_type=="vowel"]),
    c_dur=mean(phone_dur[segment_type=="consonant"])
  ) %>%
  ungroup() %>%
  filter((1/speech_rate_phone) < 0.2)
```

Exploratory plots!

```{r}
ggplot(k_rhythm, aes(x=1/speech_rate_phone, y=v_prop*100)) +
  geom_point() +
  geom_smooth(method="gam") +
  ylab("V%") +
  xlab("average segment duration (1 / speech rate)")
ggsave("graphs/k_avg_seg_dur-v_perc.png", width=6, height=4.5, dpi=300)

ggplot(k_rhythm, aes(x=1/speech_rate_phone, y=v_prop*100)) +
  facet_wrap(~speaker) +
  geom_point() +
  geom_smooth(method="gam") +
  ylab("V%") +
  xlab("average segment duration (1 / speech rate)")
ggsave("graphs/k_avg_seg_dur-v_perc-speaker.png", width=10, height=7.5, dpi=300)

ggplot(k_rhythm, aes(x=1/speech_rate_phone)) +
  geom_point(aes(y=v_dur), col="orange", alpha=0.3) +
  geom_point(aes(y=c_dur), col="purple", alpha=0.3) +
  geom_smooth(aes(y=v_dur), col="darkorange2", method="gam", se=F) +
  geom_smooth(aes(y=c_dur), col="purple4", method="gam", se=F) +
  ylab("average duration in s")
ggsave("graphs/k_avg_seg_dur-avg_C+V_dur.png", width=6, height=4.5, dpi=300)

ggplot(k_rhythm, aes(x=1/speech_rate_phone)) +
  facet_wrap(~speaker) +
  geom_point(aes(y=v_dur), col="orange", alpha=0.3) +
  geom_point(aes(y=c_dur), col="purple", alpha=0.3) +
  geom_smooth(aes(y=v_dur), col="darkorange2", method="gam", se=F) +
  geom_smooth(aes(y=c_dur), col="purple4", method="gam", se=F) +
  ylab("average duration in s")
ggsave("graphs/k_avg_seg_dur-avg_C+V_dur-speaker.png", width=10, height=7.5, dpi=300)
```

## Pampango

For now, looking at read speech only.

```{r}
library(tidyverse)
library(ggbeeswarm)

p <- read_csv("../data/kapampangan_durations.csv") %>%
  arrange(speaker, file, phone_start) %>%
  mutate(recording_type = str_extract(file, "[rs]s$"))
```

How much do speech rates vary?

```{r}
p %>%
  group_by(speaker, recording_type, uttr_id) %>%
  summarise(speech_rate=speech_rate[1]) %>%
  ungroup() %>%
ggplot(aes(x=recording_type, col=recording_type, y=speech_rate)) +
  facet_wrap(~ speaker) +
  geom_beeswarm() +
  scale_y_log10()

p %>%
  group_by(speaker, recording_type, uttr_id) %>%
  summarise(speech_rate=speech_rate[1]) %>%
  ungroup() %>%
ggplot(aes(x=recording_type, fill=recording_type, y=speech_rate)) +
  geom_violin() +
  stat_summary(fun=mean, geom="point", shape=16) +
  scale_y_log10()
```

Some data filtering.

```{r}
# utterance level filtering
# - at least 5 syllables
# - log sd based filtering for top end of speech rate

p_uttr <- p %>% 
  group_by(uttr_id, speaker) %>%
  summarise(speech_rate=speech_rate[1]) %>%
  ungroup() %>%
  group_by(speaker) %>%
  filter(log(speech_rate) < mean(log(speech_rate)) + 3*sd(log(speech_rate))) %>%
  ungroup()

p <- p %>%
  filter(
    num_syllables_uttr >= 5,
    uttr_id %in% p_uttr$uttr_id
  )
```

Plot speech rates again.

```{r}
p %>%
  group_by(speaker, recording_type, uttr_id) %>%
  summarise(speech_rate=speech_rate[1]) %>%
  ungroup() %>%
ggplot(aes(x=recording_type, col=recording_type, y=speech_rate)) +
  facet_wrap(~ speaker) +
  geom_beeswarm() +
  scale_y_log10()

p %>%
  group_by(speaker, recording_type, uttr_id) %>%
  summarise(speech_rate=speech_rate[1]) %>%
  ungroup() %>%
ggplot(aes(x=recording_type, fill=recording_type, y=speech_rate)) +
  geom_violin() +
  stat_summary(fun=mean, geom="point", shape=16) +
  scale_y_log10()
```

Right, calculate %V!

```{r}
p_rhythm <- p %>%
  group_by(
    speaker, 
    sex,
    age,
    recording_type,
    uttr_id
  ) %>%
  summarise(
    speech_rate=num_syllables_uttr[1]/sum(phone_dur[segment_type %in% c("vowel", "consonant")]),
    speech_rate_phone=length(phone_dur[segment_type %in% c("vowel", "consonant")])/sum(phone_dur[segment_type %in% c("vowel", "consonant")]),
    num_words_uttr=num_words_uttr[1],
    num_syllables_uttr=num_syllables_uttr[1],
    uttr_start=uttr_start[1],
    uttr_end=uttr_end[1],
    uttr_dur=uttr_dur[1],
    v_prop=sum(phone_dur[segment_type=="vowel"])/sum(phone_dur[segment_type %in% c("vowel", "consonant")]),
    v_dur=mean(phone_dur[segment_type=="vowel"]),
    c_dur=mean(phone_dur[segment_type=="consonant"]),
    v_count_prop=length(phone_dur[segment_type=="vowel"])/length(phone_dur)
      ) %>%
  ungroup() %>%
  filter((1/speech_rate_phone) < 0.2)
```

Exploratory plots!

```{r}
ggplot(p_rhythm, aes(x=1/speech_rate_phone, y=v_prop*100)) +
  geom_point() +
  geom_smooth(method="gam") +
  ylab("V%") +
  xlab("average segment duration (1 / speech rate)")
ggsave("graphs/p_avg_seg_dur-v_perc.png", width=6, height=4.5, dpi=300)

ggplot(p_rhythm, aes(x=1/speech_rate_phone, y=v_prop*100)) +
  facet_wrap(~speaker) +
  geom_point() +
  geom_smooth(method="gam") +
  ylab("V%") +
  xlab("average segment duration (1 / speech rate)")
ggsave("graphs/p_avg_seg_dur-v_perc-speaker.png", width=10, height=7.5, dpi=300)

ggplot(p_rhythm, aes(x=1/speech_rate_phone)) +
  geom_point(aes(y=v_dur), col="orange", alpha=0.3) +
  geom_point(aes(y=c_dur), col="purple", alpha=0.3) +
  geom_smooth(aes(y=v_dur), col="darkorange2", method="gam", se=F) +
  geom_smooth(aes(y=c_dur), col="purple4", method="gam", se=F) +
  ylab("average duration in s")
ggsave("graphs/p_avg_seg_dur-avg_C+V_dur.png", width=6, height=4.5, dpi=300)

ggplot(p_rhythm, aes(x=1/speech_rate_phone)) +
  facet_wrap(~speaker) +
  geom_point(aes(y=v_dur), col="orange", alpha=0.3) +
  geom_point(aes(y=c_dur), col="purple", alpha=0.3) +
  geom_smooth(aes(y=v_dur), col="darkorange2", method="gam", se=F) +
  geom_smooth(aes(y=c_dur), col="purple4", method="gam", se=F) +
  ylab("average duration in s")
ggsave("graphs/p_avg_seg_dur-avg_C+V_dur-speaker.png", width=10, height=7.5, dpi=300)
```

## Buckeye

Note the RDS file -- the original CSV was 200 MB, so too large for GitHub.

```{r}
library(tidyverse)
library(ggbeeswarm)

b <- readRDS("../data/buckeye_durations.rds") %>%
  arrange(speaker, file, phone_start)
```

How much do speech rates vary?

```{r}
b %>%
  group_by(speaker, uttr_id) %>%
  summarise(speech_rate=speech_rate[1]) %>%
  ungroup() %>%
ggplot(aes(x=NA, y=speech_rate)) +
  facet_wrap(~ speaker) +
  geom_beeswarm() +
  scale_y_log10()

b %>%
  group_by(speaker, uttr_id) %>%
  summarise(speech_rate=speech_rate[1]) %>%
  ungroup() %>%
ggplot(aes(x=NA, y=speech_rate)) +
  geom_violin() +
  stat_summary(fun=mean, geom="point", shape=16) +
  scale_y_log10()
```

Some data filtering.

```{r}
# utterance level filtering
# - at least 5 syllables
# - log sd based filtering for top end of speech rate

b_uttr <- b %>% 
  group_by(uttr_id, speaker) %>%
  summarise(speech_rate=speech_rate[1]) %>%
  ungroup() %>%
  group_by(speaker) %>%
  filter(log(speech_rate) < mean(log(speech_rate)) + 3*sd(log(speech_rate)),
         log(speech_rate) > mean(log(speech_rate)) - 3*sd(log(speech_rate)),
         speech_rate > 1) %>%
  ungroup()

b <- b %>%
  filter(
    num_syllables_uttr >= 5,
    uttr_id %in% b_uttr$uttr_id
  )
```

Plot speech rates again.

```{r}
b %>%
  group_by(speaker, uttr_id) %>%
  summarise(speech_rate=speech_rate[1]) %>%
  ungroup() %>%
ggplot(aes(x=NA, y=speech_rate)) +
  facet_wrap(~ speaker) +
  geom_beeswarm() +
  scale_y_log10()

b %>%
  group_by(speaker, uttr_id) %>%
  summarise(speech_rate=speech_rate[1]) %>%
  ungroup() %>%
ggplot(aes(x=NA, y=speech_rate)) +
  geom_violin() +
  stat_summary(fun=mean, geom="point", shape=16) +
  scale_y_log10()
```

Right, calculate %V! (Note that syllabic consonants are excluded from the calculation of *all* segment-based metrics.)

```{r}
b_rhythm <- b %>%
  mutate(segment_type = ifelse(phone %in% c("en","em","el","eng"), "syllabic consonant", segment_type)) %>%
  group_by(
    speaker, 
    uttr_id
  ) %>%
  summarise(
    speech_rate=num_syllables_uttr[1]/sum(phone_dur[segment_type %in% c("vowel", "consonant", "syllabic consonant")]),
    speech_rate_phone=length(phone_dur[segment_type %in% c("vowel", "consonant")])/sum(phone_dur[segment_type %in% c("vowel", "consonant")]),
    num_words_uttr=num_words_uttr[1],
    num_syllables_uttr=num_syllables_uttr[1],
    uttr_start=uttr_start[1],
    uttr_end=uttr_end[1],
    uttr_dur=uttr_dur[1],
    v_prop=sum(phone_dur[segment_type=="vowel"])/sum(phone_dur[segment_type %in% c("vowel", "consonant")]),
    v_dur=mean(phone_dur[segment_type=="vowel"]),
    c_dur=mean(phone_dur[segment_type %in% c("consonant")]),
    v_count_prop=length(phone_dur[segment_type=="vowel"])/length(phone_dur)
      ) %>%
  ungroup() %>%
  filter((1/speech_rate_phone) < 0.2)
```

Exploratory plots!

```{r}
ggplot(b_rhythm, aes(x=1/speech_rate_phone, y=v_prop*100)) +
  geom_point() +
  geom_smooth(method="gam") +
  ylab("V%") +
  xlab("average segment duration (1 / speech rate)")
ggsave("graphs/b_avg_seg_dur-v_perc.png", width=6, height=4.5, dpi=300)

ggplot(b_rhythm, aes(x=1/speech_rate_phone, y=v_prop*100)) +
  facet_wrap(~speaker) +
  geom_point() +
  geom_smooth(method="gam") +
  ylab("V%") +
  xlab("average segment duration (1 / speech rate)")
ggsave("graphs/b_avg_seg_dur-v_perc-speaker.png", width=10, height=7.5, dpi=300)

ggplot(b_rhythm, aes(x=1/speech_rate_phone)) +
  geom_point(aes(y=v_dur), col="orange", alpha=0.3) +
  geom_point(aes(y=c_dur), col="purple", alpha=0.3) +
  geom_smooth(aes(y=v_dur), col="darkorange2", method="gam", se=F) +
  geom_smooth(aes(y=c_dur), col="purple4", method="gam", se=F) +
  ylab("average duration in s")
ggsave("graphs/b_avg_seg_dur-avg_C+V_dur.png", width=6, height=4.5, dpi=300)

ggplot(b_rhythm, aes(x=1/speech_rate_phone)) +
  facet_wrap(~speaker) +
  geom_point(aes(y=v_dur), col="orange", alpha=0.3) +
  geom_point(aes(y=c_dur), col="purple", alpha=0.3) +
  geom_smooth(aes(y=v_dur), col="darkorange2", method="gam", se=F) +
  geom_smooth(aes(y=c_dur), col="purple4", method="gam", se=F) +
  ylab("average duration in s")
ggsave("graphs/b_avg_seg_dur-avg_C+V_dur-speaker.png", width=10, height=7.5, dpi=300)
```

## Plots for abstract

Make a combined data set: added columns of language + pivoting based on measure. Measures: v_prop, v_dur, c_dur.

```{r}
twm_rhythm$language <- "twm"
k_rhythm$language <- "sk"
p_rhythm$language <- "kp"
all_uttr <- bind_rows(twm_rhythm, k_rhythm, p_rhythm) %>%
  mutate(v_prop=v_prop*100) %>%
  pivot_longer(c(v_prop, v_dur, c_dur),
               names_to="measure",
               values_to="value") %>%
  mutate(measure_type = ifelse(measure=="v_prop", "v_prop", "dur"),
         measure_type = factor(measure_type, levels=c("v_prop", "dur")),
         measure = factor(measure, levels=c("v_prop", "v_dur", "c_dur"))
  )
```

Make combined plot!

```{r}
ggplot(all_uttr, aes(x=1/speech_rate_phone, y=value, col=measure)) +
  facet_grid(measure_type ~ language,
             scales = "free_y", 
             labeller = as_labeller(
               c(v_prop = "%V", dur = "Average V/C duration",
                 twm="Taiwanese Mandarin", sk="Seoul Korean",
                 kp="Kapampangan")), 
             switch = "y") +
  geom_point(alpha=0.3) +
  geom_smooth(data=filter(all_uttr, measure=="v_prop"), col="blue", method="gam", se=T) +
  geom_smooth(data=filter(all_uttr, measure=="v_dur"), col="darkorange2", method="gam", se=F) +
  geom_smooth(data=filter(all_uttr, measure=="c_dur"), col="purple4", method="gam", se=F) +
  scale_x_continuous(limits=c(0.035,0.145), 
                     breaks=seq(0.04,0.14,0.02)) +
  scale_colour_manual(values=c("black", "orange", "purple"), guide="none") +
  #scale_y_continuous(breaks=seq(0.04,0.20,0.02)) +
    # labels for axes
  xlab("Average segment duration in s (inverse speech rate)") +
  labs(y=NULL) +
  #ylab("Average V vs. C duration in s") +
  # remove clutter
  theme_minimal() +
  # some specific formatting...
  theme(panel.grid = element_blank(),
        axis.line = element_line(size=0.8),
        axis.ticks = element_line(size=0.8),
        axis.text = element_text(size=14, colour="black"),
        axis.title = element_text(size=16, face="bold"),
        strip.background.y = element_blank(), 
        strip.placement = "outside",
        strip.text=element_text(size=16, face="bold"))
ggsave("graphs/lsurc_plot.png", width=10, height=6, dpi=300)
  
```
