---
title: "Taiwanese Mandarin / Korean / Kapampangan / English / Vietnamese articulation rate vs. rhythm"
author: "Márton Sóskuthy & Roger Lo"
output: html_document
editor_options: 
  chunk_output_type: console
---

Shared data processing pipeline for the four Ls.

```{r}
library(tidyverse)
library(ggbeeswarm)
library(ggpubr)
library(geomtextpath)

setwd("Documents/RA/oops_lab-speech_rate/prelim_analysis/")

twm <- read_csv("../data/taiwanese_mandarin_durations.csv") %>%
  arrange(speaker, file, phone_start) %>%
  filter(speaker != "mn_tw_66") %>%
  select(-sex, -age) %>%
  mutate(language="Taiwan Mandarin",
         folder="taiwanese_mandarin")
k <- read_csv("../data/korean_durations.csv") %>%
  arrange(speaker, file, phone_start) %>%
  select(-sex, -age) %>%
  mutate(language="Seoul Korean",
         folder="korean")
p <- read_csv("../data/kapampangan_durations.csv") %>%
  arrange(speaker, file, phone_start) %>%
  select(-sex, -age) %>%
  mutate(language="Kapampangan",
         folder="kapampangan")
v <- read_csv("../data/vietnamese_durations.csv") %>%
  arrange(speaker, file, phone_start) %>%
  #select(-sex, -age) %>%
  mutate(language="Vietnamese",
         speaker=as.character(speaker),
         folder="vietnamese")
s <- read_csv("../data/swahili_durations.csv") %>%
  arrange(speaker, file, phone_start) %>%
  #select(-sex, -age) %>%
  mutate(language="Swahili",
         speaker=as.character(speaker),
         folder="swahili")
tu <- read_csv("../data/turkish_durations.csv") %>%
  arrange(speaker, file, phone_start) %>%
  #select(-sex, -age) %>%
  mutate(language="Turkish",
         speaker=as.character(speaker),
         folder="turkish")
a <- read_csv("../data/amharic_durations.csv") %>%
  arrange(speaker, file, phone_start) %>%
  #select(-sex, -age) %>%
  mutate(language="Amharic",
         speaker=as.character(speaker),
         folder="amharic")
g <- read_csv("../data/georgian_durations.csv") %>%
  arrange(speaker, file, phone_start) %>%
  #select(-sex, -age) %>%
  mutate(language="Georgian",
         speaker=as.character(speaker),
         folder="georgian")
b <- readRDS("../data/buckeye_durations.rds") %>%
  arrange(speaker, file, phone_start) %>%
  mutate(language="American English",
    segment_type=case_when(
      phone %in% c("{E_TRANS}", "{B_TRANS}", "id", "no", "hhn", "an") ~ "non-speech",
      phone %in% c("a", "e", "ih l", "ah r", "i", "ah l", "ah ix", "uw ix", "ah n") ~ "vowel",
      TRUE ~ segment_type
    )
  )

d <- bind_rows(twm, k, p, b, v, s, tu, a, g)
```

Coding consonant type.

```{r}
d <- d %>%
  mutate(
    c_type = case_when(
      ## Taiwan Mandarin
      language == "Taiwan Mandarin" & 
        phone %in% c("p", "b", "t", "d", "k", "g") ~ "stop",
      language == "Taiwan Mandarin" & 
        phone %in% c("f", "sh", "s", "x", "h") ~ "fricative",
      language == "Taiwan Mandarin" & 
        phone %in% c("j", "q", "zh", "ch", "z", "c") ~ "affricate",
      language == "Taiwan Mandarin" & 
        phone %in% c("m", "n", "ng") ~ "nasal",
      language == "Taiwan Mandarin" & 
        phone %in% c("y", "w", "ue") ~ "approximant",
      language == "Taiwan Mandarin" & 
        phone %in% c("l", "r") ~ "liquid",
      language == "Taiwan Mandarin" & 
        phone %in% c("i","ii","v","u") ~ "high vowel",
      ## Korean
      language == "Seoul Korean" & 
        phone %in% c("B", "Ph", "BB", "p", "D", "Th", "DD", "t", "G", "Kh", "GG", "k") ~ "stop",
      language == "Seoul Korean" & 
        phone %in% c("S", "SS", "H") ~ "fricative",
      language == "Seoul Korean" & 
        phone %in% c("J", "CHh", "JJ") ~ "affricate",
      language == "Seoul Korean" & 
        phone %in% c("M", "N", "NG") ~ "nasal",
      language == "Seoul Korean" & 
        phone %in% c("L", "R") ~ "liquid",
      language == "Seoul Korean" & 
        phone %in% c("I","U","EU","UE","iU","euI") ~ "high vowel",
      ## Kapampangan
      language == "Kapampangan" & 
        phone %in% c("b", "p", "d", "t", "ɡ", "k", "ʔ", "q", "c") ~ "stop",
      language == "Kapampangan" & 
        phone %in% c("s", "ʃ", "f", "h", "v", "x") ~ "fricative",
      language == "Kapampangan" & 
        phone %in% c("tʃ", "dʒ") ~ "affricate",
      language == "Kapampangan" & 
        phone %in% c("m", "n", "ŋ") ~ "nasal",
      language == "Kapampangan" & 
        phone %in% c("j", "w") ~ "approximant",
      language == "Kapampangan" & 
        phone %in% c("l", "r") ~ "liquid",
      language == "Kapampangan" & 
        phone %in% c("i", "ɪ", "u", "ʊ") ~ "high vowel",
      ## American English
      language == "American English" &
        phone %in% c("p", "b", "t", "d", "k", "g", "dx", "q", "tq") ~ "stop",
      language == "American English" &
        phone %in% c("m", "n", "ng", "nx") ~ "nasal",
      language == "American English" &
        phone %in% c("f", "v", "th", "dh", "s", "z", "sh", "zh", "hh", "h", "x") ~ "fricative",
      language == "American English" &
        phone %in% c("ch", "jh", "j") ~ "affricate",
      language == "American English" &
        phone %in% c("r", "l") ~ "liquid",
      language == "American English" &
        phone %in% c("w", "y") ~ "approximant",
      language == "American English" &
        phone %in% c("em", "en", "el", "eng") ~ "syllabic C",
      language == "American English" & 
        phone %in% c("ih", "iy", "uw", "ihn", "uhn", "iyn", "uwn") ~ "high vowel",
      ## Vietnamese
      language == "Vietnamese" &
        phone %in% c("p", "t", "c", "k", "tʰ", "ɓ", "ɗ", "ʔ") ~ "stop",
      language == "Vietnamese" &
        phone %in% c("m", "n", "ɲ", "ŋ") ~ "nasal",
      language == "Vietnamese" &
        phone %in% c("f", "v", "s", "z", "x", "ɣ", "h") ~ "fricative",
      language == "Vietnamese" &
        phone %in% c("tɕ") ~ "affricate",
      language == "Vietnamese" &
        phone %in% c("l") ~ "liquid",
      language == "Vietnamese" &
        phone %in% c("w", "j") ~ "approximant",
      language == "Vietnamese" & 
        phone %in% (unique(filter(v, segment_type=="vowel")$phone) %>% str_subset("^[iɨu][^ə]")) ~ "high vowel",
      ## Swahili
      language == "Swahili" &
        phone %in% c('b', 'd', 'k', 'p', 't', 'ɓ', 'ɗ', 'ɠ', 'ɡ', 'ʄ') ~ "stop",
      language == "Swahili" &
        phone %in% c('m', 'n', 'ŋ', 'ɱ', 'ɲ') ~ "nasal",
      language == "Swahili" &
        phone %in% c('f', 'h', 's', 'v', 'z', 'ð', 'ɣ', 'ʃ', 'θ') ~ "fricative",
      language == "Swahili" &
        phone %in% c('dʒ', 'tʃ') ~ "affricate",
      language == "Swahili" &
        phone %in% c("l", "ɾ") ~ "liquid",
      language == "Swahili" &
        phone %in% c("w", "j") ~ "approximant",
      language == "Swahili" & 
        phone %in% c("i","u","iː","uː") ~ "high vowel",
      ## Turkish
      language == "Turkish" &
        phone %in% c('b', 'bː', 'c', 'cː', 'd̪', 'd̪ː','k', 'kː', 'p','t̪', 't̪ː', 'ɟ', 'ɡ') ~ "stop",
      language == "Turkish" &
        phone %in% c('m', 'mː', 'n̪', 'n̪ː','ŋ') ~ "nasal",
      language == "Turkish" &
        phone %in% c('f', 'h', 'hː', 's̪', 's̪ː','v', 'z̪','ç','ʃ','ʒ') ~ "fricative",
      language == "Turkish" &
        phone %in% c('dʒ', 'tʃ') ~ "affricate",
      language == "Turkish" &
        phone %in% c('ɾ','ɫ', 'ɫː', 'ʎ', 'ʎː') ~ "liquid",
      language == "Turkish" &
        phone %in% c("j") ~ "approximant",
      language == "Turkish" & 
        phone %in% c("i","ɯ","u","y","iː","ɯː","uː","yː","ɨ","ʏ") ~ "high vowel",
      ## Georgian
      language == "Georgian" &
        phone %in% c('b', 'd', 'kʰ', 'kʼ', 'pʰ', 'pʼ', 'qʼ', 'tʰ', 'tʼ', 'ɡ') ~ "stop",
      language == "Georgian" &
        phone %in% c('m', 'n') ~ "nasal",
      language == "Georgian" &
        phone %in% c('h', 's', 'v', 'x', 'z', 'ɣ', 'ʃ', 'ʒ') ~ "fricative",
      language == "Georgian" &
        phone %in% c('dz', 'dʒ', 'ts', 'tsʼ', 'tʃ', 'tʃʼ') ~ "affricate",
      language == "Georgian" &
        phone %in% c('ɾ', 'l') ~ "liquid",
      language == "Georgian" &
        phone %in% c('i', 'u') ~ "high vowel",
      ## Amharic
      language == "Amharic" &
        phone %in% c('p', 'pʷ', 'pʼ', 'pʷʼ', 'b', 'pʷ', 't', 'tʷ', 'tʼ', 'tʷʼ', 'd', 'dʷ', 'k', 'kʷ', 'kʼ', 'kʷʼ', 'ɡ', 'ɡʷ', 'ʔ') ~ "stop",
      language == "Amharic" &
        phone %in% c('m', 'mʷ', 'n', 'nʷ', 'ɲ', 'ɲʷ') ~ "nasal",
      language == "Amharic" &
        phone %in% c('v', 'vʷ', 'f', 'fʷ', 's', 'sʷ', 'sʼ', 'sʷʼ', 'z', 'zʷ', 'ʃ', 'ʃʷ', 'ʒ', 'ʒʷ', 'h', 'hʷ') ~ "fricative",
      language == "Amharic" &
        phone %in% c('t͡ʃ', 't͡ʃʷ', 't͡ʃʼ', 't͡ʃʷʼ', 'd͡ʒ', 'd͡ʒʷ') ~ "affricate",
      language == "Amharic" &
        phone %in% c('r', 'rʷ', 'l', 'lʷ') ~ "liquid",
      language == "Amharic" &
        phone %in% c('w', 'j') ~ "approximant",
      language == "Amharic" &
        phone %in% c('i', 'u', "ɨ") ~ "high vowel"#,
      #segment_type == "vowel" &  ~ "vowel"
    ),
    c_type = ifelse(is.na(c_type), "NA", c_type),
    c_type = ifelse(segment_type == "vowel" & (c_type != "high vowel"),  "non-high vowel", c_type)
  )

# check phonemes
d %>%
  filter(c_type == "approximant", language == "American English") %>%
  pull(phone) %>%
  unique()
```

Grouping by utterance, creating predictors.

```{r}
# for analysis - segment level
d_seg <- d %>%
  filter(segment_type %in% c("consonant", "vowel"),
         phone_dur > 0) %>%
  group_by(
    language,
    speaker,
    uttr_id
  ) %>%
  mutate(
    # duration of segmental material only
    uttr_dur = sum(phone_dur),
    num_syllables_uttr = sum(segment_type == "vowel" | c_type == "syllabic C"),
    num_segments_uttr = n(),
    segment_dur_uttr = uttr_dur / num_segments_uttr,
    syllable_dur_uttr = uttr_dur / num_syllables_uttr,
    ) %>%
  ungroup()
```

Actual filtering

```{r}
d_seg %>%
  ggplot(aes(x = phone_dur)) +
  facet_wrap(~language) +
  geom_density(bw = 0.04) +
  scale_x_log10()

d_seg %>%
  mutate(phone_dur_round = round(phone_dur, 2)) %>%
  group_by(language, phone_dur_round) %>%
  summarise(count = n()) %>%
  ungroup()

d_seg <- d_seg %>% 
  filter(num_syllables_uttr > 5) %>%
  filter(segment_dur_uttr < 0.25) %>%
  filter(segment_dur_uttr > 0.04) %>%
  filter(
    log(segment_dur_uttr) > 
      mean(log(segment_dur_uttr)) - 
      3*sd(log(segment_dur_uttr)),
    log(segment_dur_uttr) < 
      mean(log(segment_dur_uttr)) +
      3*sd(log(segment_dur_uttr))
  ) %>%
  filter(phone_dur > 0.02,
         phone_dur < 0.4)

# recreating speech rate measures / getting rid of
# utterances with too little data left

d_seg %>%
  group_by(language, speaker, uttr_id) %>%
  mutate(num_segments_uttr = n()) %>%
  ungroup() %>%
  ggplot(aes(x = num_segments_uttr)) +
  facet_wrap(~language) +
  geom_histogram() +
  scale_x_log10()

d_seg %>%
  group_by(language, speaker, uttr_id) %>%
  mutate(num_segments_uttr = n()) %>%
  ungroup() %>%
  pull(num_segments_uttr) %>%
  `<`(10) %>%
  sum()

d_seg <- d_seg %>%
  group_by(language, speaker, uttr_id) %>%
  mutate(
    uttr_dur = sum(phone_dur),
    num_segments_uttr = n(),
    segment_dur_uttr = uttr_dur / num_segments_uttr
  ) %>%
  ungroup() %>%
  filter(num_segments_uttr >= 10)
```

## Importing spot-checking results

Looking at Olivia's checking results

```{r}
df_check <- read_csv("../data/spot_checking.csv") %>%
  rename_with(tolower)

df_check <- df_check %>%
  group_by(language, speaker) %>%
  summarize(prop_correct = mean(phone_alignment == "OK")) %>%
  ungroup()

df_check %>%
  ggplot(aes(x = language, y = prop_correct, color = language)) +
  ggbeeswarm::geom_beeswarm() +
  #geom_jitter(width = 0.1) +
  coord_flip()
```

Transformation for weights in regression.

```{r}
transf <- function (x) {
  1 / (1 + exp(-(x - 0.6) * 15))
}
x <- seq(0, 1, 0.01)

plot(x, transf(x), ylim = c(0, 1))

df_check <- mutate(df_check, weight = transf(prop_correct))
```

```{r}
# for checking
d_check <- d %>%
  filter(segment_type %in% c("consonant", "vowel"),
         phone_dur > 0)

# for analysis
d_uttr <- d %>%
  filter(segment_type %in% c("consonant", "vowel"), phone_dur > 0) %>%
  group_by(language, speaker, uttr_id) %>%
  summarise(
    # duration of segmental material only
    uttr_dur = sum(phone_dur),
    num_syllables_uttr = sum(segment_type == "vowel" | c_type == "syllabic C"),
    num_segments_uttr = n(),
    segment_dur_uttr = uttr_dur / num_segments_uttr,
    syllable_dur_uttr = uttr_dur / num_syllables_uttr,
    vowel_dur_uttr = mean(phone_dur[segment_type == "vowel"]),
    cons_dur_uttr = mean(phone_dur[segment_type == "consonant"]),
    nh_v_dur = mean(phone_dur[c_type == "non-high vowel"]),
    h_v_dur = mean(phone_dur[c_type == "high vowel"]),
    stop_dur = mean(phone_dur[c_type == "stop"]),
    fricative_dur = mean(phone_dur[c_type == "fricative"]),
    affricate_dur = mean(phone_dur[c_type == "affricate"]),
    nasal_dur = mean(phone_dur[c_type == "nasal"]),
    approximant_dur = mean(phone_dur[c_type == "approximant"]),
    liquid_dur = mean(phone_dur[c_type == "liquid"]),
    syllabic_c_dur = mean(phone_dur[c_type == "syllabic C"]),
    perc_v = 100 * sum(phone_dur[segment_type == "vowel"]) / uttr_dur
  ) %>%
  ungroup()

d_uttr <- d_uttr %>% 
  filter(num_syllables_uttr > 5) %>%
  filter(segment_dur_uttr < 0.25) %>%
  filter(segment_dur_uttr > 0.04) %>%
  filter(
    log(segment_dur_uttr) > 
      mean(log(segment_dur_uttr)) - 
      3*sd(log(segment_dur_uttr)),
    log(segment_dur_uttr) < 
      mean(log(segment_dur_uttr)) +
      3*sd(log(segment_dur_uttr))
  )
```

Merging into main data sets.

```{r}
d_uttr <- d_uttr %>%
  left_join(df_check, by = c("speaker", "language"))

d_uttr <- d_uttr %>%
  mutate(
    weight = case_when(
      language == "American English" ~ transf(1),
      TRUE ~ weight
    )
  )

d_seg <- d_seg %>%
  left_join(df_check, by = c("speaker", "language"))

d_seg <- d_seg %>%
  mutate(
    weight = case_when(
      language == "American English" ~ transf(1),
      TRUE ~ weight
    )
  )
```

## Segment frequencies

Get the frequency of different segment types

```{r}
included_uttr <- unique(d_uttr$uttr_id)

d_c_freq <- d %>%
  dplyr::filter(uttr_id %in% included_uttr) %>%
  group_by(language, c_type) %>%
  count() %>%
  dplyr::filter(c_type != "NA") %>%
  group_by(language) %>%
  mutate(
    prop = n / sum(n)
  )

d_c_freq %>%
  mutate(
    c_type = fct_relevel(c_type,
    "syllabic C", "stop", "affricate", "fricative", "liquid", "nasal", "approximant", "non-high vowel", "high-vowel"
    )
    #"high vowel", "non-high vowel", "approximant", "nasal", "liquid", "fricative", "affricate", "stop", "syllabic C")
  ) %>%
  ggplot(aes(x = c_type, y = prop)) +
  geom_bar(stat = "identity") +
  facet_wrap(. ~ language) +
  coord_flip()
```

Get numbers for paper

```{r paper}
# get some numbers for the paper
d_uttr %>%
  group_by(language) %>%
  summarize(
    N = length(unique(uttr_id)),
    DUR = sum(uttr_dur)/60/60
  )

unique(d_uttr$speaker)
```

# Main analysis

Exploratory plots!

```{r}
ggplot(d_uttr, aes(x=segment_dur_uttr)) +
  facet_wrap(~language) +
  geom_point(aes(y=vowel_dur_uttr), col="orange", alpha=0.3) +
  geom_point(aes(y=cons_dur_uttr), col="purple", alpha=0.3) +
  geom_smooth(aes(y=vowel_dur_uttr), col="darkorange2", method="gam", se=F) +
  geom_smooth(aes(y=cons_dur_uttr), col="purple4", method="gam", se=F) +
  ylab("average duration in s") +
  scale_x_log10() +
  scale_y_log10()
```

GAMMs!

```{r}
library(mgcv)
library(itsadug)

d_seg <- d_seg %>%
  filter(language != "Kapampangan")

d_seg$segment_type_o <- as.ordered(d_seg$segment_type)
contrasts(d_seg$segment_type_o) <- "contr.treatment"

d_seg$speaker_f <- as.factor(d_seg$speaker)
d_seg$speakerSegtype <- interaction(d_seg$speaker, d_seg$segment_type)

d_seg$language_f <- as.factor(d_seg$language)
d_seg$languageSegtype <- interaction(d_seg$language, d_seg$segment_type)

d_seg$log_segment_dur_uttr <- log(d_seg$segment_dur_uttr)
d_seg$log_dur <- log(d_seg$phone_dur)

d_seg$c_type_f <- as.factor(d_seg$c_type)

# model 1: no control for quality

cv_mod_baseline <- bam(phone_dur ~
  segment_type_o +
  s(segment_dur_uttr, bs = "cr") +
  s(segment_dur_uttr, by = segment_type_o, bs = "cr") +
  #s(segment_dur_uttr, c_type_f, bs = "fs", m = 1, xt = list("cr")) +
  s(segment_dur_uttr, speakerSegtype, bs = "fs", m = 1, xt = list("cr")) +
  s(segment_dur_uttr, languageSegtype, bs = "fs", m = 1, xt = list("cr")),
  data = d_seg,
  #weights=d_seg$weight / mean(d_seg$weight),
  discrete = TRUE,
  nthreads = 8
)
saveRDS(cv_mod_baseline, "cv_mod_baseline.rds")
cv_mod_baseline <- readRDS("cv_mod_baseline.rds")
summary(cv_mod_baseline)

# model 2: weighted regression based on olivia's ratings

cv_mod_weighted <- bam(phone_dur ~
  segment_type_o +
  s(segment_dur_uttr, bs = "cr") +
  s(segment_dur_uttr, by = segment_type_o, bs = "cr") +
  #s(segment_dur_uttr, c_type_f, bs = "fs", m = 1, xt = list("cr")) +
  s(segment_dur_uttr, speakerSegtype, bs = "fs", m = 1, xt = list("cr")) +
  s(segment_dur_uttr, languageSegtype, bs = "fs", m = 1, xt = list("cr")),
  data = d_seg,
  weights = d_seg$weight / mean(d_seg$weight),
  discrete = TRUE,
  nthreads = 8
)
saveRDS(cv_mod_weighted, "cv_mod_weighted.rds")
cv_mod_weighted <- readRDS("cv_mod_weighted.rds")

# model 3: weighted + control for segment type

cv_mod_full <- bam(phone_dur ~
  segment_type_o +
  s(segment_dur_uttr, bs = "cr") +
  s(segment_dur_uttr, by = segment_type_o, bs = "cr") +
  s(segment_dur_uttr, c_type_f, bs = "fs", m = 1, xt = list("cr")) +
  s(segment_dur_uttr, speakerSegtype, bs = "fs", m = 1, xt = list("cr")) +
  s(segment_dur_uttr, languageSegtype, bs = "fs", m = 1, xt = list("cr")),
  data = d_seg,
  weights = d_seg$weight / mean(d_seg$weight),
  discrete = TRUE,
  nthreads = 8
)
saveRDS(cv_mod_full, "cv_mod_full.rds")
cv_mod_full <- readRDS("cv_mod_full.rds")
```

## Prediction plots: languages

```{r}
# d_uttr_plot <- d_uttr %>%
#   pivot_longer(c(vowel_dur_uttr, cons_dur_uttr),
#                names_to="measure",
#                values_to="value")

d_uttr_plot <- d_seg %>%
  group_by(language, speaker, uttr_id) %>%
  summarise(
    # duration of segmental material only
    uttr_dur = sum(phone_dur),
    num_syllables_uttr = sum(segment_type=="vowel" | c_type=="syllabic C"),
    num_segments_uttr = n(),
    segment_dur_uttr = uttr_dur / num_segments_uttr,
    syllable_dur_uttr = uttr_dur / num_syllables_uttr,
    vowel_dur_uttr = mean(phone_dur[segment_type=="vowel"]),
    cons_dur_uttr = mean(phone_dur[segment_type=="consonant"]),
  ) %>%
  ungroup() %>%
  pivot_longer(c(vowel_dur_uttr, cons_dur_uttr),
               names_to="measure",
               values_to="value")

```

Make C/V duration plots.

```{r}
# model predictions
newdat_cv <- expand.grid(
  segment_dur_uttr = seq(
    min(d_seg$segment_dur_uttr),
    max(d_seg$segment_dur_uttr),
    length.out = 100
  ),
  c_type_f=d_seg$c_type_f[1],
  speakerSegtype = d_seg$speakerSegtype[1],
  #language_f=unique(d_uttr_cv$language_f),
  languageSegtype = unique(d_seg$languageSegtype)
)

newdat_cv$segment_type_o <- str_extract(newdat_cv$languageSegtype, "[^.]*$")
newdat_cv$segment_type_o <- as.ordered(newdat_cv$segment_type_o)
contrasts(newdat_cv$segment_type_o) <- "contr.treatment"

newdat_cv$language <- str_extract(newdat_cv$languageSegtype, "^[^.]*")
newdat_cv$language_f <- as.factor(newdat_cv$language)
newdat_cv$speaker_f <- as.factor(str_extract(newdat_cv$speakerSegtype, "^[^.]*"))
#newdat_cv$speaker_f <- d_uttr_cv$speaker_f[1]

preds_baseline <- predict(cv_mod_baseline, newdat_cv,
  # the following is used when predicting without speakers
  exclude=c("s(segment_dur_uttr,speakerSegtype)",
            "s(segment_dur_uttr,speaker_f)",
            "s(segment_dur_uttr,speaker_f):segment_type_ovowel_dur_uttr",
            "s(segment_dur_uttr,c_type_f)"),
  se.fit = TRUE
)
preds_weighted <- predict(cv_mod_weighted, newdat_cv,
  # the following is used when predicting without speakers
  exclude=c("s(segment_dur_uttr,speakerSegtype)",
            "s(segment_dur_uttr,speaker_f)",
            "s(segment_dur_uttr,speaker_f):segment_type_ovowel_dur_uttr",
            "s(segment_dur_uttr,c_type_f)"),
  se.fit = TRUE
)
preds_full <- predict(cv_mod_full, newdat_cv,
  # the following is used when predicting without speakers
  exclude=c("s(segment_dur_uttr,speakerSegtype)",
            "s(segment_dur_uttr,speaker_f)",
            "s(segment_dur_uttr,speaker_f):segment_type_ovowel_dur_uttr",
            "s(segment_dur_uttr,c_type_f)"),
  se.fit = TRUE
)

# baseline

newdat_cv_baseline <- newdat_cv

newdat_cv_baseline$dur <- preds_baseline$fit
newdat_cv_baseline$ul <- preds_baseline$fit + preds_baseline$se.fit * 1.96
newdat_cv_baseline$ll <- preds_baseline$fit - preds_baseline$se.fit * 1.96

newdat_cv_baseline$model <- "cv_mod_baseline"

# weighted

newdat_cv_weighted <- newdat_cv

newdat_cv_weighted$dur <- preds_weighted$fit
newdat_cv_weighted$ul <- preds_weighted$fit + preds_weighted$se.fit * 1.96
newdat_cv_weighted$ll <- preds_weighted$fit - preds_weighted$se.fit * 1.96

newdat_cv_weighted$model <- "cv_mod_weighted"

# full

newdat_cv_full <- newdat_cv

newdat_cv_full$dur <- preds_full$fit
newdat_cv_full$ul <- preds_full$fit + preds_full$se.fit * 1.96
newdat_cv_full$ll <- preds_full$fit - preds_full$se.fit * 1.96

newdat_cv_full$model <- "cv_mod_full"

newdat_all <- bind_rows(
  newdat_cv_baseline,
  newdat_cv_weighted,
  newdat_cv_full
)
```


Plot!

```{r}
d_uttr_plot %>%
  ggplot(data = ., aes(x = segment_dur_uttr, y = value, col = measure)) +
  facet_wrap(. ~ language) +
  geom_point(alpha = 0.1, pch=16) +
  #geom_ribbon(data = filter(newdat_cv_all,
  #                          model == "dur",
  #                          segment_type_o == "vowel_dur_uttr"),
  #            aes(ymin = ll, ymax = ul), col = NA, fill = "grey",
  #            alpha = 0.5) +
  #geom_ribbon(data = filter(newdat_cv,
  #                          measure_type == "dur",
  #                          segment_type_o == "cons_dur_uttr"),
   #           aes(ymin = ll, ymax = ul), col = NA, fill = "grey",
   #           alpha = 0.5) +
  geom_line(data = filter(newdat_all,
                          segment_type_o == "vowel"),
            aes(lty=model, y=dur),
            col = "darkorange1", lwd = 1) +
  geom_line(data = filter(newdat_all,
                          segment_type_o == "consonant"),
            aes(lty=model, y=dur),
            col = "purple4", lwd = 1) +
  scale_x_log10(breaks = c(0.05, 0.07, 0.1, 0.14)) +
  scale_y_log10(breaks = c(0.05, 0.07, 0.1, 0.14),
                limits = c(0.03, 0.19)) +
  #scale_x_continuous(limits=c(0.035,0.145), 
  #                   breaks=seq(0.04,0.14,0.02),
  #                   labels=c(".04",".06",".08",".10",".12",".14")) +
  scale_colour_manual(values = c("purple", "orange"), guide = "none") +
  #scale_y_continuous(breaks=seq(0.04,0.20,0.02)) +
    # labels for axes
  xlab("Average segment duration in s (inverse speech rate)") +
  labs(y = "Average C vs. V duration in s") +
  #ylab("Average V vs. C duration in s") +
  # remove clutter
  theme_minimal()
```

We now obtain a measure of effect strength: the difference in the log-log derivative at the median speech rate / language. Based on model predictions!

```{r}
eff_sizes <- list()

# model predictions
newdat_deriv <- expand.grid(
  speakerSegtype=d_seg$speakerSegtype[1],
  languageSegtype=unique(d_seg$languageSegtype),
  measure_type="dur",
  c_type_f=d_seg$c_type_f[1]
)

newdat_deriv$segment_type_o <- str_extract(newdat_deriv$languageSegtype,
                                        "[^.]*$")
newdat_deriv$segment_type_o <- as.ordered(newdat_deriv$segment_type_o)
contrasts(newdat_deriv$segment_type_o) <- "contr.treatment"

newdat_deriv$language <- str_extract(newdat_deriv$languageSegtype,
                                        "^[^.]*")
newdat_deriv$language_f <- as.factor(newdat_deriv$language)
newdat_deriv$speaker_f <- d_uttr_cv$speaker_f[1]

language_med_durs <- d_seg %>%
  group_by(language) %>%
  summarise(segment_dur_uttr=median(segment_dur_uttr)) %>%
  ungroup()

newdat_deriv <- left_join(
  newdat_deriv,
  language_med_durs,
  by="language"
)


eps <- 0.001

newdat_deriv_1 <- newdat_deriv %>%
  mutate(segment_dur_uttr = exp(log(segment_dur_uttr) - (eps/2)))
newdat_deriv_2 <- newdat_deriv %>%
  mutate(segment_dur_uttr = exp(log(segment_dur_uttr) + (eps/2)))

mods <- c("cv_mod_baseline", "cv_mod_weighted", "cv_mod_full")
for (m in mods) {
X0 <- predict(get(m), 
              newdat_deriv_1,
              exclude=c(
                "s(segment_dur_uttr,speakerSegtype)",
                "s(segment_dur_uttr,speaker_f)",
                "s(segment_dur_uttr,speaker_f):segment_type_ovowel_dur_uttr",
                "s(segment_dur_uttr,c_type_f)"
              )
) %>% log()
X1 <- predict(get(m), 
              newdat_deriv_2,
              exclude=c(
                "s(segment_dur_uttr,speakerSegtype)",
                "s(segment_dur_uttr,speaker_f)",
                "s(segment_dur_uttr,speaker_f):segment_type_ovowel_dur_uttr",
                "s(segment_dur_uttr,c_type_f)"
              )
) %>% log()

# finite difference approximation of first derivative
# the design matrix
first_deriv <- (X1 - X0) / eps
#first_deriv <- Xp %*% coef(get(m))

newdat_deriv$derivative <- first_deriv
newdat_deriv$model <- m

eff_sizes[[m]] <- newdat_deriv %>%
  dplyr::select(language, segment_type_o, segment_dur_uttr, model, derivative) %>%
  group_by(language, segment_dur_uttr, model) %>%
  summarise(eff_size = derivative[segment_type_o == "vowel"] - derivative[segment_type_o == "consonant"]) %>%
  ungroup() %>%
  arrange(desc(eff_size))
}

eff_sizes_df <- bind_rows(eff_sizes)
eff_sizes_df
```

We now order by effect size & show this number in the plot.

```{r}
for (m in mods) {
  
  eff_sizes_m <- eff_sizes[[m]]
  eff_sizes_m$language <- factor(eff_sizes_m$language, levels=eff_sizes_m$language)
  eff_sizes_m$segment_dur_uttr <- 0.05
  eff_sizes_m$value <- 0.13
  
  newdat_m <- filter(newdat_all, model==m)
  newdat_m$language <- factor(newdat_m$language, levels=eff_sizes_m$language)
  
  plot_dat <- d_uttr_plot
  plot_dat$language <- factor(plot_dat$language, levels=eff_sizes_m$language)
  
  plot_dat %>%
    ggplot(data = ., aes(x = segment_dur_uttr, y = value, col = measure)) +
    facet_wrap(. ~ language) +
    geom_point(alpha = 0.1, pch=16) +
    #geom_ribbon(data = filter(newdat_cv_all,
    #                          model == "dur",
    #                          segment_type_o == "vowel_dur_uttr"),
    #            aes(ymin = ll, ymax = ul), col = NA, fill = "grey",
    #            alpha = 0.5) +
    #geom_ribbon(data = filter(newdat_cv,
    #                          measure_type == "dur",
    #                          segment_type_o == "cons_dur_uttr"),
    #            aes(ymin = ll, ymax = ul), col = NA, fill = "grey",
    #            alpha = 0.5) +
    geom_line(data = filter(newdat_m,
                            segment_type_o == "vowel"),
              aes(lty=model, y=dur),
              col = "darkorange1", lwd = 1) +
    geom_line(data = filter(newdat_m,
                            segment_type_o == "consonant"),
              aes(lty=model, y=dur),
              col = "purple4", lwd = 1) +
    geom_text(data=eff_sizes_m, 
              aes(label=paste0("Δβ = ", round(eff_size,2))),
              col="black", size=5) +
    scale_x_log10(breaks = c(0.05, 0.07, 0.1, 0.14)) +
    scale_y_log10(breaks = c(0.05, 0.07, 0.1, 0.14),
                  limits = c(0.03, 0.19)) +
    scale_colour_manual(values = c("purple", "orange"), guide = "none") +
    xlab("Average segment duration in s (inverse speech rate)") +
    labs(y = "Average C vs. V duration in s") +
    theme_minimal() +
    theme(panel.grid = element_blank(),
        axis.line = element_line(size=0.8),
        axis.ticks = element_line(size=0.8),
        axis.text = element_text(size=16, colour="black"),
        axis.title = element_text(size=18, face="bold"),
        strip.background.y = element_blank(), 
        strip.placement = "outside",
        strip.text=element_text(size=18, face="bold"))
  ggsave(paste0("graphs_icphs/", m, ".png"), width=9, height=9, dpi=300)
}
```

## Recovering consonant type effect from full model

```{r}
# model predictions
newdat_ctype <- expand.grid(
  segment_dur_uttr = seq(
    min(d_seg$segment_dur_uttr),
    max(d_seg$segment_dur_uttr),
    length.out = 100
  ),
  c_type_f=unique(d_seg$c_type_f),
  speakerSegtype = d_seg$speakerSegtype[1], # will be set to 0
  languageSegtype = unique(d_seg$languageSegtype) # will be set to 0
) %>% filter(c_type_f != "syllabic C")

newdat_ctype$segment_type_o <- ifelse(
  newdat_ctype$c_type_f %in% c("non-high vowel", "high vowel"),
  "vowel",
  "consonant"
  )
newdat_ctype$segment_type_o <- as.ordered(newdat_ctype$segment_type_o)
contrasts(newdat_ctype$segment_type_o) <- "contr.treatment"

preds_ctype <- predict(cv_mod_full, newdat_ctype,
  # the following is used when predicting without speakers
  exclude=c("s(segment_dur_uttr,speakerSegtype)",
            "s(segment_dur_uttr,speaker_f)",
            "s(segment_dur_uttr,speaker_f):segment_type_ovowel_dur_uttr",
            "s(segment_dur_uttr,languageSegtype)"),
  se.fit = TRUE
)

newdat_ctype$dur <- preds_ctype$fit
newdat_ctype$ul <- preds_ctype$fit + preds_ctype$se.fit * 1.96
newdat_ctype$ll <- preds_ctype$fit - preds_ctype$se.fit * 1.96

newdat_ctype$segment_type_broad <- 
  case_when(
    newdat_ctype$c_type_f %in% c("non-high vowel", "high vowel", "approximant") ~ "vocalic",
    newdat_ctype$c_type_f %in% c("liquid", "nasal", "fricative") ~ "continuant + nasal",
    newdat_ctype$c_type_f %in% c("affricate", "stop") ~ "non-continuant",
  )
newdat_ctype$segment_type_broad <- 
  factor(newdat_ctype$segment_type_broad,
         levels=c("vocalic", "continuant + nasal", "non-continuant"))
```


Plot!

```{r}
ggplot(newdat_ctype, aes(x=segment_dur_uttr, y=dur, col=c_type_f)) +
  facet_wrap(~segment_type_broad) +
  geom_ribbon(aes(ymin=ll,ymax=ul, group=c_type_f), col=NA, fill="grey",
              alpha=0.3) +
  geom_textline(aes(label = c_type_f), size = 5, vjust = 0.2,
                linewidth = 1) +
  scale_x_continuous(breaks=c(-3.5,-3,-2.5,-2),
                     labels=round(exp(c(-3.5,-3,-2.5,-2)),2)) +
  scale_y_continuous(breaks=c(-3.5,-3,-2.5,-2),
                     labels=round(exp(c(-3.5,-3,-2.5,-2)),2)) +
  scale_colour_brewer(palette = "Set1", guide='none') +
  #scale_colour_manual(values=c("purple4", "darkorange2"), guide="none") +
  xlab("Overall average segment duration in s (inverse speech rate)") +
  labs(y=NULL) +
  ylab("Average duration per segment type") +
  # remove clutter
  theme_minimal() +
  # some specific formatting...
  theme(panel.grid = element_blank(),
        axis.line = element_line(size=0.8),
        axis.ticks = element_line(size=0.8),
        axis.text = element_text(size=14, colour="black"),
        axis.title = element_text(size=16, face="bold"),
        strip.background.y = element_blank(), 
        strip.placement = "outside",
        strip.text=element_text(size=16, face="bold"))
#ggsave("graphs_icphs/ctype_mod_preds.png", width=7.5, height=4, dpi=300)

```

The results here are a little odd... Some model tweaking might help:

(1) Try removing syllabic C's from the data, which are sparse and could be driving high edf
(2) Try bs="sz"? Fixed effect smooths that represent a deviation from main effect, with main effect "forced to do as much of the work as possible"

## Inter-subject variation

Three different dummy data sets, one for each model. We only want to calculate derivative at L-specific median value for each speaker.

```{r}
# model predictions
newdat_ind_deriv <- d_seg %>%
  dplyr::select(speakerSegtype, speaker, languageSegtype, language) %>%
  unique() %>%
  arrange(language, speakerSegtype) %>%
  mutate(c_type_f=d_seg$c_type_f[1],
         segment_type_o=str_extract(languageSegtype, "[^.]*$") %>% as.ordered()
         )

contrasts(newdat_ind_deriv$segment_type_o) <- "contr.treatment"

ind_med_durs <- d_seg %>%
  group_by(language, speaker) %>%
  summarise(segment_dur_uttr=median(segment_dur_uttr)) %>%
  ungroup()

newdat_ind_deriv <- left_join(
  newdat_ind_deriv,
  ind_med_durs,
  by=c("language", "speaker")
)


eps <- 0.001

newdat_ind_deriv_1 <- newdat_ind_deriv %>%
  mutate(segment_dur_uttr = exp(log(segment_dur_uttr) - (eps/2)))
newdat_ind_deriv_2 <- newdat_ind_deriv %>%
  mutate(segment_dur_uttr = exp(log(segment_dur_uttr) + (eps/2)))

mods <- c("cv_mod_baseline", "cv_mod_weighted", "cv_mod_full")
ind_eff_sizes <- list()

for (m in mods) {
X0 <- predict(get(m), 
              newdat_ind_deriv_1,
              exclude=c(
                "s(segment_dur_uttr,c_type_f)"
              )
) %>% log()
X1 <- predict(get(m), 
              newdat_ind_deriv_2,
              exclude=c(
                "s(segment_dur_uttr,c_type_f)"
              )
) %>% log()

# finite difference approximation of first derivative
# the design matrix
first_deriv <- (X1 - X0) / eps
#first_deriv <- Xp %*% coef(get(m))

newdat_ind_deriv$derivative <- first_deriv
newdat_ind_deriv$model <- m

ind_eff_sizes[[m]] <- newdat_ind_deriv %>%
  dplyr::select(language, speaker, segment_type_o, segment_dur_uttr, model, derivative) %>%
  group_by(language, speaker, segment_dur_uttr, model) %>%
  summarise(eff_size = derivative[segment_type_o=="vowel"] - derivative[segment_type_o=="consonant"]) %>%
  ungroup() %>%
  arrange(desc(eff_size))
}

ind_eff_sizes_df <- bind_rows(ind_eff_sizes)
#eff_sizes_df
```

Plot histograms (or density plots?) of by-speaker deviations by language across three model types.

```{r}
ind_eff_sizes_df$model <- factor(ind_eff_sizes_df$model,
                                 levels=c(
                                   "cv_mod_baseline",
                                   "cv_mod_weighted",
                                   "cv_mod_full"
                                  )
                                 )

ind_eff_sizes_df %>%
  ggplot(aes(x=eff_size, col=model, lty=model)) +
  #facet_wrap(~language) +
  geom_density()

ind_eff_sizes_df %>%
  ggplot(aes(x=eff_size, col=model, lty=model)) +
  facet_wrap(~language) +
  geom_density()

ind_eff_sizes_df %>%
  group_by(model) %>%
  summarise(eff_sd = sd(eff_size)) %>%
  ungroup()

ind_eff_sizes_df %>%
  group_by(language, model) %>%
  summarise(eff_sd = sd(eff_size)) %>%
  ungroup()
```
