---
title: "doreco_stretchiness_analysis"
author: "Noah Luntzlara"
date: "2025-10-01"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Set Up

Load packages.

```{r}
library(tidyverse)
# library(ggbeeswarm)
# library(ggpubr)
library(forcats)

library(mgcv)
library(itsadug)

library(geomtextpath)
library(RColorBrewer)
library(viridis)

library(broom)
```

Read in the data.

```{r}

d_all <- read_csv("../doreco_data/combined_doreco_durations_filtered.csv")

# d_all_unfilt <- read_csv("../doreco_data/combined_doreco_durations_unfiltered.csv")

```

Data frame for CV models

```{r}

d_cv <- d_all %>%
  filter(phone_type != "syllabic_consonant")

d_cv$segtype_o <- as.ordered(d_cv$segtype)
contrasts(d_cv$segtype_o) <- "contr.treatment"

d_cv$speaker_f <- as.factor(d_cv$speaker)
d_cv$speakerSegtype <- interaction(d_cv$speaker, d_cv$segtype)

d_cv$language_f <- as.factor(d_cv$language)
d_cv$languageSegtype <- interaction(d_cv$language, d_cv$segtype)

d_cv$phone_type_f <- as.factor(d_cv$phone_type)

```


# GAMMs for CV comparisons

## Fitting CV models

```{r}
# model 1: consonants vs vowels phone duration by utterance

cv_mod_baseline <- bam(phone_dur ~
  segtype_o +
  s(phone_dur_uttr, bs = "cr") +
  s(phone_dur_uttr, by = segtype_o, bs = "cr") +
  #s(phone_dur_uttr, phone_type_f, bs = "fs", m = 1, xt = list("cr")) +
  s(phone_dur_uttr, speakerSegtype, bs = "fs", m = 1, xt = list("cr")) +
  s(phone_dur_uttr, languageSegtype, bs = "fs", m = 1, xt = list("cr")),
  data = d_cv,
  #weights=d_cv$weight / mean(d_cv$weight),
  discrete = TRUE,
  nthreads = 8
)

saveRDS(cv_mod_baseline, "cv_mod_baseline.rds")
cv_mod_baseline <- readRDS("cv_mod_baseline.rds")
summary(cv_mod_baseline)

# model 2: consonants vs vowels phone duration by utterance w/ phone type

cv_mod_type <- bam(phone_dur ~
  segtype_o +
  s(phone_dur_uttr, bs = "cr") +
  s(phone_dur_uttr, by = segtype_o, bs = "cr") +
  s(phone_dur_uttr, phone_type_f, bs = "fs", m = 1, xt = list("cr")) +
  s(phone_dur_uttr, speakerSegtype, bs = "fs", m = 1, xt = list("cr")) +
  s(phone_dur_uttr, languageSegtype, bs = "fs", m = 1, xt = list("cr")),
  data = d_cv,
  #weights=d_cv$weight / mean(d_cv$weight),
  discrete = TRUE,
  nthreads = 8
)

saveRDS(cv_mod_type, "cv_mod_type.rds")
cv_mod_type <- readRDS("cv_mod_type.rds")
summary(cv_mod_type)
```

Just reading the models from files

```{r}
cv_mod_baseline <- readRDS("cv_mod_baseline.rds")
cv_mod_type <- readRDS("cv_mod_type.rds")
```


## Model Predictions for CV Models

```{r}
# model predictions for models with with segment inclusive speech rate measures
newdat_cv <- expand.grid(
  phone_dur_uttr = seq(
    min(d_cv$phone_dur_uttr),
    max(d_cv$phone_dur_uttr),
    length.out = 100
  ),
  phone_type_f=d_cv$phone_type_f[1],
  #phone_type_new_f=d_cv$phone_type_new_f[1],
  speakerSegtype = d_cv$speakerSegtype[1],
  languageSegtype = unique(d_cv$languageSegtype)
)

newdat_cv$segtype_o <- str_extract(newdat_cv$languageSegtype, "[^.]*$")
newdat_cv$segtype_o <- as.ordered(newdat_cv$segtype_o)
contrasts(newdat_cv$segtype_o) <- "contr.treatment"

newdat_cv$language <- str_extract(newdat_cv$languageSegtype, "^[^.]*")
newdat_cv$language_f <- as.factor(newdat_cv$language)
newdat_cv$speaker_f <- as.factor(str_extract(newdat_cv$speakerSegtype, "^[^.]*"))
#newdat_cv$speaker_f <- newdat_cv$speaker_f[1]

preds_baseline <- predict(cv_mod_baseline, newdat_cv,
  # the following is used when predicting without speakers
  exclude=c("s(phone_dur_uttr,speakerSegtype)",
            "s(phone_dur_uttr,speaker_f)",
            "s(phone_dur_uttr,speaker_f):segtype_ovowel_dur_uttr",
            "s(phone_dur_uttr,phone_type_f)"),
  se.fit = TRUE
)

preds_type <- predict(cv_mod_type, newdat_cv,
  # the following is used when predicting without speakers
  exclude=c("s(phone_dur_uttr,speakerSegtype)",
            "s(phone_dur_uttr,speaker_f)",
            "s(phone_dur_uttr,speaker_f):segtype_ovowel_dur_uttr",
            "s(phone_dur_uttr,phone_type_f)"),
  se.fit = TRUE
)

# baseline

newdat_cv_baseline <- newdat_cv

newdat_cv_baseline$dur <- preds_baseline$fit
newdat_cv_baseline$ul <- preds_baseline$fit + preds_baseline$se.fit * 1.96
newdat_cv_baseline$ll <- preds_baseline$fit - preds_baseline$se.fit * 1.96

newdat_cv_baseline$model <- "cv_mod_baseline"

# controlling for phone type

newdat_cv_type <- newdat_cv

newdat_cv_type$dur <- preds_type$fit
newdat_cv_type$ul <- preds_type$fit + preds_type$se.fit * 1.96
newdat_cv_type$ll <- preds_type$fit - preds_type$se.fit * 1.96

newdat_cv_type$model <- "cv_mod_type"


newdat_cv_all <- bind_rows(
  newdat_cv_baseline,
  newdat_cv_type
)

newdat_cv_all %>% write_csv("./newdat_cv_all.csv")
newdat_cv_all <- read_csv("./newdat_cv_all.csv")
```


## Plotting CV Model Predictions

New data with C and V averages by utterance

```{r}
d_uttr_cv <- d_cv %>%
  group_by(language,speaker,uttr_id) %>%
  summarize(vowel_dur_uttr = mean(phone_dur[segtype == "syllabic"]),
            cons_dur_uttr = mean(phone_dur[segtype == "consonant"]),
            phone_dur_uttr = phone_dur_uttr[1]
  ) %>%
  pivot_longer(c(vowel_dur_uttr, cons_dur_uttr),
               names_to="measure",
               values_to="value")
```

Plot CV model predictions

```{r}
newdat_cv_all %>%
  filter(language %in% d_uttr_cv$language) %>%
  ggplot(data = ., aes(x = phone_dur_uttr)) +
  facet_wrap(. ~ language) +
  geom_point(data = d_uttr_cv,
             aes(y = value, col=measure), alpha = 0.1, pch=16) +
  geom_line(aes(y = dur, lty=model, col = segtype_o),, lwd = 1) +
  # geom_ribbon(data = filter(newdat_cv_all,
  #                          model == "cv_mod_baseline",
  #                          segtype_o == "consonant"),
  #           aes(ymin = ll, ymax = ul), col = NA, fill = "skyblue",
  #           alpha = 0.5) +
  # geom_ribbon(data = filter(newdat_cv_all,
  #                          model == "cv_mod_baseline",
  #                          segtype_o == "syllabic"),
  #           aes(ymin = ll, ymax = ul), col = NA, fill = "yellow",
  #           alpha = 0.5) +
  scale_color_manual(values = c("consonant" = "purple4", "syllabic" = "darkorange",
                                "cons_dur_uttr" = "purple", "vowel_dur_uttr" = "orange"),
                     guide = "none") +
  scale_x_log10(breaks = c(0.05, 0.07, 0.1, 0.14)) +
  scale_y_log10(breaks = c(0.05, 0.07, 0.1, 0.14),
                #limits = c(0.03, 0.2)
                ) +
  #scale_x_continuous(limits=c(0.035,0.145), 
  #                   breaks=seq(0.04,0.14,0.02),
  #                   labels=c(".04",".06",".08",".10",".12",".14")) +
  #scale_y_continuous(breaks=seq(0.04,0.20,0.02)) +
  # labels for axes
  labs(x = "Average phone duration by utterance",
       y = "Average C vs. V duration in s") +
  # remove clutter
  theme_minimal() +
  theme(panel.grid = element_blank(),
        axis.line = element_line(size=0.8),
        axis.ticks = element_line(size=0.8),
        axis.text = element_text(size=16, colour="black"),
        axis.title = element_text(size=18, face="bold"),
        strip.background.y = element_blank(), 
        strip.placement = "outside",
        strip.text=element_text(size=18, face="bold"))
  ggsave("doreco_graphs_v1/cv/cv_mod_all.png", width=15, height=15, dpi=300)
```

# CV Model Phone Type Comparisons

## CV Model Predictions for Phone Type

CV Model predictions for phone type

```{r}
# model predictions
newdat_phonetype <- expand.grid(
  phone_dur_uttr = seq(
    min(d_cv$phone_dur_uttr),
    max(d_cv$phone_dur_uttr),
    length.out = 100
  ),
  phone_type_f=unique(d_cv$phone_type_f),
  speakerSegtype = d_cv$speakerSegtype[1], # will be set to 0
  languageSegtype = unique(d_cv$languageSegtype) # will be set to 0
) #%>% filter(phone_type_f != "syllabic C")

newdat_phonetype$segtype_o <- ifelse(
  newdat_phonetype$phone_type_f %in% c("vowel_diph",
                                         "vowel_mono_long_high",
                                         "vowel_mono_long_nonhigh",
                                         "vowel_mono_short_high",
                                         "vowel_mono_short_nonhigh"),
  "syllabic",
  "consonant"
  )
newdat_phonetype$segtype_o <- as.ordered(newdat_phonetype$segtype_o)
contrasts(newdat_phonetype$segtype_o) <- "contr.treatment"

preds_ctype <- predict(cv_mod_type, newdat_phonetype,
  # the following is used when predicting without speakers
  exclude=c("s(phone_dur_uttr,speakerSegtype)",
            "s(phone_dur_uttr,speaker_f)",
            "s(phone_dur_uttr,speaker_f):segtype_ovowel_dur_uttr",
            "s(phone_dur_uttr,languageSegtype)"),
  se.fit = TRUE
)

newdat_phonetype$dur <- preds_ctype$fit
newdat_phonetype$ul <- preds_ctype$fit + preds_ctype$se.fit * 1.96
newdat_phonetype$ll <- preds_ctype$fit - preds_ctype$se.fit * 1.96

newdat_phonetype$segtype_broad <- 
  case_when(
    newdat_phonetype$phone_type_f %in% c("vowel_diph",
                                         "vowel_mono_long_high",
                                         "vowel_mono_long_nonhigh",
                                         "vowel_mono_short_high",
                                         "vowel_mono_short_nonhigh",
                                         "approximant",
                                         "syllabic_consonant") ~ "vocalic",
    newdat_phonetype$phone_type_f %in% c("liquid", "nasal", "fricative", "nasal_double") ~ "continuant + nasal",
    newdat_phonetype$phone_type_f %in% c("plosive", "plosive_double", "affricate", "flap", "glottal", "trill", "click", "click_contour") ~ "non-continuant"
  )
newdat_phonetype$segtype_broad <- 
  factor(newdat_phonetype$segtype_broad,
         levels=c("vocalic", "continuant + nasal", "non-continuant", "complex segment"))
```

## Plotting CV Model Phone Type Effects

Plotting phone type effects

```{r}
mycolors = c(brewer.pal(name="Dark2", n = 8), brewer.pal(name="Paired", n = 8))

ggplot(newdat_phonetype, aes(x=phone_dur_uttr, y=dur, col=phone_type_f)) +
  facet_wrap(~segtype_broad) +
  geom_ribbon(aes(ymin=ll,ymax=ul, group=phone_type_f), col=NA, fill="grey",
              alpha=0.3) +
  geom_textline(aes(label = phone_type_f), size = 5, vjust = 0.2,
                linewidth = 1) +
  scale_x_continuous(breaks=c(-3.5,-3,-2.5,-2),
                     labels=round(exp(c(-3.5,-3,-2.5,-2)),2)) +
  scale_y_continuous(breaks=c(-3.5,-3,-2.5,-2),
                     labels=round(exp(c(-3.5,-3,-2.5,-2)),2)) +
  #scale_colour_brewer(palette = "Set3", guide='none') +
  scale_color_manual(values = mycolors) +
  #scale_colour_manual(values=c("purple4", "darkorange2"), guide="none") +
  xlab("Overall average segment duration in s (inverse speech rate)") +
  labs(y=NULL) +
  ylab("Average duration per segment type") +
  # remove clutter
  theme_minimal() +
  # some specific formatting...
  theme(panel.grid = element_blank(),
        axis.line = element_line(size=0.8),
        axis.ticks = element_line(size=0.8),
        axis.text = element_text(size=14, colour="black"),
        axis.title = element_text(size=16, face="bold"),
        strip.background.y = element_blank(), 
        strip.placement = "outside",
        strip.text=element_text(size=16, face="bold"))
ggsave("doreco_graphs_v1/phonetype_mod_preds.png", width=10, height=6, dpi=300)

```

# CV Model Effect Sizes at Medians

## Computing CV Model Effect Sizes at Medians
Measuring and comparing derivatives at the medians for CV models

```{r}
eff_sizes_cv <- list()

# model predictions
newdat_cv_deriv <- expand.grid(
  phone_type_f=d_cv$phone_type_f[1],
  #phone_type_new_f=d_cv$phone_type_new_f[1],
  speakerSegtype = d_cv$speakerSegtype[1],
  #language_f=unique(d_cv$language_f),
  languageSegtype = unique(d_cv$languageSegtype),
  measure_type = "dur"
)


newdat_cv_deriv$segtype_o <- str_extract(newdat_cv_deriv$languageSegtype,
                                        "[^.]*$")
newdat_cv_deriv$segtype_o <- as.ordered(newdat_cv_deriv$segtype_o)
contrasts(newdat_cv_deriv$segtype_o) <- "contr.treatment"

newdat_cv_deriv$language <- str_extract(newdat_cv_deriv$languageSegtype,
                                        "^[^.]*")
newdat_cv_deriv$language_f <- as.factor(newdat_cv_deriv$language)
newdat_cv_deriv$speaker_f <- as.factor(str_extract(newdat_cv_deriv$speakerSegtype, "^[^.]*"))
#newdat_cv_deriv$speaker_f <- newdat_cv$speaker_f[1]

language_med_durs <- d_cv %>%
  group_by(language) %>%
  summarise(phone_dur_uttr = median(phone_dur_uttr)) %>%
  ungroup()

newdat_cv_deriv <- left_join(
  newdat_cv_deriv,
  language_med_durs,
  by = "language"
)

eps <- 0.001

newdat_cv_deriv_1 <- newdat_cv_deriv %>%
  mutate(phone_dur_uttr = exp(log(phone_dur_uttr) - (eps/2)))
newdat_cv_deriv_2 <- newdat_cv_deriv %>%
  mutate(phone_dur_uttr = exp(log(phone_dur_uttr) + (eps/2)))

mods_cv <- c("cv_mod_baseline", "cv_mod_type"
             #, "cv_mod_type2"
             )

for (m in mods_cv) {
X1 <- predict(get(m), 
              newdat_cv_deriv_1,
              exclude=c(
                "s(phone_dur_uttr,speakerSegtype)",
                "s(phone_dur_uttr,speaker_f)",
                "s(phone_dur_uttr,speaker_f):segtype_ovowel_dur_uttr",
                "s(phone_dur_uttr,phone_type_f)"
                #"s(phone_dur_uttr,phone_type_new_f)"
              )
) %>% log()
X2 <- predict(get(m), 
              newdat_cv_deriv_2,
              exclude=c(
                "s(phone_dur_uttr,speakerSegtype)",
                "s(phone_dur_uttr,speaker_f)",
                "s(phone_dur_uttr,speaker_f):segtype_ovowel_dur_uttr",
                "s(phone_dur_uttr,phone_type_f)"
                #"s(phone_dur_uttr,phone_type_new_f)"
              )
) %>% log()

X0 <- predict(get(m), 
              newdat_cv_deriv,
              exclude=c(
                "s(phone_dur_uttr,speakerSegtype)",
                "s(phone_dur_uttr,speaker_f)",
                "s(phone_dur_uttr,speaker_f):segtype_ovowel_dur_uttr",
                "s(phone_dur_uttr,phone_type_f)"
                #"s(phone_dur_uttr,phone_type_new_f)"
              )
)

# finite difference approximation of first derivative
# the design matrix
first_deriv <- (X2 - X1) / eps
#first_deriv <- Xp %*% coef(get(m))



newdat_cv_deriv$derivative <- first_deriv
newdat_cv_deriv$model <- m

newdat_cv_deriv$phone_dur <- X0

eff_sizes_cv[[m]] <- newdat_cv_deriv %>%
  dplyr::select(language, segtype_o, phone_dur_uttr, phone_dur, model, derivative) %>%
  group_by(language, phone_dur_uttr, model) %>%
  summarise(derivative_c = derivative[segtype_o == "consonant"],
            derivative_v = derivative[segtype_o == "syllabic"],
            eff_size = derivative_v - derivative_c,
            phone_dur_v = phone_dur[segtype_o == "syllabic"],
            phone_dur_c = phone_dur[segtype_o == "consonant"],) %>%
  ungroup() %>%
  arrange(desc(eff_size))
}


eff_sizes_cv_df <- bind_rows(eff_sizes_cv)
eff_sizes_cv_df %>% write_csv("eff_sizes_cv_df.csv")
eff_sizes_cv_df <- read_csv("./eff_sizes_cv_df.csv")
```





## Making CV Prediction Plots Showing Effect Sizes at Medians

Ordering by effect size (delta-beta) & showing this number in the plot of CV predictions.

```{r}
for (m in mods_cv) {
  
  eff_sizes_cv_m <- eff_sizes_cv[[m]]
  eff_sizes_cv_m$language <- factor(eff_sizes_cv_m$language, levels=eff_sizes_cv_m$language)
  eff_sizes_cv_m$phone_dur_uttr_median <- eff_sizes_cv_m$phone_dur_uttr
  eff_sizes_cv_m$phone_dur_uttr <- 0.05
  eff_sizes_cv_m$value <- 0.13
  
  newdat_m <- filter(newdat_cv_all, model==m)
  newdat_m$language <- factor(newdat_m$language, levels=eff_sizes_cv_m$language)
  
  plot_dat <- d_uttr_cv
  plot_dat$language <- factor(plot_dat$language, levels=eff_sizes_cv_m$language)
  
  plot_dat %>%
    ggplot(data = ., aes(x = phone_dur_uttr, y = value, col = measure)) +
    facet_wrap(. ~ language) +
    geom_point(alpha = 0.1, pch=16) + 
    #geom_ribbon(data = filter(newdat_cv_all,
    #                          model == "dur",
    #                          segtype_o == "vowel_dur_uttr"),
    #            aes(ymin = ll, ymax = ul), col = NA, fill = "grey",
    #            alpha = 0.5) +
    #geom_ribbon(data = filter(newdat_cv,
    #                          measure_type == "dur",
    #                          segtype_o == "cons_dur_uttr"),
    #            aes(ymin = ll, ymax = ul), col = NA, fill = "grey",
    #            alpha = 0.5) +
    geom_line(data = filter(newdat_m,
                            segtype_o == "syllabic"),
              aes(lty=model, y=dur),
              col = "darkorange1", lwd = 1) +
    geom_line(data = filter(newdat_m,
                            segtype_o == "consonant"),
              aes(lty=model, y=dur),
              col = "purple4", lwd = 1) +
    geom_text(data=eff_sizes_cv_m, 
              aes(label=paste0("Δβ = ", round(eff_size,2))),
              col="black", size=5) +
    geom_text(data=eff_sizes_cv_m, 
              aes(y=value/1.2,
                  #x=phone_dur_uttr*2,
                  label=paste0("med = ", round(phone_dur_uttr_median,2))),
              col="black", size=5) +
    scale_x_log10(breaks = c(0.05, 0.07, 0.1, 0.14)) +
    scale_y_log10(breaks = c(0.05, 0.07, 0.1, 0.14),
                  limits = c(0.03, 0.19)) +
    scale_colour_manual(values = c("purple", "orange"), guide = "none") +
    labs(title = "Filtering by number of phones remaining (v1)",
         x = "Average segment duration in s (inverse speech rate)",
         y = "Average C vs. V duration in s") +
    theme_minimal() +
    theme(panel.grid = element_blank(),
        title = element_text(size=24, face="bold"),
        axis.line = element_line(size=0.8),
        axis.ticks = element_line(size=0.8),
        axis.text = element_text(size=16, colour="black"),
        axis.title = element_text(size=18, face="bold"),
        strip.background.y = element_blank(), 
        strip.placement = "outside",
        strip.text=element_text(size=18, face="bold"))
  ggsave(paste0("doreco_graphs_v1/cv/", m, ".png"), width=15, height=15, dpi=300)
}
```

## Comparing CV Model Effect Sizes at Medians

Comparing effect sizes at medians and values of the median

```{r}
for (m in mods_cv) {
  eff_sizes_cv_m <- eff_sizes_cv[[m]] %>%
  ggplot(aes(x=phone_dur_uttr,y=eff_size)) +
    geom_point() +
    xlab("Median average phone duration by utterance") +
    labs(y = "Δβ ") +
    theme_minimal() +
    theme(panel.grid = element_blank(),
        axis.line = element_line(size=0.8),
        axis.ticks = element_line(size=0.8),
        axis.text = element_text(size=16, colour="black"),
        axis.title = element_text(size=18, face="bold"),
        strip.background.y = element_blank(), 
        strip.placement = "outside",
        strip.text=element_text(size=18, face="bold"))
  ggsave(paste0("doreco_graphs_v1/cv/cv_eff_size_comp_", m, ".png"), width=5, height=5, dpi=300)
}
```

Comparing predicted phone durations and slopes at medians

```{r}
for (m in mods_cv) {
  eff_sizes_cv_m <- eff_sizes_cv[[m]] %>%
  ggplot() +
    geom_point(aes(x=phone_dur_v,y=derivative_v),col="orange") +
    geom_point(aes(x=phone_dur_c,y=derivative_c),col="purple") +
    geom_segment(aes(x=phone_dur_c,xend=phone_dur_v,
                     y=derivative_c,yend=derivative_v,
                     col = phone_dur_uttr)) +
    xlab("Predicted phone duration at median") +
    labs(y = "β") +
    theme_minimal() +
    theme(panel.grid = element_blank(),
        axis.line = element_line(size=0.8),
        axis.ticks = element_line(size=0.8),
        axis.text = element_text(size=16, colour="black"),
        axis.title = element_text(size=18, face="bold"),
        strip.background.y = element_blank(), 
        strip.placement = "outside",
        strip.text=element_text(size=18, face="bold"))
  ggsave(paste0("doreco_graphs_v1/cv/cv_deriv_pred_", m, ".png"), width=5, height=5, dpi=300)
}
```

# CV Model Density-Weighted Effect Sizes

## Computing CV Model Effect Curves and Density-Weighted Effect Sizes

```{r}
eff_curves_cv <- list()
eff_sizes_dw_cv <- list()

newdat_cv_deriv_curve <- expand.grid(
  phone_dur_uttr = seq(
    min(d_cv$phone_dur_uttr),
    max(d_cv$phone_dur_uttr),
    length.out = 100
  ),
  phone_type_f=d_cv$phone_type_f[1],
  #phone_type_new_f=d_cv$phone_type_new_f[1],
  speakerSegtype = d_cv$speakerSegtype[1],
  languageSegtype = unique(d_cv$languageSegtype),
  measure_type = "dur"
)




newdat_cv_deriv_curve$segtype_o <- str_extract(newdat_cv_deriv_curve$languageSegtype, "[^.]*$")
newdat_cv_deriv_curve$segtype_o <- as.ordered(newdat_cv_deriv_curve$segtype_o)
contrasts(newdat_cv_deriv_curve$segtype_o) <- "contr.treatment"

newdat_cv_deriv_curve$language <- str_extract(newdat_cv_deriv_curve$languageSegtype, "^[^.]*")
newdat_cv_deriv_curve$language_f <- as.factor(newdat_cv_deriv_curve$language)
newdat_cv_deriv_curve$speaker_f <- as.factor(str_extract(newdat_cv_deriv_curve$speakerSegtype, "^[^.]*"))
#newdat_cv_deriv_curve$speaker_f <- newdat_cv_deriv_curve$speaker_f[1]





# preds_baseline <- predict(cv_mod_baseline, newdat_cv_deriv_curve,
#   # the following is used when predicting without speakers
#   exclude=c("s(phone_dur_uttr,speakerSegtype)",
#             "s(phone_dur_uttr,speaker_f)",
#             "s(phone_dur_uttr,speaker_f):segtype_ovowel_dur_uttr",
#             "s(phone_dur_uttr,phone_type_f)"),
#   se.fit = TRUE
# )




language_dist_stats <- d_cv %>%
  select(language,phone_dur_uttr) %>%
  group_by(language) %>%
  nest() %>%
  mutate(
    phone_dur_uttr_med = as.numeric(map(data,~median(.x$phone_dur_uttr))),
    dens_data = map(data,~broom::tidy(density(.x$phone_dur_uttr,
                                              n = 100,
                                              from = min(d_cv$phone_dur_uttr),
                                              to = max(d_cv$phone_dur_uttr))))
         ) %>%
  unnest(dens_data) %>%
  mutate(phone_dur_uttr = x,
         density = y/max(y)) %>%
  ungroup() %>%
  select(language,phone_dur_uttr_med,phone_dur_uttr,density)
  

newdat_cv_deriv_curve <- left_join(
  mutate(newdat_cv_deriv_curve,
         phone_dur_uttr = round(phone_dur_uttr,15)),
  mutate(language_dist_stats,
         phone_dur_uttr = round(phone_dur_uttr,15)),
  by = c("language","phone_dur_uttr")
)


eps <- 0.001

newdat_cv_deriv_curve_1 <- newdat_cv_deriv_curve %>%
  mutate(phone_dur_uttr = exp(log(phone_dur_uttr) - (eps/2)))
newdat_cv_deriv_curve_2 <- newdat_cv_deriv_curve %>%
  mutate(phone_dur_uttr = exp(log(phone_dur_uttr) + (eps/2)))

mods_cv <- c("cv_mod_baseline", "cv_mod_type"
             #, "cv_mod_type2"
             )

for (m in mods_cv) {
X1_curve <- predict(get(m), 
              newdat_cv_deriv_curve_1,
              exclude=c(
                "s(phone_dur_uttr,speakerSegtype)",
                "s(phone_dur_uttr,speaker_f)",
                "s(phone_dur_uttr,speaker_f):segtype_ovowel_dur_uttr",
                "s(phone_dur_uttr,phone_type_f)"
                #"s(phone_dur_uttr,phone_type_new_f)"
              )
) %>% log()
X2_curve <- predict(get(m), 
              newdat_cv_deriv_curve_2,
              exclude=c(
                "s(phone_dur_uttr,speakerSegtype)",
                "s(phone_dur_uttr,speaker_f)",
                "s(phone_dur_uttr,speaker_f):segtype_ovowel_dur_uttr",
                "s(phone_dur_uttr,phone_type_f)"
                #"s(phone_dur_uttr,phone_type_new_f)"
              )
) %>% log()

X0_curve <- predict(get(m), 
              newdat_cv_deriv_curve,
              exclude=c(
                "s(phone_dur_uttr,speakerSegtype)",
                "s(phone_dur_uttr,speaker_f)",
                "s(phone_dur_uttr,speaker_f):segtype_ovowel_dur_uttr",
                "s(phone_dur_uttr,phone_type_f)"
                #"s(phone_dur_uttr,phone_type_new_f)"
              )
)

# finite difference approximation of first derivative
# the design matrix
first_deriv_curve <- (X2_curve - X1_curve) / eps
#first_deriv <- Xp %*% coef(get(m))



newdat_cv_deriv_curve$derivative <- first_deriv_curve
newdat_cv_deriv_curve$model <- m

newdat_cv_deriv_curve$phone_dur <- X0_curve



eff_curves_cv[[m]] <- newdat_cv_deriv_curve %>%
  dplyr::select(language, segtype_o, phone_dur_uttr_med, phone_dur_uttr, density, phone_dur, model, derivative) %>%
  group_by(language, phone_dur_uttr_med, phone_dur_uttr, density, model) %>%
  summarise(derivative_c = derivative[segtype_o == "consonant"],
            derivative_v = derivative[segtype_o == "syllabic"],
            delta_deriv = derivative_v - derivative_c,
            phone_dur_v = phone_dur[segtype_o == "syllabic"],
            phone_dur_c = phone_dur[segtype_o == "consonant"]) %>%
  group_by(language) %>%
  mutate(derivative_v_w = derivative_v*density,
         derivative_c_w = derivative_c*density,
         delta_deriv_w = delta_deriv*density,
         phone_dur_v_w = phone_dur_v*density,
         phone_dur_c_w = phone_dur_c*density) %>%
  ungroup()

eff_sizes_dw_cv[[m]] <- eff_curves_cv[[m]] %>%
  group_by(language, phone_dur_uttr_med, model) %>%
  summarize(density_area = sum(density),
            derivative_v_w_area = sum(derivative_v_w),
            derivative_c_w_area = sum(derivative_c_w),
            delta_deriv_w_area = sum(delta_deriv_w),
            phone_dur_v_w_area = sum(phone_dur_v_w),
            phone_dur_c_w_area = sum(phone_dur_c_w),
            eff_size_dw = delta_deriv_w_area/density_area,
            derivative_v_dw = derivative_v_w_area/density_area,
            derivative_c_dw = derivative_c_w_area/density_area,
            phone_dur_v_dw = phone_dur_v_w_area/density_area,
            phone_dur_c_dw = phone_dur_c_w_area/density_area,
            )

eff_sizes_dw_cv[[m]] <- left_join(
  eff_sizes_dw_cv[[m]],
  rename(eff_sizes_cv[[m]],eff_size_med = eff_size),
  by = "language")

eff_sizes_dw_cv[[m]] <- eff_sizes_dw_cv[[m]] %>%
  arrange(desc(eff_size_dw))
}

eff_curves_cv_df <- bind_rows(eff_curves_cv)
eff_curves_cv_df %>% write_csv("eff_curves_cv_df.csv")
eff_curves_cv_df <- read_csv("./eff_curves_cv_df.csv")

eff_sizes_dw_cv_df <- bind_rows(eff_sizes_dw_cv)
eff_sizes_dw_cv_df %>% write_csv("eff_sizes_dw_cv_df.csv")
eff_sizes_dw_cv_df <- read_csv("./eff_sizes_dw_cv_df.csv")
```


## Making CV Prediction Plots Showing Density-Weighted Effect Sizes
Ordering by effect size (delta-beta) & showing this number in the plot of CV predictions.

```{r}
for (m in mods_cv) {
  
  eff_sizes_dw_cv_m <- eff_sizes_dw_cv[[m]]
  eff_sizes_dw_cv_m$language <- factor(eff_sizes_dw_cv_m$language, levels=eff_sizes_dw_cv_m$language)
  eff_sizes_dw_cv_m$phone_dur_uttr_median <- eff_sizes_dw_cv_m$phone_dur_uttr
  eff_sizes_dw_cv_m$phone_dur_uttr <- 0.055
  eff_sizes_dw_cv_m$value <- 0.13
  
  newdat_m <- filter(newdat_cv_all, model==m)
  newdat_m$language <- factor(newdat_m$language, levels=eff_sizes_dw_cv_m$language)
  
  plot_dat <- d_uttr_cv
  plot_dat$language <- factor(plot_dat$language, levels=eff_sizes_dw_cv_m$language)
  
  plot_dat %>%
    ggplot(data = ., aes(x = phone_dur_uttr, y = value, col = measure)) +
    facet_wrap(. ~ language) +
    geom_point(alpha = 0.1, pch=16) + 
    #geom_ribbon(data = filter(newdat_cv_all,
    #                          model == "dur",
    #                          segtype_o == "vowel_dur_uttr"),
    #            aes(ymin = ll, ymax = ul), col = NA, fill = "grey",
    #            alpha = 0.5) +
    #geom_ribbon(data = filter(newdat_cv,
    #                          measure_type == "dur",
    #                          segtype_o == "cons_dur_uttr"),
    #            aes(ymin = ll, ymax = ul), col = NA, fill = "grey",
    #            alpha = 0.5) +
    geom_line(data = filter(newdat_m,
                            segtype_o == "syllabic"),
              aes(lty=model, y=dur),
              col = "darkorange1", lwd = 1) +
    geom_line(data = filter(newdat_m,
                            segtype_o == "consonant"),
              aes(lty=model, y=dur),
              col = "purple4", lwd = 1) +
    geom_text(data=eff_sizes_dw_cv_m, 
              aes(label=paste0("Δβ[dw] == ",
                               round(eff_sizes_dw_cv_m$eff_size_dw,2))),
              col="black", size=5, parse = TRUE) +
    geom_text(data=eff_sizes_dw_cv_m,
              aes(y=value/1.2,
                  label=paste0("Δβ[m] == ",
                               round(eff_sizes_dw_cv_m$eff_size_med,2))),
              col="black", size=5, parse = TRUE) +
    scale_x_log10(breaks = c(0.05, 0.07, 0.1, 0.14)) +
    scale_y_log10(breaks = c(0.05, 0.07, 0.1, 0.14),
                  limits = c(0.03, 0.19)) +
    scale_colour_manual(values = c("purple", "orange"), guide = "none") +
    labs(x = "Average segment duration in s (inverse speech rate)",
         y = "Average C vs. V duration in s") +
    theme_minimal() +
    theme(panel.grid = element_blank(),
        title = element_text(size=24, face="bold"),
        axis.line = element_line(size=0.8),
        axis.ticks = element_line(size=0.8),
        axis.text = element_text(size=16, colour="black"),
        axis.title = element_text(size=18, face="bold"),
        strip.background.y = element_blank(), 
        strip.placement = "outside",
        strip.text=element_text(size=18, face="bold"))
  ggsave(paste0("doreco_graphs_v1/cv/", m, ".png"), width=15, height=15, dpi=300)
}



```

## Plotting CV Model Effect Curves

Plot effect size curves

```{r}
for (m in mods_cv) {

  eff_curves_cv_m <- eff_curves_cv[[m]]
  eff_sizes_dw_cv_m <- eff_sizes_dw_cv[[m]]
  
  eff_sizes_dw_cv_m$language <- factor(eff_sizes_dw_cv_m$language,
                                     levels=eff_sizes_dw_cv_m$language)
  eff_curves_cv_m$language <- factor(eff_curves_cv_m$language,
                                     levels=eff_sizes_dw_cv_m$language)
  
  eff_curves_cv_m %>%
    ggplot(data = ., aes(x = phone_dur_uttr)) +
    facet_wrap(. ~ language) +
    #geom_ribbon(data = filter(newdat_cv_all,
    #                          model == "dur",
    #                          segtype_o == "vowel_dur_uttr"),
    #            aes(ymin = ll, ymax = ul), col = NA, fill = "grey",
    #            alpha = 0.5) +
    #geom_ribbon(data = filter(newdat_cv,
    #                          measure_type == "dur",
    #                          segtype_o == "cons_dur_uttr"),
    #            aes(ymin = ll, ymax = ul), col = NA, fill = "grey",
    #            alpha = 0.5) +
    # geom_density(data = d_cv,
    #              aes(x = phone_dur_uttr, after_stat(scaled)),
    #              n = 100, trim = FALSE, #bounds = c(min(d_cv$phone_dur_uttr),max(d_cv$phone_dur_uttr)),
    #              fill = "forestgreen",lwd=0, alpha=0.3) +
    geom_area(aes(y = density),
              fill = "darkgreen",alpha=0.3) +
    # geom_line(aes(y = derivative_v),
    #           col = "orange", lwd = 1) +
    # geom_line(aes(y = derivative_c),
    #           col = "purple", lwd = 1) +
    geom_line(aes(y = delta_deriv),
              col = "darkblue", lwd = 1) +
    # geom_line(aes(y = delta_deriv_w),
    #           col = "firebrick", lwd = 1) +
    geom_text(data = eff_sizes_dw_cv_m,
              aes(label=paste0("Δβ[w] == ", round(eff_size_dw,2)),
                  x = 0.05, y = 0.9),
              col="darkblue", size=5, parse=TRUE) +
    geom_text(data = eff_sizes_dw_cv_m,
              aes(label=paste0("Δβ[m] == ", round(eff_size_med,2)),
                  x = 0.05, y = 0.75),
              col="red", size=5, parse=TRUE) +
    geom_hline(data = eff_sizes_dw_cv_m,
               aes(yintercept = eff_size_dw),
               linetype = 'dashed',
               col = "darkblue") +
    geom_hline(data = eff_sizes_dw_cv_m,
               aes(yintercept = eff_size_med),
               linetype = 'dashed',
               col = "red") +
    geom_hline(yintercept = 0, linetype = 'dashed') +
    geom_vline(aes(xintercept = phone_dur_uttr_med),
               col = 'red') +
    
    scale_x_log10(breaks = c(0.05, 0.07, 0.1, 0.14)) +
    scale_y_continuous(
      limits=c(-0.2,1.1)
      ) +
    labs(title = "Effect Curves",
         x = "Average segment duration in s (inverse speech rate)",
         y = "Delta derivative") +
    theme_minimal() +
    theme(panel.grid = element_blank(),
        title = element_text(size=24, face="bold"),
        axis.line = element_line(size=0.8),
        axis.ticks = element_line(size=0.8),
        axis.text = element_text(size=16, colour="black"),
        axis.title = element_text(size=18, face="bold"),
        strip.background.y = element_blank(), 
        strip.placement = "outside",
        strip.text=element_text(size=18, face="bold"))
  ggsave(paste0("doreco_graphs_v1/cv/effect_curves_", m, ".png"), width=15, height=15, dpi=300)
}
```

Derivative curves

```{r}
for (m in mods_cv) {

  eff_curves_cv_m <- eff_curves_cv[[m]]
  eff_sizes_dw_cv_m <- eff_sizes_dw_cv[[m]]
  
  eff_sizes_dw_cv_m$language <- factor(eff_sizes_dw_cv_m$language,
                                     levels=eff_sizes_dw_cv_m$language)
  eff_curves_cv_m$language <- factor(eff_curves_cv_m$language,
                                     levels=eff_sizes_dw_cv_m$language)
  
  eff_curves_cv_m %>%
    ggplot(data = ., aes(x = phone_dur_uttr)) +
    facet_wrap(. ~ language) +
    #geom_ribbon(data = filter(newdat_cv_all,
    #                          model == "dur",
    #                          segtype_o == "vowel_dur_uttr"),
    #            aes(ymin = ll, ymax = ul), col = NA, fill = "grey",
    #            alpha = 0.5) +
    #geom_ribbon(data = filter(newdat_cv,
    #                          measure_type == "dur",
    #                          segtype_o == "cons_dur_uttr"),
    #            aes(ymin = ll, ymax = ul), col = NA, fill = "grey",
    #            alpha = 0.5) +
    # geom_density(data = d_cv,
    #              aes(x = phone_dur_uttr, after_stat(scaled)),
    #              n = 100, trim = FALSE, #bounds = c(min(d_cv$phone_dur_uttr),max(d_cv$phone_dur_uttr)),
    #              fill = "forestgreen",lwd=0, alpha=0.3) +
    # geom_line(aes(y = density),
    #           col = "darkgreen", lwd = 0.5) +
    geom_area(aes(y = density/2+1),
              fill = "darkgreen",alpha=0.3) +
    geom_line(aes(y = derivative_v),
              col = "orange", lwd = 1) +
    geom_line(aes(y = derivative_c),
              col = "purple", lwd = 1) +
    geom_hline(yintercept = 1, linetype = 'dashed') +
    # geom_line(aes(y = delta_deriv),
    #           col = "darkblue", lwd = 1) +
    geom_vline(aes(xintercept = phone_dur_uttr_med),
               col = 'red') +
    scale_x_log10(breaks = c(0.05, 0.07, 0.1, 0.14)) +
    scale_y_log10(n.breaks = 8) +
    labs(title = "Derivative Curves",
         x = "Average segment duration in s (inverse speech rate)",
         y = "Predicted phone duration") +
    theme_minimal() +
    theme(panel.grid = element_blank(),
        title = element_text(size=24, face="bold"),
        axis.line = element_line(size=0.8),
        axis.ticks = element_line(size=0.8),
        axis.text = element_text(size=16, colour="black"),
        axis.title = element_text(size=18, face="bold"),
        strip.background.y = element_blank(), 
        strip.placement = "outside",
        strip.text=element_text(size=18, face="bold"))
  ggsave(paste0("doreco_graphs_v1/cv/derivative_curves_", m, ".png"), width=15, height=15, dpi=300)
}
```

## Comparing CV Model Density-Weighted Effect Sizes

Comparing effect sizes computed in different ways

```{r}
for (m in mods_cv) {
  eff_sizes_dw_cv_m <- eff_sizes_dw_cv[[m]] %>%
  ggplot(aes(x=eff_size_dw, y=eff_size_med)) +
    geom_segment(x=-1,xend=1,
                 y=-1,yend=1,
                 col="red", lwd = 0.25) +
    geom_point() +
    scale_x_continuous(limits = c(0,0.65)) +
    scale_y_continuous(limits = c(0,0.65)) +
    labs(x = "Δβ duration-weighted",
         y = "Δβ at median") +
    theme_minimal() +
    theme(panel.grid = element_blank(),
        axis.line = element_line(size=0.8),
        axis.ticks = element_line(size=0.8),
        axis.text = element_text(size=16, colour="black"),
        axis.title = element_text(size=18, face="bold"),
        strip.background.y = element_blank(), 
        strip.placement = "outside",
        strip.text=element_text(size=18, face="bold"))
  ggsave(paste0("doreco_graphs_v1/cv/cv_eff_type_comp_", m, ".png"), width=5, height=5, dpi=300)
  
  eff_sizes_dw_cv_m <- eff_sizes_dw_cv[[m]] %>%
  ggplot(aes(x=phone_dur_uttr_med, y = eff_size_med - eff_size_dw)) +
    geom_hline(yintercept=0, col="red", lwd = 0.25) +
    geom_point() +
    xlab("Median") +
    labs(y = "Δβ duration-weighted - Δβ at median") +
    theme_minimal() +
    theme(panel.grid = element_blank(),
        axis.line = element_line(size=0.8),
        axis.ticks = element_line(size=0.8),
        axis.text = element_text(size=16, colour="black"),
        axis.title = element_text(size=18, face="bold"),
        strip.background.y = element_blank(), 
        strip.placement = "outside",
        strip.text=element_text(size=18, face="bold"))
  ggsave(paste0("doreco_graphs_v1/cv/cv_eff_type_diff_", m, ".png"), width=5, height=5, dpi=300)
}
```
Comparing predicted phone durations and slopes
```{r}
for (m in mods_cv) {
  eff_sizes_dw_cv_m <- eff_sizes_dw_cv[[m]] %>%
  ggplot() +
    geom_point(aes(x=phone_dur_v_dw,y=derivative_v_dw),col="orange") +
    geom_point(aes(x=phone_dur_c_dw,y=derivative_c_dw),col="purple") +
    geom_segment(aes(x=phone_dur_c_dw,xend=phone_dur_v_dw,
                     y=derivative_c_dw,yend=derivative_v_dw)) +
    xlab("Predicted density-weighted phone duration") +
    labs(y = "β ") +
    theme_minimal() +
    theme(panel.grid = element_blank(),
        axis.line = element_line(size=0.8),
        axis.ticks = element_line(size=0.8),
        axis.text = element_text(size=16, colour="black"),
        axis.title = element_text(size=18, face="bold"),
        strip.background.y = element_blank(), 
        strip.placement = "outside",
        strip.text=element_text(size=18, face="bold"))
  ggsave(paste0("doreco_graphs_v1/cv/cv_deriv_pred_dw_", m, ".png"), width=5, height=5, dpi=300)
}

```



