---
title: "doreco_stretchiness_analysis"
author: "Noah Luntzlara"
date: "2025-10-01"
output: pdf_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Set up

Load packages.

```{r}
library(tidyverse)
library(forcats)

library(ggbeeswarm)
library(halfmoon)
# library(geomtextpath)
# library(ggpubr)
# 
# library(mgcv)
# library(itsadug)
# 
# library(RColorBrewer)
# library(viridis)
# 
# library(broom)
# library(dplyr)
# library(tidyr)
```

Read in the data.

```{r message=FALSE}

apah <- read_csv("../doreco_data/doreco_apah1238_durations.csv",
                 col_types = cols(.default = "?", speaker = "c", sex = "c", all_speakers = "c")) %>%
  mutate(language="Yali") %>%
  arrange(speaker, discourse, phone_start)
arap <- read_csv("../doreco_data/doreco_arap1274_durations.csv",
                 col_types = cols(.default = "?", speaker = "c", sex = "c", all_speakers = "c")) %>%
  mutate(language="Arapaho") %>%
  arrange(speaker, discourse, phone_start)
bain <- read_csv("../doreco_data/doreco_bain1259_durations.csv",
                 col_types = cols(.default = "?", speaker = "c", sex = "c", all_speakers = "c")) %>%
  mutate(language="Gubeeher") %>%
  arrange(speaker, discourse, phone_start)
beja <- read_csv("../doreco_data/doreco_beja1238_durations.csv",
                 col_types = cols(.default = "?", speaker = "c", sex = "c", all_speakers = "c")) %>%
  mutate(language="Beja") %>%
  arrange(speaker, discourse, phone_start)
bora <- read_csv("../doreco_data/doreco_bora1263_durations.csv",
                 col_types = cols(.default = "?", speaker = "c", sex = "c", all_speakers = "c")) %>%
  mutate(language="Bora") %>%
  arrange(speaker, discourse, phone_start)
dolg <- read_csv("../doreco_data/doreco_dolg1241_durations.csv",
                 col_types = cols(.default = "?", speaker = "c", sex = "c", all_speakers = "c")) %>%
  mutate(language="Dolgan") %>%
  arrange(speaker, discourse, phone_start)
even <- read_csv("../doreco_data/doreco_even1259_durations.csv",
                 col_types = cols(.default = "?", speaker = "c", sex = "c", all_speakers = "c")) %>%
  mutate(language="Evenki") %>%
  arrange(speaker, discourse, phone_start)
guri <- read_csv("../doreco_data/doreco_guri1247_durations.csv",
                 col_types = cols(.default = "?", speaker = "c", sex = "c", all_speakers = "c")) %>%
  mutate(language="Gurindji") %>%
  arrange(speaker, discourse, phone_start)
jeju <- read_csv("../doreco_data/doreco_jeju1234_durations.csv",
                 col_types = cols(.default = "?", speaker = "c", sex = "c", all_speakers = "c")) %>%
  mutate(language="Jejuan") %>%
  arrange(speaker, discourse, phone_start)
kama <- read_csv("../doreco_data/doreco_kama1351_durations.csv",
                 col_types = cols(.default = "?", speaker = "c", sex = "c", all_speakers = "c")) %>%
  mutate(language="Kamas") %>%
  arrange(speaker, discourse, phone_start)
lowe <- read_csv("../doreco_data/doreco_lowe1385_durations.csv",
                 col_types = cols(.default = "?", speaker = "c", sex = "c", all_speakers = "c")) %>%
  mutate(language="Lower_Sorbian") %>%
  arrange(speaker, discourse, phone_start)
movi <- read_csv("../doreco_data/doreco_movi1243_durations.csv",
                 col_types = cols(.default = "?", speaker = "c", sex = "c", all_speakers = "c")) %>%
  mutate(language="Movima") %>%
  arrange(speaker, discourse, phone_start)
ngal <- read_csv("../doreco_data/doreco_ngal1292_durations.csv",
                 col_types = cols(.default = "?", speaker = "c", sex = "c", all_speakers = "c")) %>%
  mutate(language="Dalabon") %>%
  arrange(speaker, discourse, phone_start)
nngg <- read_csv("../doreco_data/doreco_nngg1234_durations.csv",
                 col_types = cols(.default = "?", speaker = "c", sex = "c", all_speakers = "c")) %>%
  mutate(language="Nuu") %>%
  arrange(speaker, discourse, phone_start)
orko <- read_csv("../doreco_data/doreco_orko1234_durations.csv",
                 col_types = cols(.default = "?", speaker = "c", sex = "c", all_speakers = "c")) %>%
  mutate(language="Fanbyak") %>%
  arrange(speaker, discourse, phone_start)
resi <- read_csv("../doreco_data/doreco_resi1247_durations.csv",
                 col_types = cols(.default = "?", speaker = "c", sex = "c", all_speakers = "c")) %>%
  mutate(language="Resigaro") %>%
  arrange(speaker, discourse, phone_start)
ruul <- read_csv("../doreco_data/doreco_ruul1235_durations.csv",
                 col_types = cols(.default = "?", speaker = "c", sex = "c", all_speakers = "c")) %>%
  mutate(language="Ruuli") %>%
  arrange(speaker, discourse, phone_start)
sadu <- read_csv("../doreco_data/doreco_sadu1234_durations.csv",
                 col_types = cols(.default = "?", speaker = "c", sex = "c", all_speakers = "c")) %>%
  mutate(language="Sadu") %>%
  arrange(speaker, discourse, phone_start)
sanz <- read_csv("../doreco_data/doreco_sanz1248_durations.csv",
                col_types = cols(.default = "?", speaker = "c", sex = "c", all_speakers = "c")) %>%
 mutate(language="Sanzhi_Dargwa") %>%
 arrange(speaker, discourse, phone_start)
sout <- read_csv("../doreco_data/doreco_sout3282_durations.csv",
                 col_types = cols(.default = "?", speaker = "c", sex = "c", all_speakers = "c")) %>%
  mutate(language="English") %>%
  arrange(speaker, discourse, phone_start)
svan <- read_csv("../doreco_data/doreco_svan1243_durations.csv",
                 col_types = cols(.default = "?", speaker = "c", sex = "c", all_speakers = "c")) %>%
  mutate(language="Svan") %>%
  arrange(speaker, discourse, phone_start)
trin <- read_csv("../doreco_data/doreco_trin1278_durations.csv",
                col_types = cols(.default = "?", speaker = "c", sex = "c", all_speakers = "c")) %>%
 mutate(language="Mojeno_Trinitario") %>%
 arrange(speaker, discourse, phone_start)
urum <- read_csv("../doreco_data/doreco_urum1249_durations.csv",
                 col_types = cols(.default = "?", speaker = "c", sex = "c", all_speakers = "c")) %>%
  mutate(language="Urum") %>%
  arrange(speaker, discourse, phone_start)
vera <- read_csv("../doreco_data/doreco_vera1241_durations.csv",
                 col_types = cols(.default = "?", speaker = "c", sex = "c", all_speakers = "c")) %>%
  mutate(language="Veraa") %>%
  arrange(speaker, discourse, phone_start)
yong <- read_csv("../doreco_data/doreco_yong1270_durations.csv",
                 col_types = cols(.default = "?", speaker = "c", sex = "c", all_speakers = "c")) %>%
  mutate(language="Yongning_Na") %>%
  arrange(speaker, discourse, phone_start)
```
Combine into one data frame.

```{r}
d <- bind_rows(apah,
               arap,
               bain,
               beja,
               bora,
               dolg,
               even,
               guri,
               jeju,
               kama,
               lowe,
               movi,
               ngal,
               nngg,
               orko,
               resi,
               ruul,
               sadu,
               sanz,
               sout,
               svan,
               trin,
               urum,
               vera,
               yong
               )
```

Transformation of discourse metadata to weights for regression (not used).

```{r}
bad_wt <- 0.5
medium_wt <- 0.75
good_wt <- 1

d <- d %>%
  mutate(sound_quality_wt = case_when(
                              sound_quality == "bad" ~ bad_wt,
                              sound_quality == "medium" ~ medium_wt,
                              sound_quality == "good" ~ good_wt
                              ),
         background_noise_wt = case_when(
                              background_noise == "constant" ~ bad_wt,
                              background_noise == "punctual" ~ medium_wt,
                              background_noise == "none_from_doreco_metadata" ~ good_wt
                              ),
         genre_wt = case_when(
                              genre == "procedural" ~ bad_wt,
                              genre == "stimulus retelling" ~ bad_wt,
                              genre == "traditional narrative" ~ medium_wt,
                              genre == "personal narrative" ~ good_wt,
                              genre == "conversation" ~ good_wt
                              )
         )
```

Add the typographical information of vowel length contrast.

```{r}
d <- d %>%
  mutate(syll_pos = if_else(syll_end == word_end,"final","nonfinal")) %>%
  mutate(vowel_length_contrast = case_when(language %in% c("Dalabon",
                                                    "Gurindji",
                                                    "Jejuan",
                                                    "Lower_Sorbian",
                                                    "Nuu",
                                                    "Sadu",
                                                    "Sanzhi_Dargwa",
                                                    "English",
                                                    "Urum",
                                                    "Veraa",
                                                    "Yali",
                                                    "Yongning_Na") ~ "none",
                                    language %in% c("Kamas",
                                                    "Mojeno_Trinitario",
                                                    "Ruuli") ~ "restricted",
                                    language %in% c("Arapaho",
                                                    "Beja",
                                                    "Bora",
                                                    "Dolgan",
                                                    "Evenki",
                                                    "Fanbyak",
                                                    "Gubeeher",
                                                    "Movima",
                                                    "Resigaro",
                                                    "Svan") ~ "unrestricted")
         ) %>%
  mutate(segtype_vowel_length_revised = if_else(
    (segtype_syllabic == "vowel") &
      (vowel_length_contrast == "none" |
         (vowel_length_contrast == "restricted" & syll_pos == "final")) ,
                                        "short",
                                        segtype_vowel_length),     
        vlcLn = ifelse((vowel_length_contrast == "none") |
                (vowel_length_contrast == "restricted" & syll_pos == "final"),
                "no_contrast",
                paste("contrast",segtype_vowel_length_revised,sep="_"))
         )

```


Unite phone type into a single variable

```{r}
d <- d %>%
  mutate(phone_type = case_when(is.na(segtype) ~ 'nonspeech',
                                segtype_syllabic == 'vowel'
                                ~ paste("vowel",
                                        # if the vowel is a monophthong
                                        case_when(segtype_vowel_dipthongality == "monophthong"
                                        # label it "_mono_<long/short>_<high/nonhigh>"
                                        ~ paste("mono",
                                                segtype_vowel_length_revised,
                                                segtype_vowel_height,
                                                sep="_"),
                                        # if the vowel is a diphthong
                                        segtype_vowel_dipthongality == "diphthong"
                                        # just label it "_diph"
                                        ~ "diph"),
                                        sep="_"),
                                segtype_syllabic == 'syll_c'
                                ~ 'syllabic_consonant',
                                segtype_consonant == 'sonorant' #& segtype_complex != 'double_stop'
                                ~ segtype_sonorant,
                                segtype_consonant == 'obstruent' #&
                                  #segtype_obstruent != 'click' &
                                  #segtype_complex != 'double_stop'
                                ~ segtype_obstruent,
                                #segtype_obstruent == 'click' & segtype_click == 'simple'
                                #~ segtype_obstruent,
                                #segtype_complex == 'double_stop' & segtype_consonant == 'obstruent'
                                #~ paste(segtype_obstruent,'double',sep="_"),
                                #segtype_complex == 'double_stop' & segtype_consonant == 'sonorant'
                                #~ paste(segtype_sonorant,'double',sep="_"),
                                #segtype_obstruent == 'click' & segtype_click != 'simple'
                                #~ 'click_contour'
                                )
         # phone_type_new = case_when(
         #                        # if the phone is a vowel
         #                        segtype_syllabic == 'vowel'
         #                        ~ paste("vowel",
         #                                # if the vowel is a monophthong
         #                                case_when(segtype_vowel_dipthongality == "monophthong"
         #                                # label it "_mono_<long/short>_<high/nonhigh>"
         #                                ~ paste("mono",
         #                                        segtype_vowel_length_revised,
         #                                        segtype_vowel_height,
         #                                        sep="_"),
         #                                # if the vowel is a diphthong
         #                                segtype_vowel_dipthongality == "diphthong"
         #                                # just label it "_diph"
         #                                ~ "diph"),
         #                                sep="_"),
         #                        .default = phone_type
         #                        )
  )
```



Calculate utterance level measures and compare with those generated by PGDB.
```{r}
# for analysis - segment level
d_phones <- d %>%
  filter(phone_type != 'nonspeech'
         ) %>%
  group_by(
    language,
    speaker,
    uttr_id
  ) %>%
  # rename old measures
  rename(uttr_dur_pgdb = uttr_dur,
         word_dur_pgdb = word_dur,
         syll_dur_pgdb = syll_dur,
         num_phones_uttr_pgdb = num_phones_uttr,
         num_sylls_uttr_pgdb = num_sylls_uttr,
         num_words_uttr_pgdb = num_words_uttr,
         num_phones_word_pgdb = num_phones_word,
         num_sylls_word_pgdb = num_sylls_word,
         num_phones_syll_pgdb = num_phones_syll,
         ) %>%
  # Encode utterance level measures before filtering
  mutate(
    uttr_dur_init = sum(phone_dur),
    num_phones_uttr_init = n(),
    num_sylls_uttr_init = n_distinct(syll_id),
    num_words_uttr_init = n_distinct(word_id),
    num_cons_uttr_init = sum(segtype == "consonant"),
    num_vowel_uttr_init = sum(segtype == "syllabic"),
    cv_ratio_uttr_init = num_cons_uttr_init/num_vowel_uttr_init
    ) %>%
  ungroup() %>%
  # Encode word level measures before filtering
  group_by(language,speaker,uttr_id,word_id) %>%
  mutate(word_dur_init = sum(phone_dur),
         num_phones_word_init = n(),
         num_sylls_word_init = n_distinct(syll_id)) %>%
  ungroup() %>%
  # Encode syllable level measures before filtering
  group_by(language,speaker,uttr_id,word_id,syll_id) %>%
  mutate(syll_dur_init = sum(phone_dur),
         num_phones_syll_init = n()) %>%
  ungroup() %>%
  # Encode rate measures before filtering
  mutate(
    phone_dur_uttr_init = uttr_dur_init / num_phones_uttr_init,
    phone_dur_uttr_ex_init = (uttr_dur_init - phone_dur) / (num_phones_uttr_init - 1),
    syll_dur_uttr_init = uttr_dur_init / num_sylls_uttr_init,
    syll_dur_uttr_ex_init = if_else(num_sylls_uttr_init > 1, 
                                    (uttr_dur_init - syll_dur_init) / (num_sylls_uttr_init - 1),
                                    syll_dur_uttr_init),
   word_dur_uttr_init = uttr_dur_init / num_words_uttr_init,
   word_dur_uttr_ex_init = if_else(num_words_uttr_init > 1, 
                                    (uttr_dur_init - word_dur_init) / (num_words_uttr_init - 1),
                                    word_dur_uttr_init)
   )

unique(d_phones$phone_type)
d_phones %>%
  filter(phone %in% c("h", "?")) %>%
  select(language, phone_type, phone, phone_ipa)

#Checks that the utterance-level metrics compare as expected with the ones generated by PolyglotDB.

d_phones_check <- d_phones %>%
  mutate(num_phones_uttr_check = (num_phones_uttr_init == num_phones_uttr_pgdb),
         num_sylls_uttr_check = (num_sylls_uttr_init == num_sylls_uttr_pgdb),
         num_words_uttr_check = (num_words_uttr_init == num_words_uttr_pgdb),
         uttr_dur_check = abs(uttr_dur_pgdb - uttr_dur_init) < 0.0001) %>%
  group_by(num_phones_uttr_check,
           num_sylls_uttr_check,
           num_words_uttr_check,
           uttr_dur_check) %>%
  summarize(num_uttr = n_distinct(uttr_id))
d_phones_check

d_phones_uttr_dur_check <- d_phones

d_phones_uttr_dur_check %>%
  group_by(language,speaker,uttr_id) %>%
  summarize(uttr_dur_pgdb = uttr_dur_pgdb[1],
            uttr_dur_init = uttr_dur_init[1]) %>%
  mutate(uttr_dur_diff = uttr_dur_pgdb - uttr_dur_init) %>%
  filter(uttr_dur_diff > 0.0001) %>%
  ggplot(aes(x = uttr_dur_diff)) +
  facet_wrap(~language) +
  geom_density(bw = 0.04) +
  scale_x_log10(labels = scales::label_number(drop0trailing=TRUE)) +
  labs(x = "discrepancy in pgdb utterance duration")
```

# Filtering

Setting thresholds.
```{r}
threshold_sylls_min <- 5

threshold_uttr_min <- 0.04
threshold_uttr_max <- 0.2

threshold_individual_min <- 0.03 # changed from 0.025
threshold_individual_max <- 0.5 # changed from 0.4

threshold_z_uttr_dur <- 3
threshold_z_phone_dur <- 3

threshold_num_phones_uttr <- 10
threshold_num_phones_removed <- 5
threshold_prop_phones_removed <- 0.1
```

Performing filtering
```{r}
# First round of filtering

# Perform utterance-level filtering

# Create a data frame summarized by utterance
d_uttr <- d_phones %>%
  group_by(language,speaker,uttr_id) %>%
  summarize(num_sylls_uttr_init = num_sylls_uttr_init[1],
            phone_dur_uttr_init = phone_dur_uttr_init[1])

# Remove utterances with 5 syllables or fewer
d_uttr_filt_1.1 <- d_uttr %>%
  filter(num_sylls_uttr_init > threshold_sylls_min)

# Remove utterances with average phone duration outside (40ms, 200ms)
d_uttr_filt_1.2 <- d_uttr_filt_1.1 %>%
  filter(phone_dur_uttr_init > threshold_uttr_min,
         phone_dur_uttr_init < threshold_uttr_max) %>%
  # create thresholds for log average phone duration by utterance
  # 3 standard deviations above/below the mean by language
  group_by(language) %>%
  mutate(
    threshold_uttr_by_lang_min =
    exp(
      mean(log(phone_dur_uttr_init)) -
      threshold_z_uttr_dur*sd(log(phone_dur_uttr_init))
      ),
    threshold_uttr_by_lang_max =
    exp(
      mean(log(phone_dur_uttr_init)) +
      threshold_z_uttr_dur*sd(log(phone_dur_uttr_init))
    )
  ) %>%
  ungroup()
  
d_uttr_filt_1.3 <- d_uttr_filt_1.2 %>%
  # remove utterances with log average phone duration by utterance
  # # 3 standard deviations above/below the mean by language
  filter(phone_dur_uttr_init > threshold_uttr_by_lang_min,
         phone_dur_uttr_init < threshold_uttr_by_lang_max)

d_uttr_filt_1 <- d_uttr_filt_1.3

d_phones_filt_1.1 <- d_phones %>%
  # remove utterances removed by utterance-level filtering
  filter(uttr_id %in% d_uttr_filt_1$uttr_id)

d_phones_filt_1.2 <- d_phones_filt_1.1 %>%
  # remove phones with duration outside (20ms, 500ms)
  filter(
    phone_dur > threshold_individual_min,
    phone_dur < threshold_individual_max
         )

d_phones_filt_1 <- d_phones_filt_1.2

# Counting phones removed
d_phones_filt_1 <- d_phones_filt_1 %>%
  group_by(language,speaker,uttr_id) %>%
  mutate(num_phones_uttr = n(),
         num_phones_uttr_removed = num_phones_uttr_init - num_phones_uttr,
         prop_phones_uttr_removed = num_phones_uttr_removed/num_phones_uttr_init) %>%
  ungroup()

# Recalculating rate measures
d_phones_filt_1 <- d_phones_filt_1 %>%
  # Encode utterance level measures after filtering
  group_by(language,speaker,uttr_id) %>%
  mutate(
    uttr_dur = sum(phone_dur),
    num_phones_uttr = n(),
    num_sylls_uttr = n_distinct(syll_id),
    num_words_uttr = n_distinct(word_id),
    num_cons_uttr = sum(segtype == "consonant"),
    num_vowel_uttr = sum(segtype == "syllabic"),
    cv_ratio_uttr = num_cons_uttr/num_vowel_uttr
  ) %>%
  ungroup() %>%
  # Encode word level measures after filtering
  group_by(language,speaker,uttr_id,word_id) %>%
  mutate(word_dur = sum(phone_dur),
         num_phones_word = n(),
         num_sylls_word = n_distinct(syll_id)) %>%
  ungroup() %>%
  # Encode syllable level measures after filtering
  group_by(language,speaker,uttr_id,word_id,syll_id) %>%
  mutate(syll_dur = sum(phone_dur),
         num_phones_syll = n()) %>%
  ungroup() %>%
  # Encode rate measures after filtering
  mutate(
    phone_dur_uttr = uttr_dur / num_phones_uttr,
    phone_dur_uttr_ex = (uttr_dur - phone_dur) / (num_phones_uttr - 1),
    syll_dur_uttr = uttr_dur / num_sylls_uttr,
    syll_dur_uttr_ex = if_else(num_sylls_uttr > 1, 
                               (uttr_dur - syll_dur) / (num_sylls_uttr - 1),
                               syll_dur_uttr),
    word_dur_uttr = uttr_dur / num_words_uttr,
    word_dur_uttr_ex = if_else(num_words_uttr > 1, 
                               (uttr_dur - word_dur) / (num_words_uttr - 1),
                               word_dur_uttr)
    )

# Second round of filtering

d_phones_filt_2 <- d_phones_filt_1 %>%
  # remove utterances with at least 10% of phones removed OR
  # at least 5 phones removed
  filter(num_phones_uttr_removed < threshold_num_phones_removed,
         prop_phones_uttr_removed < threshold_prop_phones_removed)

# d_phones_filt_2 <- d_phones_filt_1 %>%
#   # remove utterances with 10 phones or fewer
#   filter(num_phones_uttr >= threshold_num_phones_uttr)

# DoReCo
#  [1] "vowel_mono_short_nonhigh" "plosive" -> stop                  "fricative"                "nasal"
#  [5] "trill"                    "vowel_mono_short_high"    "approximant"              "liquid"
#  [9] "affricate"                "glottal"                  "vowel_mono_long_nonhigh"  "vowel_diph"
# [13] "vowel_mono_long_high"     "flap"                     "click"                    "syllabic_consonant"

# Oops
#  [1] "stop"           "non-high vowel" "nasal"          "fricative"      "high vowel"     "affricate"      "approximant"
#  [8] "liquid"         "NA"             "syllabic C"

# Additional categories: click, monophthong (high vs non-high), diphthong (a single category), (trill, tap/flap) -> liquid

# Third round of filtering

# Create a data frame summarized by utterance
d_uttr_filt_2.1 <- d_phones_filt_2 %>%
  group_by(language,speaker,uttr_id) %>%
  summarize(phone_dur_uttr = phone_dur_uttr[1])

d_uttr_thresholds_by_lang_1 <- d_uttr_filt_1 %>%
  group_by(language) %>%
  summarize(threshold_uttr_by_lang_min=threshold_uttr_by_lang_min[1],
            threshold_uttr_by_lang_max=threshold_uttr_by_lang_max[1])

d_uttr_filt_2.1 <- d_uttr_filt_2.1 %>%
  left_join(d_uttr_thresholds_by_lang_1,by="language")
  # group_by(language) %>%
  # mutate(
  #   threshold_uttr_by_lang_2_min =
  #   exp(
  #     mean(log(phone_dur_uttr)) -
  #     threshold_z_uttr_dur*sd(log(phone_dur_uttr))
  #     ),
  #   threshold_uttr_by_lang_2_max =
  #   exp(
  #     mean(log(phone_dur_uttr)) +
  #     threshold_z_uttr_dur*sd(log(phone_dur_uttr))
  #   )
  # ) %>%
  # ungroup()

d_uttr_filt_2 <- d_uttr_filt_2.1

d_uttr_filt_3 <- d_uttr_filt_2 %>%
  filter(phone_dur_uttr > threshold_uttr_by_lang_min,
         phone_dur_uttr < threshold_uttr_by_lang_max)
  # filter(phone_dur_uttr > threshold_uttr_by_lang_2_min,
  #        phone_dur_uttr < threshold_uttr_by_lang_2_max)

d_phones_filt_3 <- d_phones_filt_2 %>%
  filter(uttr_id %in% d_uttr_filt_3$uttr_id)
```

```{r}

# Final filtered dataframe
d_phones_filt <- d_phones_filt_1 # d_phones_filt_3

# Renaming measures in unfiltered dataframe
d_phones_unfilt <- d_phones %>%
  rename_with(~str_remove(., '_init'))

d_phones_filt %>% write_csv("../doreco_data/combined_doreco_durations_filtered.csv")

d_phones_unfilt %>% write_csv("../doreco_data/combined_doreco_durations_unfiltered.csv")
```

# Data exploration

## Metadata

Make dataframes to look at metadata by discourse (i.e. recording), language, and speaker.
```{r}
d <- d %>%
  group_by(discourse) %>%
  mutate(discourse_dur = max(phone_end)) %>%
  ungroup()

d_discourses <- d %>%
  group_by(discourse) %>%
  summarize(num_speakers = n_distinct(speaker),
            language = language[1],
            genre = genre[1],
            sound_quality = sound_quality[1],
            background_noise = background_noise[1],
            all_speakers = all_speakers[1],
            discourse_dur = discourse_dur[1])

d_languages <- d %>%
  group_by(language) %>%
  summarize(num_speakers = n_distinct(speaker),
            hours = sum(discourse_dur)/3600)

d_discourses_speakers <- d_discourses %>%
  group_by(language,all_speakers) %>%
  summarize(num_speakers = num_speakers[1],
            hours = sum(discourse_dur)/3600)

d_discourses_genres <- d_discourses %>%
  mutate(genre = factor(genre,
                        levels=c("procedural",
                                 "stimulus retelling",
                                 "conversation",
                                 "traditional narrative",
                                 "personal narrative"))) %>%
  group_by(genre,language) %>%
  summarize(hours = sum(discourse_dur)/3600,
            num_discourses = n())
d_discourses_genres_summary <- d_discourses_genres %>%
  group_by(genre) %>%
  summarize(hours = sum(hours),
            num_discourses=sum(num_discourses),
            num_languages=n_distinct(language))

d_discourses_quality <- d_discourses %>%
  mutate(sound_quality = factor(sound_quality,
                                levels=c("bad",
                                         "medium",
                                         "good"))) %>%
  group_by(language,sound_quality) %>%
  summarize(hours = sum(discourse_dur)/3600,
            num_discourses = n())
d_discourses_quality_summary <- d_discourses_quality %>%
  group_by(sound_quality) %>%
  summarize(hours = sum(hours),
            num_discourses=sum(num_discourses),
            num_languages=n_distinct(language))

d_discourses_noise <- d_discourses %>%
  mutate(background_noise = factor(background_noise,
                        levels=c("constant",
                                 "punctual",
                                 "none_from_doreco_metadata"),
                        labels=c("constant",
                                 "punctual",
                                 "none")
                        )
         ) %>%
  group_by(language,background_noise) %>%
  summarize(hours = sum(discourse_dur)/3600,
            num_discourses = n())
d_discourses_noise_summary <- d_discourses_noise %>%
  group_by(background_noise) %>%
  summarize(hours = sum(hours),
            num_discourses=sum(num_discourses),
            num_languages=n_distinct(language))

d_discourses_quality_noise <- d_discourses %>%
  mutate(sound_quality = factor(sound_quality,
                                levels=c("bad",
                                         "medium",
                                         "good")),
         background_noise = factor(background_noise,
                        levels=c("constant",
                                 "punctual",
                                 "none_from_doreco_metadata"),
                        labels=c("constant",
                                 "punctual",
                                 "none"),
                        ),
         quality_noise = interaction(background_noise,sound_quality)) %>%
  group_by(language,quality_noise) %>%
  summarize(hours = sum(discourse_dur)/3600,
            num_discourses = n())

#d_discourses_genres
#d_discourses_quality
#d_discourses_noise

d_discourses_genres_summary
d_discourses_quality_summary
d_discourses_noise_summary

```


```{r}
d_languages %>%
  ggplot(aes(y=language,x=num_speakers)) +
  geom_bar(stat = "identity") + 
  scale_x_continuous(n.breaks = 17, 
                     limits = c(0,32)
                     ) +
  labs(title = "Number of speakers per language")

d_discourses_speakers %>%
  mutate(num_speakers = as.factor(num_speakers)) %>%
  ggplot(aes(y=language,
             x=hours,
             color = num_speakers)) +
  geom_point() +
  scale_x_continuous(n.breaks = 8) +
  labs(title = "Number of hours per discourse per language",
       x = "hours / discourse")

d_discourses_genres %>%
  ggplot(aes(x = hours, 
             y = language, 
             fill = genre)) +
  geom_bar(stat = "identity") +
  scale_x_continuous(n.breaks = 8) +
  scale_fill_manual(values = c("traditional narrative"="violet",
                               "personal narrative"="yellowgreen",
                               "conversation"="gold",
                               "stimulus retelling"="goldenrod",
                               "procedural"="orangered") ) +
  labs(title = "Number of hours per language by genre")

d_discourses_quality %>%
  ggplot(aes(x = hours, 
             y = language, 
             fill = sound_quality)) +
  geom_bar(stat = "identity") +
  scale_x_continuous(n.breaks = 8) +
  scale_fill_manual(values = c("bad" = "red2", "medium"="orange", "good"="green3") ) +
  labs(title = "Number of hours per language by sound quality")

d_discourses_noise %>%
  ggplot(aes(x = hours, 
             y = language, 
             fill = background_noise)) +
  geom_bar(stat = "identity") +
  scale_x_continuous(n.breaks = 8) +
  scale_fill_manual(values = c("constant" = "yellow3",
                               "punctual" = "purple",
                               "none" = "cyan4") ) +
  labs(title = "Number of hours per language by background noise")

d_discourses_quality_noise %>%
  ggplot(aes(x = hours, 
             y = language, 
             fill = quality_noise)) +
  geom_bar(stat = "identity") +
  scale_x_continuous(n.breaks = 8) +
  scale_fill_manual(values = c("constant.bad" = "darkred",
                               "punctual.bad" = "red",
                               "none.bad" = "indianred",
                               "constant.medium" = "darkorange",
                               "punctual.medium" = "orange",
                               "none.medium" = "yellow",
                               "constant.good" = "purple",
                               "punctual.good" = "forestgreen",
                               "none.good" = "green") ) +
  labs(title = "Number of hours per language by quality and background noise")

```




## Phone categories

Make some summaries to look at phone type


```{r}
d %>%
  group_by(phone_type) %>%
  summarize(num_langs = n_distinct(language),
            num_phones = n_distinct(phone),
            count = n()) %>%
  arrange(num_langs)
  
  
d %>%
  ggplot(aes(y = language, 
             fill = phone_type)) +
  geom_bar() +
  scale_x_continuous(n.breaks = 8)
```
```{r}
d %>%
  filter(phone_type %in% c("click",
                           #"click_contour",
                           #"nasal_double",
                           #"plosive_double",
                           "syllabic_consonant")) %>%
  group_by(phone_type,language) %>%
  summarize(count = n())



d %>%
  filter(language %in% filter(d,phone_type=="click")$language,
         segtype == "consonant") %>%
  group_by(language) %>%
  mutate(total_c = n()) %>%
  ungroup() %>%
  group_by(language,phone_type,phone_ipa) %>%
  filter(phone_type %in% c("click")) %>%
  summarize(count = n(),
            total_c = total_c[1]) %>%
  mutate(percent_c = round(100*count/total_c,2)) %>%
  arrange(language,desc(percent_c)) %>%
  select(-total_c)


# d %>%
#   filter(language %in% filter(d,phone_type=="plosive_double")$language,
#          segtype == "consonant") %>%
#   group_by(language) %>%
#   mutate(total_c = n()) %>%
#   ungroup() %>%
#   group_by(language,phone_type) %>%
#   filter(phone_type %in% c("nasal_double","plosive_double")) %>%
#   summarize(count = n(),
#             total_c = total_c[1]) %>%
#   mutate(percent_c = round(100*count/total_c,2)) %>%
#   arrange(language,desc(percent_c)) %>%
#   select(-total_c)

d %>%
  filter(language %in% filter(d,phone_type=="syllabic_consonant")$language,
         segtype == "syllabic") %>%
  group_by(language) %>%
  mutate(total_v = n()) %>%
  ungroup() %>%
  group_by(language,phone_type,phone_ipa) %>%
  filter(phone_type %in% c("syllabic_consonant")) %>%
  summarize(count = n(),
            total_v = total_v[1]) %>%
  mutate(percent_v = round(100*count/total_v,2)) %>%
  arrange(language,desc(percent_v)) %>%
  select(-total_v)
```

Make some summaries to look at vowels and consonants




```{r}
d_vowels_plot <- d %>%
  filter(segtype_syllabic == 'vowel')

d_consonants_plot <- d %>%
  filter(segtype == "consonant")

ggplot(d_consonants_plot, aes(x = "",
                     y = phone_dur,
                     fill = phone_type)) +
  scale_y_log10(n.breaks = 10) +
  geom_boxplot() +
  labs(title = "Consonant durations by phone type.")

ggplot(d_vowels_plot, aes(x = "",
                     y = phone_dur,
                     fill = phone_type)) +
  scale_y_log10(n.breaks = 10) +
  geom_boxplot() +
  labs(title = "Vowel durations by phone type.")

ggplot(d_vowels_plot, aes(x = vowel_length_contrast,
                     y = phone_dur,
                     fill = vlcLn)) +
  scale_fill_manual(values = c("blue","limegreen","magenta")) +
  scale_y_log10(n.breaks = 10) +
  geom_boxplot() +
  labs(title = "Vowel durations by length contrast.")
```
```{r}
d_vowels_plot %>%
  group_by(phone_type) %>%
  summarize(count = n())
```


```{r warning=FALSE}
d_monos_plot <- d_phones_filt_2 %>%
  filter(segtype_syllabic == 'vowel') %>%
  filter(phone_type != 'vowel_diph') %>%
  mutate(phone_short = str_replace(phone_ipa,'ː',''))

d_monos_plot %>%
  group_by(vowel_length_contrast,language,phone_short,segtype_vowel_length_revised) %>%
  summarize() %>%
  group_by(vowel_length_contrast,language) %>%
  summarize(count = n()) %>%
  arrange(desc(vowel_length_contrast),count)


for (l in d_monos_plot$languages) {
d_monos_plot %>%
  filter(language == l) %>%
  ggplot() +
  facet_grid(phone_short~speaker) +
  geom_point(aes(x=phone_dur_uttr,y=phone_dur,fill=segtype_vowel_length_revised),pch=21,stroke=NA) +   geom_smooth(aes(x=phone_dur_uttr,y=phone_dur,fill=segtype_vowel_length_revised,col=segtype_vowel_length_revised),method='gam') +
  scale_x_log10() +scale_y_log10() +
  scale_color_manual(values = c('long'='blue','short'='limegreen'),guide='legend') +
  scale_fill_manual(values = c('long'='royalblue','short'='green'),guide='none') +
guides(col = guide_legend(title = "Length",
                          override.aes = list(lwd = 2),fill=NA))
}
```

## Visualizations for filtering.

First round of filtering.

```{r}
d_uttr %>%
  ggplot(aes(x=num_sylls_uttr_init)) +
  facet_wrap(~language) +
  geom_histogram() +
  geom_vline(xintercept = c(threshold_sylls_min),
             color = "red",
             linetype = "dashed") +
  scale_x_log10(breaks = c(1,3,10,30,100,
                           threshold_sylls_min)) +
  labs(title = "Utterances by number of syllables per utterance.")

d_uttr %>%
  filter(num_sylls_uttr_init <= threshold_sylls_min) %>%
  mutate(num_sylls = factor(num_sylls_uttr_init)) %>%
  mutate(language = reorder(language, desc(language))) %>%
  ggplot(aes(y=language)) +
  geom_bar(aes(fill = num_sylls)) +
  labs(title = "Utterances removed for having too few syllables.")
  


d_uttr_filt_1.1 %>%
  ggplot(aes(x=phone_dur_uttr_init)) +
  facet_wrap(~language) +
  geom_density(bw = 0.01,fill="grey50") +
  geom_vline(xintercept = c(threshold_uttr_min,threshold_uttr_max),
             color = "red",
             linetype = "dashed") +
  scale_x_log10(breaks = c(threshold_uttr_min,
                           0.08,
                           threshold_uttr_max)) +
  labs(title = "Utterances after syllable filtering by average phone duration.")

d_uttr_filt_1.1 %>%
  filter(phone_dur_uttr_init <= threshold_uttr_min |
         phone_dur_uttr_init >= threshold_uttr_max) %>%
  mutate(error_type = ifelse(phone_dur_uttr_init <= threshold_uttr_min,
                           "below_min","above_max")) %>%
  mutate(language = reorder(language, desc(language))) %>%
  ggplot(aes(y=language)) +
  geom_bar(aes(fill=error_type)) +
  labs(title = "Utterances removed for having too high or low average phone duration.")



d_uttr_filt_1.2 %>%
  ggplot(aes(x=phone_dur_uttr_init)) +
  facet_wrap(~language) +
  geom_density(bw = 0.01,fill="grey50") +
  geom_vline(aes(xintercept = threshold_uttr_by_lang_min),
             color = "blue",
             linetype = "dashed") +
  geom_vline(aes(xintercept = threshold_uttr_by_lang_max),
             color = "blue",
             linetype = "dashed") +
  scale_x_log10(
    breaks=c(round(min(d_uttr_filt_1.2$threshold_uttr_by_lang_min),3),
             0.075,
             round(max(d_uttr_filt_1.2$threshold_uttr_by_lang_max),3))
    ) +
  labs(title = "Utterances after bulk filtering by average phone duration.")

d_uttr_filt_1.2 %>%
  filter(phone_dur_uttr_init <= threshold_uttr_by_lang_min |
         phone_dur_uttr_init >= threshold_uttr_by_lang_max) %>%
  mutate(error_type = ifelse(phone_dur_uttr_init <= threshold_uttr_by_lang_min,
                           "below_min","above_max")) %>%
  mutate(language = reorder(language, desc(language))) %>%
  ggplot(aes(y=language)) +
  geom_bar(aes(fill=error_type)) +
  labs(title = "Utterances removed for having outlying average phone duration by language.")


d_phones_filt_1.1 %>%
  ggplot(aes(x=phone_dur)) +
  facet_wrap(~language) +
  geom_density(bw = 0.04,fill="grey50") +
  geom_vline(xintercept = c(threshold_individual_min,threshold_individual_max),
             color = "red",
             linetype = "dashed") +
  scale_x_log10(
    breaks=c(threshold_individual_min,
             0.1,
             threshold_individual_max)
    ) +
  labs(title = "Phone durations after utterance-level filtering.")
  
d_phones_filt_1.1 %>%
  filter(phone_dur <= threshold_individual_min |
         phone_dur >= threshold_individual_max) %>%
  mutate(error_type = ifelse(phone_dur <= threshold_individual_min,
                           "below_min","above_max")) %>%
  mutate(language = reorder(language, desc(language))) %>%
  ggplot(aes(y=language)) +
  geom_bar(aes(fill=error_type)) +
  labs(title = "Phones removed for having too high or too low duration.")
```

Second round of filtering.

```{r warning=FALSE}

d_uttr_filt_1.4 <- d_phones_filt_1 %>%
  group_by(language,speaker,uttr_id) %>%
  summarize(num_phones_uttr = num_phones_uttr[1],
            num_phones_uttr_removed = num_phones_uttr_removed[1],
            prop_phones_uttr_removed = prop_phones_uttr_removed[1])

d_uttr_filt_1.4 %>%
  ggplot(aes(x=num_phones_uttr)) +
  facet_wrap(.~language) +
  geom_histogram() +
  geom_vline(aes(xintercept = threshold_num_phones_uttr),
             col = "red",
             linetype = "dashed") +
  scale_x_log10(breaks = c(1,3,10,30,100,
                           threshold_num_phones_uttr)) +
  labs(title = "Utterances by number of phones remaining after filtering (v1).")


d_uttr_filt_1.4 %>%
  filter(num_phones_uttr < threshold_num_phones_uttr) %>%
  ggplot(aes(y=reorder(language, desc(language)))) +
  geom_bar(aes(fill=num_phones_uttr_removed==0)) +
  labs(title = "Utterances removed for having too few phones remaining after filtering (v1).")

d_uttr_filt_1.4 %>%
  filter(num_phones_uttr_removed != 0) %>%
  group_by(num_phones_uttr_removed,prop_phones_uttr_removed) %>%
  summarize(num_uttr = n()) %>%
  ggplot(aes(x=num_phones_uttr_removed,prop_phones_uttr_removed,size=num_uttr)) +
  #facet_wrap(.~language) +
  geom_point(alpha=0.3) +
  geom_hline(aes(yintercept = threshold_prop_phones_removed),
             col = "red",
             linetype = "dashed") +
  geom_vline(aes(xintercept = threshold_num_phones_removed),
             col = "red",
             linetype = "dashed") +
  labs(title = "Utterances with ≥1 phone removed by number and proportion removed (v2).")

d_uttr_filt_1.4 %>%
  filter(num_phones_uttr_removed >= threshold_num_phones_removed |
         prop_phones_uttr_removed >= threshold_prop_phones_removed) %>%
  mutate(error_type = case_when(
    num_phones_uttr_removed < threshold_num_phones_removed ~ "proportion",
    prop_phones_uttr_removed < threshold_prop_phones_removed ~ "number",
    num_phones_uttr_removed >= threshold_num_phones_removed &
         prop_phones_uttr_removed >= threshold_prop_phones_removed ~ "both"
  )) %>%
  mutate(language = reorder(language, desc(language))) %>%
  ggplot(aes(y=language)) +
  geom_bar(aes(fill=error_type)) +
  #scale_x_continuous(breaks = c(0,2,4,6,8,10,12)) +
  labs(title = "Utterances removed for having too many phones removed (v2).")
```


Third round of filtering.



```{r warning=FALSE}


d_uttr_filt_2 %>%
  ggplot(aes(x=phone_dur_uttr)) +
  facet_wrap(~language) +
  geom_density(bw = 0.01,fill="grey50") +
  # geom_vline(aes(xintercept = threshold_uttr_by_lang_2_min),
  #            color = "darkgreen",
  #            linetype = "solid") +
  # geom_vline(aes(xintercept = threshold_uttr_by_lang_2_max),
  #            color = "darkgreen",
  #            linetype = "solid") +
  geom_vline(aes(xintercept = threshold_uttr_by_lang_min),
             color = "blue",
             linetype = "dashed") +
  geom_vline(aes(xintercept = threshold_uttr_by_lang_max),
             color = "blue",
             linetype = "dashed") +
  scale_x_log10(
    breaks=c(round(min(d_uttr_filt_1.2$threshold_uttr_by_lang_min),3),
             0.075,
             round(max(d_uttr_filt_1.2$threshold_uttr_by_lang_min),3))
    ) +
  labs(title = "Utterances after bulk filtering by average phone duration.")

d_uttr_filt_2 %>%
  filter(phone_dur_uttr <= threshold_uttr_by_lang_min |
         phone_dur_uttr >= threshold_uttr_by_lang_max) %>%
  mutate(error_type = ifelse(phone_dur_uttr <= threshold_uttr_by_lang_min,
                           "below_min","above_max")) %>%
  ggplot(aes(y=reorder(language, desc(language)))) +
  geom_bar(aes(fill = error_type)) +
  labs(title = "Utterances removed for having outlying recalculated average phone duration.")
```


```{r warning=FALSE}
d_phones_unfilt %>%
  ggplot(aes(x = phone_dur_uttr,
             y = phone_dur)) +
  facet_wrap(~language) +
  geom_point(aes(fill=segtype), alpha=0.3, shape=21, stroke=0) +
  scale_fill_manual(values=c("purple","orange")) +
  geom_smooth(aes(col=segtype), method="gam", se=F) +
  scale_color_manual(values=c("purple4","darkorange2")) +
  scale_x_log10(limits=c(min(d_phones_unfilt$phone_dur_uttr),
                              max(d_phones_unfilt$phone_dur_uttr))) +
  scale_y_continuous(limits=c(min(d_phones_unfilt$phone_dur),
                              1)) +
  labs(title = "Unfiltered duration data",
       x = "Initial Mean Duration by Utterance (Inverse Rate)")
  
d_phones_filt_1 %>%
  ggplot(aes(x = phone_dur_uttr_init,
             y = phone_dur)) +
  facet_wrap(~language) +
  geom_point(aes(fill=segtype), alpha=0.3, shape=21, stroke=0) +
  scale_fill_manual(values=c("purple","orange")) +
  geom_smooth(aes(col=segtype), method="gam", se=F) +
  scale_color_manual(values=c("purple4","darkorange2")) +
  scale_x_log10(limits=c(min(d_phones_unfilt$phone_dur_uttr),
                              max(d_phones_unfilt$phone_dur_uttr))) +
  scale_y_continuous(limits=c(min(d_phones_unfilt$phone_dur),
                              1)) +
  labs(title = "Duration data after first round of filtering (by phone duration and initial rate)",
       x = "Initial Mean Duration by Utterance (Inverse Speech Rate)")

d_phones_filt_1 %>%
  ggplot(aes(x = phone_dur_uttr,
             y = phone_dur)) +
  facet_wrap(~language) +
  geom_point(aes(fill=segtype), alpha=0.3, shape=21, stroke=0) +
  scale_fill_manual(values=c("purple","orange")) +
  geom_smooth(aes(col=segtype), method="gam", se=F) +
  scale_color_manual(values=c("purple4","darkorange2")) +
  scale_x_log10(limits=c(min(d_phones_unfilt$phone_dur_uttr),
                              max(d_phones_unfilt$phone_dur_uttr))) +
  scale_y_continuous(limits=c(min(d_phones_unfilt$phone_dur),
                              1)) +
  labs(title = "Duration data after first round of filtering (by phone duration and initial rate)",
       x = "Recalculated Mean Duration by Utterance (Inverse Speech Rate)")

d_phones_filt_2 %>%
  ggplot(aes(x = phone_dur_uttr,
             y = phone_dur)) +
  facet_wrap(~language) +
  geom_point(aes(fill=segtype), alpha=0.3, shape=21, stroke=0) +
  scale_fill_manual(values=c("purple","orange")) +
  geom_smooth(aes(col=segtype), method="gam", se=F) +
  scale_color_manual(values=c("purple4","darkorange2")) +
  scale_x_log10(limits=c(min(d_phones_unfilt$phone_dur_uttr),
                              max(d_phones_unfilt$phone_dur_uttr))) +
  scale_y_continuous(limits=c(min(d_phones_unfilt$phone_dur),1)) +
  labs(title = "Duration data after second round of filtering (by phones removed per utterance)",
       x = "Recalculated Mean Duration by Utterance (Inverse Speech Rate)")

d_phones_filt_3 %>%
  ggplot(aes(x = phone_dur_uttr,
             y = phone_dur)) +
  facet_wrap(~language) +
  geom_point(aes(fill=segtype), alpha=0.3, shape=21, stroke=0) +
  scale_fill_manual(values=c("purple","orange")) +
  geom_smooth(aes(col=segtype), method="gam", se=F) +
  scale_color_manual(values=c("purple4","darkorange2")) +
  scale_x_log10(limits=c(min(d_phones_unfilt$phone_dur_uttr),
                              max(d_phones_unfilt$phone_dur_uttr))) +
  scale_y_continuous(limits=c(min(d_phones_unfilt$phone_dur),1)) + 
  labs(title = "Duration data after third round of filtering (by recalculated rate)",
       x = "Recalculated Mean Duration by Utterance (Inverse Speech Rate)")
```



## Further Visualizations

```{r}
d_phones_filt_language_nums <- d_phones_filt %>%
  group_by(language) %>%
  summarize() %>%
  mutate(language_number = row_number())
  

d_phones_filt_speaker_nums <- d_phones_filt %>%
  group_by(language,sex,speaker) %>%
  summarize(phone_dur_uttr_av=mean(phone_dur_uttr)) %>%
  group_by(language) %>%
  mutate(num_speakers = n()) %>%
  mutate(phone_dur_uttr_av_scaled = if_else(num_speakers > 1,
                                            (phone_dur_uttr_av-min(phone_dur_uttr_av))/
                                              (max(phone_dur_uttr_av)-min(phone_dur_uttr_av)),
                                            0)) %>%
  arrange(phone_dur_uttr_av_scaled) %>%
  mutate(speaker_number = as.factor(row_number()))

d_phones_filt_speakers <- left_join(d_phones_filt,d_phones_filt_speaker_nums,
                                    by = c("language","sex","speaker"))
d_phones_filt_speakers <- left_join(d_phones_filt_speakers,d_phones_filt_language_nums,
                                    by = "language")


d_phones_filt_speakers %>%
  group_by(language,sex,speaker) %>%
  summarize(speaker_number = speaker_number[1],num_speakers = num_speakers[1]) %>%
  group_by(language) %>%
  arrange(num_speakers)
```

```{r}
d_phones_filt_speakers %>%
  group_by(language,language_number,sex,speaker,speaker_number,phone_dur_uttr_av_scaled,vlcLn,genre,uttr_id) %>%
  summarize(phone_dur_uttr = phone_dur_uttr[1]) %>%
  filter(language_number <= 5) %>%
  ggplot() +
  facet_grid(genre~language) +
  geom_density(aes(x=phone_dur_uttr,groups=speaker,fill=sex,after_stat(scaled)),linewidth=0,alpha=0.5)

d_phones_filt_speakers %>%
  group_by(language,language_number,sex,speaker,speaker_number,phone_dur_uttr_av_scaled,vlcLn,genre,uttr_id) %>%
  summarize(phone_dur_uttr = phone_dur_uttr[1]) %>%
  filter(language_number > 5 &
           language_number <= 10) %>%
  ggplot() +
  facet_grid(genre~language) +
  geom_density(aes(x=phone_dur_uttr,groups=speaker,fill=sex,after_stat(scaled)),linewidth=0,alpha=0.5)

d_phones_filt_speakers %>%
  group_by(language,language_number,sex,speaker,speaker_number,phone_dur_uttr_av_scaled,vlcLn,genre,uttr_id) %>%
  summarize(phone_dur_uttr = phone_dur_uttr[1]) %>%
  filter(language_number > 10 &
           language_number <= 15) %>%
  ggplot() +
  facet_grid(genre~language) +
  geom_density(aes(x=phone_dur_uttr,groups=speaker,fill=sex,after_stat(scaled)),linewidth=0,alpha=0.5)

d_phones_filt_speakers %>%
  group_by(language,language_number,sex,speaker,speaker_number,phone_dur_uttr_av_scaled,vlcLn,genre,uttr_id) %>%
  summarize(phone_dur_uttr = phone_dur_uttr[1]) %>%
  filter(language_number > 15 &
           language_number <= 20) %>%
  ggplot() +
  facet_grid(genre~language) +
  geom_density(aes(x=phone_dur_uttr,groups=speaker,fill=sex,after_stat(scaled)),linewidth=0,alpha=0.5)

d_phones_filt_speakers %>%
  group_by(language,language_number,sex,speaker,speaker_number,phone_dur_uttr_av_scaled,vlcLn,genre,uttr_id) %>%
  summarize(phone_dur_uttr = phone_dur_uttr[1]) %>%
  filter(language_number > 20) %>%
  ggplot() +
  facet_grid(genre~language) +
  geom_density(aes(x=phone_dur_uttr,groups=speaker,fill=sex,after_stat(scaled)),linewidth=0,alpha=0.5) +
  scale_color_manual(values = c("traditional narrative"="violet",
                               "personal narrative"="yellowgreen",
                               "conversation"="gold",
                               "stimulus retelling"="goldenrod",
                               "procedural"="orangered") )

```

