---
title: "doreco_stretchiness_analysis"
author: "Noah Luntzlara"
date: "2025-10-01"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Set Up

Load packages.

```{r}
library(tidyverse)
# library(ggbeeswarm)
# library(ggpubr)
library(forcats)

library(mgcv)
library(itsadug)

library(geomtextpath)
library(RColorBrewer)
library(viridis)

library(broom)
```

Read in the data.

```{r}

d_all <- read_csv("../doreco_data/combined_doreco_durations_filtered.csv")

# d_all_unfilt <- read_csv("../doreco_data/combined_doreco_durations_unfiltered.csv")

```



Data frame for VL models

```{r}
d_vl <- d_all %>%
  filter(segtype_syllabic == "vowel",
         segtype_vowel_dipthongality == "monophthong")

#d_vl$segtype_o <- as.ordered(d_vl$segtype)
#contrasts(d_vl$segtype_o) <- "contr.treatment"

d_vl$speaker_f <- as.factor(d_vl$speaker)
d_vl$speakerSegtype <- interaction(d_vl$speaker, d_vl$segtype)

d_vl$language_f <- as.factor(d_vl$language)
d_vl$languageSegtype <- interaction(d_vl$language, d_vl$segtype)

d_vl$phone_type_f <- as.factor(d_vl$phone_type)
#d_vl$phone_type_new_f <- as.factor(d_vl$phone_type_new)


d_vl$vlcLn_f <- factor(d_vl$vlcLn)
d_vl$vlcLn_f <- relevel(d_vl$vlcLn_f, ref = "no_contrast")
d_vl$vlcLn_o <- as.ordered(d_vl$vlcLn_f)
contrasts(d_vl$vlcLn_o) <- "contr.treatment"

d_vl$vht_f <- as.factor(d_vl$segtype_vowel_height)
d_vl$vht_o <- as.ordered(d_vl$vht_f)
contrasts(d_vl$vht_o) <- "contr.treatment"

d_vl$syll_pos_f <- as.factor(d_vl$syll_pos)
d_vl$syll_pos_o <- as.ordered(d_vl$syll_pos_f) # ??? not ordered?
contrasts(d_vl$syll_pos_o) <- "contr.treatment" #added, re-run

```


Lists of languages by vowel length typology
```{r}
d_vl_language_types <- d_vl %>%
  group_by(language,vowel_length_contrast) %>%
  summarize()

languages <- d_vl_language_types$language

languages_no_contrast <- filter(d_vl_language_types,
                                vowel_length_contrast=="none")$language

languages_restricted <- filter(d_vl_language_types,
                                vowel_length_contrast=="restricted")$language

languages_unrestricted <- filter(d_vl_language_types,
                                vowel_length_contrast=="unrestricted")$language
```

# GAMMs for VL comparisons

## Fitting VL models

```{r}
# vln_mod_mono_baseline <- bam(phone_dur ~
#   vlcLn_o +
#   #vht_o +
#   #syll_pos_o +
#   s(phone_dur_uttr, bs = "cr") +
#   s(phone_dur_uttr, by = vlcLn_o, bs = "cr") +
#   #s(phone_dur_uttr, by = vht_o, bs = "cr") +
#   #s(phone_dur_uttr, by = syll_pos_o, bs = "cr") +
#   s(phone_dur_uttr, speaker_f, bs = "fs", m = 1, xt = list("cr")) +
#   s(phone_dur_uttr, language_f, bs = "fs", m = 1, xt = list("cr")),
#   data = d_vl,
#   #weights=d_vl$weight / mean(d_vl$weight),
#   discrete = TRUE,
#   nthreads = 8
# )
# 
# saveRDS(vln_mod_mono_baseline, "vln_mod_mono_baseline.rds")
# vln_mod_mono_baseline <- readRDS("vln_mod_mono_baseline.rds")
# summary(vln_mod_mono_baseline)
# 
# vln_mod_mono_height <- bam(phone_dur ~
#   vlcLn_o +
#   vht_o +
#   #syll_pos_o +
#   s(phone_dur_uttr, bs = "cr") +
#   s(phone_dur_uttr, by = vlcLn_o, bs = "cr") +
#   s(phone_dur_uttr, by = vht_o, bs = "cr") +
#   #s(phone_dur_uttr, by = syll_pos_o, bs = "cr") +
#   s(phone_dur_uttr, speaker_f, bs = "fs", m = 1, xt = list("cr")) +
#   s(phone_dur_uttr, language_f, bs = "fs", m = 1, xt = list("cr")),
#   data = d_vl,
#   #weights=d_vl$weight / mean(d_vl$weight),
#   discrete = TRUE,
#   nthreads = 8
# )
# 
# saveRDS(vln_mod_mono_height, "vln_mod_mono_height.rds")
# vln_mod_mono_height <- readRDS("vln_mod_mono_height.rds")
# summary(vln_mod_mono_height)
# 
# vln_mod_mono_pos <- bam(phone_dur ~
#   vlcLn_o +
#   #vht_o +
#   syll_pos_o +
#   s(phone_dur_uttr, bs = "cr") +
#   s(phone_dur_uttr, by = vlcLn_o, bs = "cr") +
#   #s(phone_dur_uttr, by = vht_o, bs = "cr") +
#   s(phone_dur_uttr, by = syll_pos_o, bs = "cr") +
#   s(phone_dur_uttr, speaker_f, bs = "fs", m = 1, xt = list("cr")) +
#   s(phone_dur_uttr, language_f, bs = "fs", m = 1, xt = list("cr")),
#   data = d_vl,
#   #weights=d_vl$weight / mean(d_vl$weight),
#   discrete = TRUE,
#   nthreads = 8
# )
# 
# saveRDS(vln_mod_mono_pos, "vln_mod_mono_pos.rds")
# vln_mod_mono_pos <- readRDS("vln_mod_mono_pos.rds")
# summary(vln_mod_mono_pos)
# 
# vln_mod_mono_full <- bam(phone_dur ~
#   vlcLn_o +
#   vht_o +
#   syll_pos_o +
#   s(phone_dur_uttr, bs = "cr") +
#   s(phone_dur_uttr, by = vlcLn_o, bs = "cr") + # keep
#   s(phone_dur_uttr, by = vht_o, bs = "cr") + # keep
#   s(phone_dur_uttr, by = syll_pos_o, bs = "cr") +
#   s(phone_dur_uttr, speaker_f, bs = "fs", m = 1, xt = list("cr")) + # keep
#   s(phone_dur_uttr, language_f, bs = "fs", m = 1, xt = list("cr")), # replace with random smooths by langauge x vlcLn and langauge x vht
#   data = d_vl,
#   #weights=d_vl$weight / mean(d_vl$weight),
#   discrete = TRUE,
#   nthreads = 8
# )
# 
# saveRDS(vln_mod_mono_full, "vln_mod_mono_full.rds")
# vln_mod_mono_full <- readRDS("vln_mod_mono_full.rds")
# summary(vln_mod_mono_full)
```
Just reading the models from files

```{r}
vln_mod_mono_baseline <- readRDS("vln_mod_mono_baseline.rds")
vln_mod_mono_height <- readRDS("vln_mod_mono_height.rds")
vln_mod_mono_pos <- readRDS("vln_mod_mono_pos.rds")
vln_mod_mono_full <- readRDS("vln_mod_mono_full.rds")

# plot diff (check Marci's tutorial=evaluating GAMMs paper 2019 or other)
# if you want to put a significance value, simultaneous confidence interval
```


## Model Predictions for VL Models

Model predictions for all VL models


```{r}
newdat_vln_mono <- expand.grid(
  phone_dur_uttr = seq(
    min(d_vl$phone_dur_uttr),
    max(d_vl$phone_dur_uttr),
    length.out = 100
  ),
  vht = d_vl$segtype_vowel_height[1], #unique(d_vl$segtype_vowel_height)
  syll_pos = unique(d_vl$syll_pos),
  speaker = d_vl$speaker[1],
  language = unique(d_vl$language),
  vlcLn = unique(d_vl$vlcLn)
) %>%
  mutate(vowel_length_contrast = case_when(
           language %in% languages_no_contrast ~ "none",
           language %in% languages_restricted ~ "restricted",
           language %in% languages_unrestricted ~ "unrestricted")) %>%
  filter(vowel_length_contrast == "none" & vlcLn == "no_contrast" |
           (vowel_length_contrast %in% c("unrestricted") &
             vlcLn %in% c("contrast_long","contrast_short")) |
           ((vowel_length_contrast %in% c("restricted") &
             ((vlcLn %in% c("contrast_long","contrast_short") & syll_pos == "nonfinal") |
                (vlcLn %in% c("no_contrast") & syll_pos == "final"))))
         )

newdat_vln_mono$vlcLn_o <- as.ordered(newdat_vln_mono$vlcLn)
contrasts(newdat_vln_mono$vlcLn_o) <- "contr.treatment"

newdat_vln_mono$vht_o <- as.ordered(newdat_vln_mono$vht)
#contrasts(newdat_vln_mono$vht_o) <- "contr.treatment"

newdat_vln_mono$syll_pos_o <- as.ordered(newdat_vln_mono$syll_pos)
contrasts(newdat_vln_mono$syll_pos_o) <- "contr.treatment"

newdat_vln_mono$language_f <- as.factor(newdat_vln_mono$language)
newdat_vln_mono$speaker_f <- as.factor(newdat_vln_mono$speaker)

preds_vln_mono_baseline <- predict(vln_mod_mono_baseline, newdat_vln_mono,
  # the following is used when predicting without speakers
  exclude=c("s(phone_dur_uttr,speaker_f)",
            "s(phone_dur_uttr):vht_ononhigh"),
  se.fit = TRUE
)

preds_vln_mono_height <- predict(vln_mod_mono_height, newdat_vln_mono,
  # the following is used when predicting without speakers
  exclude=c("s(phone_dur_uttr,speaker_f)",
            "s(phone_dur_uttr):vht_ononhigh"),
  se.fit = TRUE
)

preds_vln_mono_pos <- predict(vln_mod_mono_pos, newdat_vln_mono,
  # the following is used when predicting without speakers
  exclude=c("s(phone_dur_uttr,speaker_f)",
            "s(phone_dur_uttr):vht_ononhigh"),
  se.fit = TRUE
)

preds_vln_mono_full <- predict(vln_mod_mono_full, newdat_vln_mono,
  # the following is used when predicting without speakers
  exclude=c("s(phone_dur_uttr,speaker_f)",
            "s(phone_dur_uttr):vht_ononhigh"
            ),
  se.fit = TRUE
)

# baseline

newdat_vln_mono_baseline <- newdat_vln_mono

newdat_vln_mono_baseline$dur <- preds_vln_mono_baseline$fit
newdat_vln_mono_baseline$ul <- preds_vln_mono_baseline$fit + preds_vln_mono_baseline$se.fit * 1.96
newdat_vln_mono_baseline$ll <- preds_vln_mono_baseline$fit - preds_vln_mono_baseline$se.fit * 1.96

newdat_vln_mono_baseline$model <- "vln_mod_mono_baseline"

# vowel height

newdat_vln_mono_height <- newdat_vln_mono

newdat_vln_mono_height$dur <- preds_vln_mono_height$fit
newdat_vln_mono_height$ul <- preds_vln_mono_height$fit + preds_vln_mono_height$se.fit * 1.96
newdat_vln_mono_height$ll <- preds_vln_mono_height$fit - preds_vln_mono_height$se.fit * 1.96

newdat_vln_mono_height$model <- "vln_mod_mono_height"


# syllable position

newdat_vln_mono_pos <- newdat_vln_mono

newdat_vln_mono_pos$dur <- preds_vln_mono_pos$fit
newdat_vln_mono_pos$ul <- preds_vln_mono_pos$fit + preds_vln_mono_pos$se.fit * 1.96
newdat_vln_mono_pos$ll <- preds_vln_mono_pos$fit - preds_vln_mono_pos$se.fit * 1.96

newdat_vln_mono_pos$model <- "vln_mod_mono_pos"

# syllable position and vowel height

newdat_vln_mono_full <- newdat_vln_mono

newdat_vln_mono_full$dur <- preds_vln_mono_full$fit
newdat_vln_mono_full$ul <- preds_vln_mono_full$fit + preds_vln_mono_full$se.fit * 1.96
newdat_vln_mono_full$ll <- preds_vln_mono_full$fit - preds_vln_mono_full$se.fit * 1.96

newdat_vln_mono_full$model <- "vln_mod_mono_full"



newdat_vln_mono_all <- bind_rows(
  newdat_vln_mono_baseline,
  newdat_vln_mono_height,
  newdat_vln_mono_pos,
  newdat_vln_mono_full
)

newdat_vln_mono_all %>% write_csv("./newdat_vln_mono_all.csv")
newdat_vln_mono_all <- read_csv("./newdat_vln_mono_all.csv")


newdat_vln_mono_all_nopos <- newdat_vln_mono_all %>%
  filter(model %in% c("vln_mod_mono_baseline", "vln_mod_mono_height")) %>%
  group_by(across(-c(syll_pos,syll_pos_o,
                     dur,ul,ll))) %>%
  summarize(dur = dur[1],
            ul = ul[1],
            ll = ll[1],
            #count = n(),
            )

newdat_vln_mono_all_pos <- newdat_vln_mono_all %>%
  filter(model %in% c("vln_mod_mono_pos", "vln_mod_mono_full"))

```

## Plotting VL Model Predictions

New data with long, short, and no contrast averages by utterance

```{r}
d_uttr_vln_mono <- d_vl %>%
  group_by(language,speaker,uttr_id) %>%
  summarize(nocontrast_dur_uttr = mean(phone_dur[vlcLn == "no_contrast"]),
            long_dur_uttr = mean(phone_dur[vlcLn == "contrast_long"]),
            short_dur_uttr = mean(phone_dur[vlcLn == "contrast_short"]),
            phone_dur_uttr = phone_dur_uttr[1],
            vowel_length_contrast = vowel_length_contrast[1]
  ) %>%
  pivot_longer(c(nocontrast_dur_uttr, long_dur_uttr, short_dur_uttr),
               names_to="measure",
               values_to="value")
```

Plot VL model predictions

```{r}

newdat_vln_mono_all %>%
  ggplot(data = ., aes(x = phone_dur_uttr)) +
  facet_wrap(. ~ factor(language,
                        levels=c(languages_no_contrast,
                                 languages_restricted,
                                 languages_unrestricted))) +
  geom_point(data = d_uttr_vln_mono,
             aes(y = value, col=measure), alpha = 0.1, pch=16) +
  geom_line(aes(y = dur, lty=model, col = vlcLn_o, lwd = syll_pos_o)) +
  # geom_ribbon(data = filter(newdat_cv_all,
  #                          model == "cv_mod_baseline",
  #                          segtype_o == "consonant"),
  #           aes(ymin = ll, ymax = ul), col = NA, fill = "skyblue",
  #           alpha = 0.5) +
  # geom_ribbon(data = filter(newdat_cv_all,
  #                          model == "cv_mod_baseline",
  #                          segtype_o == "syllabic"),
  #           aes(ymin = ll, ymax = ul), col = NA, fill = "yellow",
  #           alpha = 0.5) +
  scale_color_manual(values = c("no_contrast" = "magenta",
                                "contrast_long" = "blue",
                                "contrast_short" = "limegreen",
                                "nocontrast_dur_uttr" = "hotpink", 
                                "long_dur_uttr" = "cyan", 
                                "short_dur_uttr" = "green"),
                     guide = "none") +
  scale_linetype_manual(values = c("vln_mod_mono_baseline" = "dotted",
                                   "vln_mod_mono_height" = "dashed",
                                   "vln_mod_mono_pos" = "longdash",
                                   "vln_mod_mono_full" = "solid"
                                   )) +
  scale_linewidth_manual(values = c("final" = 1,
                                    "nonfinal" = 0.5)) +
  scale_x_log10(breaks = c(0.05, 0.07, 0.1, 0.14)) +
  scale_y_log10(breaks = c(0.05, 0.07, 0.1, 0.14),
                #limits = c(0.03, 0.2)
                ) +
  #scale_x_continuous(limits=c(0.035,0.145), 
  #                   breaks=seq(0.04,0.14,0.02),
  #                   labels=c(".04",".06",".08",".10",".12",".14")) +
  #scale_y_continuous(breaks=seq(0.04,0.20,0.02)) +
    # labels for axes
  xlab("Average phone duration by utterance") +
  labs(y = "Average C vs. V duration in s") +
  #ylab("Average V vs. C duration in s") +
  # remove clutter
  theme_minimal() +
  theme(panel.grid = element_blank(),
        axis.line = element_line(size=0.8),
        axis.ticks = element_line(size=0.8),
        axis.text = element_text(size=16, colour="black"),
        axis.title = element_text(size=18, face="bold"),
        strip.background.y = element_blank(), 
        strip.placement = "outside",
        strip.text=element_text(size=18, face="bold"))
  ggsave("doreco_graphs_v1/vl/vln_mod_mono_all.png", width=15, height=15, dpi=300)
```

# VL Model Predictions Aggregated over Languages.

## VL Model Predictions Aggregated over All Languages

```{r}

eff_sizes_vln_mono_comb <- list()

# model predictions for models with with segment inclusive speech rate measures
newdat_vln_mono_comb <- expand.grid(
  phone_dur_uttr = seq(
    min(d_vl$phone_dur_uttr),
    max(d_vl$phone_dur_uttr),
    length.out = 100
  ),
  vht = d_vl$segtype_vowel_height[1],
  syll_pos = unique(d_vl$syll_pos),
  speaker = d_vl$speaker[1],
  language = d_vl$language[1],
  vlcLn = unique(d_vl$vlcLn)
)


newdat_vln_mono_comb$vlcLn_o <- as.ordered(newdat_vln_mono_comb$vlcLn)
contrasts(newdat_vln_mono_comb$vlcLn_o) <- "contr.treatment"

newdat_vln_mono_comb$vht_o <- as.ordered(newdat_vln_mono_comb$vht)
#contrasts(newdat_vln_mono_comb$vht_o) <- "contr.treatment"

newdat_vln_mono_comb$syll_pos_o <- as.ordered(newdat_vln_mono_comb$syll_pos)
contrasts(newdat_vln_mono_comb$syll_pos_o) <- "contr.treatment"

newdat_vln_mono_comb$language_f <- as.factor(newdat_vln_mono_comb$language)
newdat_vln_mono_comb$speaker_f <- as.factor(newdat_vln_mono_comb$speaker)


preds_vln_mono_comb_baseline <- predict(vln_mod_mono_baseline, newdat_vln_mono_comb,
  # the following is used when predicting without speakers
  exclude=c("s(phone_dur_uttr,speaker_f)",
            "s(phone_dur_uttr,language_f)",
            "s(phone_dur_uttr):vht_ononhigh"),
  se.fit = TRUE
)

preds_vln_mono_comb_height <- predict(vln_mod_mono_height, newdat_vln_mono_comb,
  # the following is used when predicting without speakers
  exclude=c("s(phone_dur_uttr,speaker_f)",
            "s(phone_dur_uttr,language_f)",
            "s(phone_dur_uttr):vht_ononhigh"),
  se.fit = TRUE
)

preds_vln_mono_comb_pos <- predict(vln_mod_mono_pos, newdat_vln_mono_comb,
  # the following is used when predicting without speakers
  exclude=c("s(phone_dur_uttr,speaker_f)",
            "s(phone_dur_uttr,language_f)",
            "s(phone_dur_uttr):vht_ononhigh"),
  se.fit = TRUE
)

preds_vln_mono_comb_full <- predict(vln_mod_mono_full, newdat_vln_mono_comb,
  # the following is used when predicting without speakers
  exclude=c("s(phone_dur_uttr,speaker_f)",
            "s(phone_dur_uttr,language_f)",
            "s(phone_dur_uttr):vht_ononhigh"),
  se.fit = TRUE
)

mods_vln_mono_nopos <- c("vln_mod_mono_baseline",
                         "vln_mod_mono_height")
mods_vln_mono_pos <- c("vln_mod_mono_pos",
                       "vln_mod_mono_full")
mods_vln_mono <- c(mods_vln_mono_nopos,mods_vln_mono_pos)

# baseline

newdat_vln_mono_comb_baseline <- newdat_vln_mono_comb

newdat_vln_mono_comb_baseline$dur <- preds_vln_mono_comb_baseline$fit
newdat_vln_mono_comb_baseline$ul <- preds_vln_mono_comb_baseline$fit + preds_vln_mono_comb_baseline$se.fit * 1.96
newdat_vln_mono_comb_baseline$ll <- preds_vln_mono_comb_baseline$fit - preds_vln_mono_comb_baseline$se.fit * 1.96

newdat_vln_mono_comb_baseline$model <- "vln_mod_mono_baseline"

# vowel height

newdat_vln_mono_comb_height <- newdat_vln_mono_comb

newdat_vln_mono_comb_height$dur <- preds_vln_mono_comb_height$fit
newdat_vln_mono_comb_height$ul <- preds_vln_mono_comb_height$fit + preds_vln_mono_comb_height$se.fit * 1.96
newdat_vln_mono_comb_height$ll <- preds_vln_mono_comb_height$fit - preds_vln_mono_comb_height$se.fit * 1.96

newdat_vln_mono_comb_height$model <- "vln_mod_mono_height"

# syllable position

newdat_vln_mono_comb_pos <- newdat_vln_mono_comb

newdat_vln_mono_comb_pos$dur <- preds_vln_mono_comb_pos$fit
newdat_vln_mono_comb_pos$ul <- preds_vln_mono_comb_pos$fit + preds_vln_mono_comb_pos$se.fit * 1.96
newdat_vln_mono_comb_pos$ll <- preds_vln_mono_comb_pos$fit - preds_vln_mono_comb_pos$se.fit * 1.96

newdat_vln_mono_comb_pos$model <- "vln_mod_mono_pos"

# vowel height and syllable position

newdat_vln_mono_comb_full <- newdat_vln_mono_comb

newdat_vln_mono_comb_full$dur <- preds_vln_mono_comb_full$fit
newdat_vln_mono_comb_full$ul <- preds_vln_mono_comb_full$fit + preds_vln_mono_comb_full$se.fit * 1.96
newdat_vln_mono_comb_full$ll <- preds_vln_mono_comb_full$fit - preds_vln_mono_comb_full$se.fit * 1.96

newdat_vln_mono_comb_full$model <- "vln_mod_mono_full"


newdat_vln_mono_comb_all <- bind_rows(
  newdat_vln_mono_comb_baseline,
  newdat_vln_mono_comb_height,
  newdat_vln_mono_comb_pos,
  newdat_vln_mono_comb_full
)

newdat_vln_mono_comb_all %>% write_csv("./newdat_vln_mono_comb_all.csv")
newdat_vln_mono_comb_all <- read_csv("./newdat_vln_mono_comb_all.csv")
```


Plotting model predictions aggregated over languages.


```{r}
for (m in mods_vln_mono) {
  
  eff_sizes_vln_mono_comb_m <- eff_sizes_vln_mono_comb[[m]]
  eff_sizes_vln_mono_comb_m$phone_dur_uttr <- 0.05
  eff_sizes_vln_mono_comb_m$value <- 0.13
  
  newdat_m <- filter(newdat_vln_mono_comb_all, model==m)
  
  plot_dat <- newdat_m
  
  plot_dat %>%
  ggplot(data = ., aes(x = phone_dur_uttr)) +
  geom_point(data = d_uttr_vln_mono,
             aes(y = value, col=measure), alpha = 0.1, pch=16) +
  geom_ribbon(data = filter(newdat_m,
                            vlcLn == "no_contrast",
                            syll_pos_o == "nonfinal"),
            aes(ymin = ll, ymax = ul), col = NA, fill = "hotpink",
            alpha = 0.25) +
    geom_ribbon(data = filter(newdat_m,
                            vlcLn == "no_contrast",
                            syll_pos_o == "final"),
            aes(ymin = ll, ymax = ul), col = NA, fill = "hotpink",
            alpha = 0.5) +
  geom_ribbon(data = filter(newdat_m,
                            vlcLn == "contrast_long",
                            syll_pos_o == "nonfinal"),
            aes(ymin = ll, ymax = ul), col = NA, fill = "cyan",
            alpha = 0.25) +
  geom_ribbon(data = filter(newdat_m,
                            vlcLn == "contrast_long",
                            syll_pos_o == "final"),
            aes(ymin = ll, ymax = ul), col = NA, fill = "cyan",
            alpha = 0.5) +
  geom_ribbon(data = filter(newdat_m,
                            vlcLn == "contrast_short",
                            syll_pos_o == "nonfinal"),
            aes(ymin = ll, ymax = ul), col = NA, fill = "green",
            alpha = 0.25) +
  geom_ribbon(data = filter(newdat_m,
                            vlcLn == "contrast_short",
                            syll_pos_o == "final"),
            aes(ymin = ll, ymax = ul), col = NA, fill = "green",
            alpha = 0.5) +
  geom_line(aes(y=dur, col = vlcLn_o, lwd = syll_pos_o)) +
  # geom_text(data=eff_sizes_vln_mono_comb_m, 
  #           aes(y=value*1.2*1.2*1.2,label=paste0("β_nc = ", round(derivative_nc,2))),
  #           col="black", size=5) +
  #   geom_text(data=eff_sizes_vln_mono_comb_m, 
  #           aes(y=value*1.2*1.2,label=paste0("β_l = ", round(derivative_cl,2))),
  #           col="black", size=5) +
  #   geom_text(data=eff_sizes_vln_mono_comb_m, 
  #           aes(y=value*1.2,label=paste0("β_s = ", round(derivative_cs,2))),
  #           col="black", size=5) +
  #   geom_text(data=eff_sizes_vln_mono_comb_m, 
  #           aes(y=value,label=paste0("Δβ = ", round(eff_size,2))),
  #           col="black", size=5) +
  scale_color_manual(values = c("no_contrast" = "magenta",
                                "contrast_long" = "blue",
                                "contrast_short" = "limegreen",
                                "nocontrast_dur_uttr" = "hotpink", 
                                "long_dur_uttr" = "cyan", 
                                "short_dur_uttr" = "green"),
                     guide = "none") +
  scale_linewidth_manual(values = c("final" = 1,
                                    "nonfinal" = 0.5)) +
  scale_x_log10(breaks = c(0.05, 0.07, 0.1, 0.14)) +
  scale_y_log10(breaks = c(0.05, 0.07, 0.1, 0.14),
                #limits = c(0.03, 0.2)
                ) +
  #scale_x_continuous(limits=c(0.035,0.145), 
  #                   breaks=seq(0.04,0.14,0.02),
  #                   labels=c(".04",".06",".08",".10",".12",".14")) +
  #scale_y_continuous(breaks=seq(0.04,0.20,0.02)) +
    # labels for axes
  xlab("Average phone duration by utterance") +
  labs(y = "Average long vs short vs no-contrast monophthong duration in s") +
  #ylab("Average V vs. C duration in s") +
  # remove clutter
  theme_minimal() +
  theme(panel.grid = element_blank(),
        axis.line = element_line(size=0.8),
        axis.ticks = element_line(size=0.8),
        axis.text = element_text(size=16, colour="black"),
        axis.title = element_text(size=18, face="bold"),
        strip.background.y = element_blank(), 
        strip.placement = "outside",
        strip.text=element_text(size=18, face="bold"))
  ggsave(paste0("doreco_graphs_v1/vl/comb/", m, "_comb.png"), width=9, height=9, dpi=300)
}
```

## VL Model Predictions Aggregated by Vowel Contrast Type

```{r}

eff_sizes_vln_mono_comb2 <- list()

# model predictions for models with with segment inclusive speech rate measures
newdat_vln_mono_comb2 <- expand.grid(
  phone_dur_uttr = seq(
    min(d_vl$phone_dur_uttr),
    max(d_vl$phone_dur_uttr),
    length.out = 100
  ),
  vht = d_vl$segtype_vowel_height[1],
  syll_pos = unique(d_vl$syll_pos),
  speaker = d_vl$speaker[1],
  language = d_vl$language[1],
  vlcLn = unique(d_vl$vlcLn),
  vowel_length_contrast = unique(d_vl$vowel_length_contrast)
) %>%
  filter(vowel_length_contrast == "none" & vlcLn == "no_contrast" |
           (vowel_length_contrast %in% c("unrestricted") &
             vlcLn %in% c("contrast_long","contrast_short")) |
           ((vowel_length_contrast %in% c("restricted") &
             ((vlcLn %in% c("contrast_long","contrast_short") & syll_pos == "nonfinal") |
                (vlcLn %in% c("no_contrast") & syll_pos == "final"))))
         )


newdat_vln_mono_comb2$vlcLn_o <- as.ordered(newdat_vln_mono_comb2$vlcLn)
contrasts(newdat_vln_mono_comb2$vlcLn_o) <- "contr.treatment"

newdat_vln_mono_comb2$vht_o <- as.ordered(newdat_vln_mono_comb2$vht)
#contrasts(newdat_vln_mono_comb2$vht_o) <- "contr.treatment"

newdat_vln_mono_comb2$syll_pos_o <- as.ordered(newdat_vln_mono_comb2$syll_pos)
contrasts(newdat_vln_mono_comb2$syll_pos_o) <- "contr.treatment"

newdat_vln_mono_comb2$language_f <- as.factor(newdat_vln_mono_comb2$language)
newdat_vln_mono_comb2$speaker_f <- as.factor(newdat_vln_mono_comb2$speaker)


preds_vln_mono_comb2_baseline <- predict(vln_mod_mono_baseline, newdat_vln_mono_comb2,
  # the following is used when predicting without speakers
  exclude=c("s(phone_dur_uttr,speaker_f)",
            "s(phone_dur_uttr,language_f)",
            "s(phone_dur_uttr):vht_ononhigh"),
  se.fit = TRUE
)

preds_vln_mono_comb2_height <- predict(vln_mod_mono_height, newdat_vln_mono_comb2,
  # the following is used when predicting without speakers
  exclude=c("s(phone_dur_uttr,speaker_f)",
            "s(phone_dur_uttr,language_f)",
            "s(phone_dur_uttr):vht_ononhigh"),
  se.fit = TRUE
)

preds_vln_mono_comb2_pos <- predict(vln_mod_mono_pos, newdat_vln_mono_comb2,
  # the following is used when predicting without speakers
  exclude=c("s(phone_dur_uttr,speaker_f)",
            "s(phone_dur_uttr,language_f)",
            "s(phone_dur_uttr):vht_ononhigh"),
  se.fit = TRUE
)

preds_vln_mono_comb2_full <- predict(vln_mod_mono_full, newdat_vln_mono_comb2,
  # the following is used when predicting without speakers
  exclude=c("s(phone_dur_uttr,speaker_f)",
            "s(phone_dur_uttr,language_f)",
            "s(phone_dur_uttr):vht_ononhigh"),
  se.fit = TRUE
)

mods_vln_mono_nopos <- c("vln_mod_mono_baseline",
                         "vln_mod_mono_height")
mods_vln_mono_pos <- c("vln_mod_mono_pos",
                       "vln_mod_mono_full")
mods_vln_mono <- c(mods_vln_mono_nopos,mods_vln_mono_pos)

# baseline

newdat_vln_mono_comb2_baseline <- newdat_vln_mono_comb2

newdat_vln_mono_comb2_baseline$dur <- preds_vln_mono_comb2_baseline$fit
newdat_vln_mono_comb2_baseline$ul <- preds_vln_mono_comb2_baseline$fit + preds_vln_mono_comb2_baseline$se.fit * 1.96
newdat_vln_mono_comb2_baseline$ll <- preds_vln_mono_comb2_baseline$fit - preds_vln_mono_comb2_baseline$se.fit * 1.96

newdat_vln_mono_comb2_baseline$model <- "vln_mod_mono_baseline"

# vowel height

newdat_vln_mono_comb2_height <- newdat_vln_mono_comb2

newdat_vln_mono_comb2_height$dur <- preds_vln_mono_comb2_height$fit
newdat_vln_mono_comb2_height$ul <- preds_vln_mono_comb2_height$fit + preds_vln_mono_comb2_height$se.fit * 1.96
newdat_vln_mono_comb2_height$ll <- preds_vln_mono_comb2_height$fit - preds_vln_mono_comb2_height$se.fit * 1.96

newdat_vln_mono_comb2_height$model <- "vln_mod_mono_height"

# syllable position

newdat_vln_mono_comb2_pos <- newdat_vln_mono_comb2

newdat_vln_mono_comb2_pos$dur <- preds_vln_mono_comb2_pos$fit
newdat_vln_mono_comb2_pos$ul <- preds_vln_mono_comb2_pos$fit + preds_vln_mono_comb2_pos$se.fit * 1.96
newdat_vln_mono_comb2_pos$ll <- preds_vln_mono_comb2_pos$fit - preds_vln_mono_comb2_pos$se.fit * 1.96

newdat_vln_mono_comb2_pos$model <- "vln_mod_mono_pos"

# vowel height and syllable position

newdat_vln_mono_comb2_full <- newdat_vln_mono_comb2

newdat_vln_mono_comb2_full$dur <- preds_vln_mono_comb2_full$fit
newdat_vln_mono_comb2_full$ul <- preds_vln_mono_comb2_full$fit + preds_vln_mono_comb2_full$se.fit * 1.96
newdat_vln_mono_comb2_full$ll <- preds_vln_mono_comb2_full$fit - preds_vln_mono_comb2_full$se.fit * 1.96

newdat_vln_mono_comb2_full$model <- "vln_mod_mono_full"


newdat_vln_mono_comb2_all <- bind_rows(
  newdat_vln_mono_comb2_baseline,
  newdat_vln_mono_comb2_height,
  newdat_vln_mono_comb2_pos,
  newdat_vln_mono_comb2_full
)

newdat_vln_mono_comb2_all %>% write_csv("./newdat_vln_mono_comb2_all.csv")
newdat_vln_mono_comb2_all <- read_csv("./newdat_vln_mono_comb2_all.csv")
```


Plotting model predictions aggregated by contrast type.


```{r}
for (m in mods_vln_mono) {
  
  eff_sizes_vln_mono_comb_m <- eff_sizes_vln_mono_comb2[[m]]
  eff_sizes_vln_mono_comb_m$phone_dur_uttr <- 0.05
  eff_sizes_vln_mono_comb_m$value <- 0.13
  
  newdat_m <- filter(newdat_vln_mono_comb2_all, model==m)
  
  plot_dat <- newdat_m
  
  plot_dat %>%
  ggplot(data = ., aes(x = phone_dur_uttr)) +
  facet_wrap(~vowel_length_contrast,labeller=as_labeller(c("none"="No Contrast",
                                                    "restricted" = "Restricted Contrast\n(Only in Nonfinal Position)",
                                                    "unrestricted" = "Unrestricted Contrast"))) +
  # geom_point(data = d_uttr_vln_mono,
  #            aes(y = value, fill=measure), alpha = 0.1, pch = 21, stroke=NA) +
  geom_ribbon(data = filter(newdat_m,
                            vlcLn == "no_contrast",
                            syll_pos_o == "nonfinal"),
            aes(ymin = ll, ymax = ul), col = NA, fill = "magenta",
            alpha = 0.25) +
    geom_ribbon(data = filter(newdat_m,
                            vlcLn == "no_contrast",
                            syll_pos_o == "final"),
            aes(ymin = ll, ymax = ul), col = NA, fill = "magenta",
            alpha = 0.1) +
  geom_ribbon(data = filter(newdat_m,
                            vlcLn == "contrast_long",
                            syll_pos_o == "nonfinal"),
            aes(ymin = ll, ymax = ul), col = NA, fill = "blue",
            alpha = 0.25) +
  geom_ribbon(data = filter(newdat_m,
                            vlcLn == "contrast_long",
                            syll_pos_o == "final"),
            aes(ymin = ll, ymax = ul), col = NA, fill = "blue",
            alpha = 0.1) +
  geom_ribbon(data = filter(newdat_m,
                            vlcLn == "contrast_short",
                            syll_pos_o == "nonfinal"),
            aes(ymin = ll, ymax = ul), col = NA, fill = "limegreen",
            alpha = 0.25) +
  geom_ribbon(data = filter(newdat_m,
                            vlcLn == "contrast_short",
                            syll_pos_o == "final"),
            aes(ymin = ll, ymax = ul), col = NA, fill = "limegreen",
            alpha = 0.1) +
  geom_segment(aes(y=min(d_vl$phone_dur_uttr),
                   yend=max(d_vl$phone_dur_uttr),
                   x=min(d_vl$phone_dur_uttr),
                   xend=max(d_vl$phone_dur_uttr)),
               col = "black",
               lwd = 0.5,
               linetype="dotted"
               ) +
  geom_line(aes(y=dur, col = vlcLn_o, lty = syll_pos_o)) +
  # geom_text(data=eff_sizes_vln_mono_comb_m, 
  #           aes(y=value*1.2*1.2*1.2,label=paste0("β_nc = ", round(derivative_nc,2))),
  #           col="black", size=5) +
  #   geom_text(data=eff_sizes_vln_mono_comb_m, 
  #           aes(y=value*1.2*1.2,label=paste0("β_l = ", round(derivative_cl,2))),
  #           col="black", size=5) +
  #   geom_text(data=eff_sizes_vln_mono_comb_m, 
  #           aes(y=value*1.2,label=paste0("β_s = ", round(derivative_cs,2))),
  #           col="black", size=5) +
  #   geom_text(data=eff_sizes_vln_mono_comb_m, 
  #           aes(y=value,label=paste0("Δβ = ", round(eff_size,2))),
  #           col="black", size=5) +
  scale_color_manual(values = c("no_contrast" = "magenta",
                                "contrast_long" = "blue",
                                "contrast_short" = "limegreen"),
                     labels = c("no_contrast" = "no contrast",
                               "contrast_long" = "long",
                               "contrast_short" = "short"),
                     guide = "legend") +
  scale_fill_manual(values = c("nocontrast_dur_uttr" = "hotpink", 
                                "long_dur_uttr" = "cyan", 
                                "short_dur_uttr" = "green"),
                     guide = "none") +
  scale_linetype_manual(values = c("nonfinal" = "solid",
                                    "final" = "dashed")) +
  scale_x_log10(breaks = c(0.03, 0.05, 0.07, 0.1, 0.14, 0.2)) +
  scale_y_log10(breaks = c(0.03, 0.05, 0.07, 0.1, 0.14, 0.2)) +
  coord_fixed() +
  #scale_x_continuous(limits=c(0.035,0.145), 
  #                   breaks=seq(0.04,0.14,0.02),
  #                   labels=c(".04",".06",".08",".10",".12",".14")) +
  #scale_y_continuous(breaks=seq(0.04,0.20,0.02)) +
  labs(title = "Duration-Rate Curves by Type of Vowel Length Contrast",
       x = "Average phone duration by utterance (inverse rate) (s)",
       y = "Phone duration (s)") +
  guides(col = guide_legend(title = "Vowel Length"),
         lty = guide_legend(title = "Position in Word")) +
  #ylab("Average V vs. C duration in s") +
  # remove clutter
  theme_minimal() +
  theme(plot.title = element_text(size=24, face="bold"),
        panel.grid = element_blank(),
        axis.line = element_line(size=0.8),
        axis.ticks = element_line(size=0.8),
        axis.text = element_text(size=16, colour="black"),
        axis.title = element_text(size=18, face="bold"),
        strip.background.y = element_blank(), 
        strip.placement = "outside",
        strip.text=element_text(size=18, face="bold"))
  ggsave(paste0("doreco_graphs_v1/vl/comb/", m, "_comb2.png"), width=12, height=9, dpi=300)
}
```

# VL Model Effect Sizes at Medians

Measuring and comparing derivatives at the medians for VL models

```{r}
eff_sizes_vln_mono <- list()   
eff_sizes_vln_mono_nopos <- list()       

# model predictions
newdat_vln_mono_deriv <- expand.grid(
  vht_o = d_vl$vht_o[1],
  speaker = d_vl$speaker[1],
  syll_pos = unique(d_vl$syll_pos),
  language = unique(d_vl$language),
  vlcLn = unique(d_vl$vlcLn)
) %>%
  mutate(vowel_length_contrast = case_when(
           language %in% languages_no_contrast ~ "none",
           language %in% languages_restricted ~ "restricted",
           language %in% languages_unrestricted ~ "unrestricted")) %>%
  filter(vowel_length_contrast == "none" & vlcLn == "no_contrast" |
           (vowel_length_contrast %in% c("unrestricted") &
             vlcLn %in% c("contrast_long","contrast_short")) |
           ((vowel_length_contrast %in% c("restricted") &
             ((vlcLn %in% c("contrast_long","contrast_short") & syll_pos == "nonfinal") |
                (vlcLn %in% c("no_contrast") & syll_pos == "final"))))
         )

newdat_vln_mono_deriv$vlcLn_o <- as.ordered(newdat_vln_mono_deriv$vlcLn)
contrasts(newdat_vln_mono_deriv$vlcLn_o) <- "contr.treatment"

newdat_vln_mono_deriv$syll_pos_o <- as.ordered(newdat_vln_mono_deriv$syll_pos)
contrasts(newdat_vln_mono_deriv$syll_pos_o) <- "contr.treatment"

newdat_vln_mono_deriv$language_f <- as.factor(newdat_vln_mono_deriv$language)
newdat_vln_mono_deriv$speaker_f <- as.factor(newdat_vln_mono_deriv$speaker)

language_med_durs <- d_all %>%
  group_by(language) %>%
  summarise(phone_dur_uttr = median(phone_dur_uttr)) %>%
  ungroup()

newdat_vln_mono_deriv <- left_join(
  newdat_vln_mono_deriv,
  language_med_durs,
  by = "language"
)

eps <- 0.001

newdat_vln_mono_deriv_1 <- newdat_vln_mono_deriv %>%
  mutate(phone_dur_uttr = exp(log(phone_dur_uttr) - (eps/2)))
newdat_vln_mono_deriv_2 <- newdat_vln_mono_deriv %>%
  mutate(phone_dur_uttr = exp(log(phone_dur_uttr) + (eps/2)))

mods_vln_mono_nopos <- c("vln_mod_mono_baseline",
                         "vln_mod_mono_height")
mods_vln_mono_pos <- c("vln_mod_mono_pos",
                       "vln_mod_mono_full")
mods_vln_mono <- c(mods_vln_mono_nopos,mods_vln_mono_pos)

for (m in mods_vln_mono) {
X1 <- predict(get(m), 
              newdat_vln_mono_deriv_1,
              exclude=c(
                "s(phone_dur_uttr,speakerSegtype)",
                "s(phone_dur_uttr,speaker_f)",
                "s(phone_dur_uttr,speaker_f):segtype_ovowel_dur_uttr",
                "s(phone_dur_uttr,vht_o)"
              )
) %>% log()
X2 <- predict(get(m), 
              newdat_vln_mono_deriv_2,
              exclude=c(
                "s(phone_dur_uttr,speakerSegtype)",
                "s(phone_dur_uttr,speaker_f)",
                "s(phone_dur_uttr,speaker_f):segtype_ovowel_dur_uttr",
                "s(phone_dur_uttr,vht_o)"
              )
) %>% log()

X0 <- predict(get(m),
              newdat_vln_mono_deriv,
              exclude=c(
                "s(phone_dur_uttr,speakerSegtype)",
                "s(phone_dur_uttr,speaker_f)",
                "s(phone_dur_uttr,speaker_f):segtype_ovowel_dur_uttr",
                "s(phone_dur_uttr,vht_o)"
              )
)

# finite difference approximation of first derivative
# the design matrix
first_deriv <- (X2 - X1) / eps
#first_deriv <- Xp %*% coef(get(m))

newdat_vln_mono_deriv$derivative <- first_deriv
newdat_vln_mono_deriv$model <- m

newdat_vln_mono_deriv$phone_dur <- X0

eff_sizes_vln_mono_contrast <- newdat_vln_mono_deriv %>%
  dplyr::select(vowel_length_contrast, language, vlcLn, syll_pos,
                phone_dur_uttr, model, phone_dur, derivative) %>%
  filter(vlcLn != "no_contrast") %>%
  group_by(vowel_length_contrast, language, phone_dur_uttr, model, syll_pos) %>%
  summarise(derivative_cs = derivative[vlcLn == "contrast_short"],
            derivative_cl = derivative[vlcLn == "contrast_long"],
            eff_size = derivative_cl - derivative_cs,
            phone_dur_cs = phone_dur[vlcLn == "contrast_short"],
            phone_dur_cl = phone_dur[vlcLn == "contrast_long"]) %>%
  ungroup() %>%
  arrange(desc(eff_size))

derivs_vln_mono_nocontrast <- newdat_vln_mono_deriv %>%
  dplyr::select(vowel_length_contrast, language, vlcLn, syll_pos,
                phone_dur_uttr, model, phone_dur, derivative) %>%
  filter(vowel_length_contrast != "unrestricted") %>%
  group_by(vowel_length_contrast, language, phone_dur_uttr, model, syll_pos) %>%
  summarise(derivative_nc = derivative[vlcLn == "no_contrast"],
            phone_dur_nc = phone_dur[vlcLn == "no_contrast"]) %>%
  ungroup() %>%
  arrange(desc(derivative_nc))

eff_sizes_vln_mono[[m]] <- full_join(eff_sizes_vln_mono_contrast,
                                     derivs_vln_mono_nocontrast,
                                     by=c("vowel_length_contrast",
                                          "language",
                                          "phone_dur_uttr",
                                          "model",
                                          "syll_pos"))
if (m %in% mods_vln_mono_nopos){
  eff_sizes_vln_mono_nopos[[m]] <- eff_sizes_vln_mono[[m]] %>%
  group_by(across(-c(syll_pos,
                   eff_size,
                   derivative_cs,derivative_cl,derivative_nc,
                   phone_dur_cs,phone_dur_cl,phone_dur_nc))) %>%
  summarize(eff_size = eff_size[1],
            derivative_cs = derivative_cs[1],
            derivative_cl = derivative_cl[1],
            derivative_nc = derivative_nc[1],
            phone_dur_cs = phone_dur_cs[1],
            phone_dur_cl = phone_dur_cl[1],
            phone_dur_nc = phone_dur_nc[1],
            )
  }
}

eff_sizes_vln_mono_df <- bind_rows(eff_sizes_vln_mono)
eff_sizes_vln_mono_df %>% write_csv("eff_sizes_vln_mono_df.csv")
eff_sizes_vln_mono_df <- read_csv("./eff_sizes_vln_mono_df.csv")

eff_sizes_vln_mono_nopos_df <- bind_rows(eff_sizes_vln_mono_nopos)
eff_sizes_vln_mono_nopos_df %>% write_csv("eff_sizes_vln_mono_nopos_df.csv")
eff_sizes_vln_mono_nopos_df <- read_csv("./eff_sizes_vln_mono_nopos_df.csv")
```
## Making VL Prediction Plots Showing Effect Sizes at Medians

Ordering by effect size at medians & showing this number in the plot of VL model predictions.

```{r}

for (m in mods_vln_mono) {
  eff_sizes_vln_mono_m <- eff_sizes_vln_mono[[m]]
  eff_sizes_vln_mono_m$language <- factor(eff_sizes_vln_mono_m$language,
                                          levels=unique(eff_sizes_vln_mono_m$language))
  eff_sizes_vln_mono_m$phone_dur_uttr_median <- eff_sizes_vln_mono_m$phone_dur_uttr
  eff_sizes_vln_mono_m$phone_dur_uttr <- 0.05
  eff_sizes_vln_mono_m$value <- 0.13
  
  newdat_m <- filter(newdat_vln_mono_all, model==m)
  newdat_m$language <- factor(newdat_m$language, levels=unique(eff_sizes_vln_mono_m$language))
  
  plot_dat <- newdat_m
  plot_dat$language <- factor(plot_dat$language, levels=unique(eff_sizes_vln_mono_m$language))
  
  d_uttr_vln_mono$language <- factor(d_uttr_vln_mono$language, levels=unique(eff_sizes_vln_mono_m$language))
  
  plot_dat %>%
  #filter(vowel_length_contrast != "none") %>%
  ggplot(data = ., aes(x = phone_dur_uttr, y=value)) +
  facet_wrap(. ~ language) +
  geom_point(data = d_uttr_vln_mono,
             aes(y = value, col=measure), alpha = 0.3, pch=16) +
  geom_line(aes(y = dur, col = vlcLn_o, lwd = syll_pos_o)) +
  # geom_ribbon(data = filter(newdat_cv_all,
  #                          model == "cv_mod_baseline",
  #                          segtype_o == "consonant"),
  #           aes(ymin = ll, ymax = ul), col = NA, fill = "skyblue",
  #           alpha = 0.5) +
  # geom_ribbon(data = filter(newdat_cv_all,
  #                          model == "cv_mod_baseline",
  #                          segtype_o == "syllabic"),
  #           aes(ymin = ll, ymax = ul), col = NA, fill = "yellow",
  #           alpha = 0.5) +
  geom_text(data=filter(eff_sizes_vln_mono_m,!is.na(derivative_cs),syll_pos=='nonfinal'), 
            aes(y=value,label=paste0("Δβ = ", round(eff_size,2))),
            col="black", size=5) +
    geom_text(data=filter(eff_sizes_vln_mono_m,!is.na(derivative_cs),syll_pos=='nonfinal'), 
            aes(y=value*1.4,label=paste0("β_s = ", round(derivative_cs,2))),
            col="black", size=5) +
    geom_text(data=filter(eff_sizes_vln_mono_m,!is.na(derivative_cl),syll_pos=='nonfinal'), 
            aes(y=value*1.4*1.4,label=paste0("β_l = ", round(derivative_cl,2))),
            col="black", size=5) +
    geom_text(data=filter(eff_sizes_vln_mono_m,!is.na(derivative_nc),syll_pos=='nonfinal'), 
            aes(y=value*1.4*1.4*1.4,label=paste0("β = ", round(derivative_nc,2))),
            col="black", size=5) +
  geom_text(data=filter(eff_sizes_vln_mono_m,!is.na(derivative_cs),syll_pos=='final'), 
            aes(x=phone_dur_uttr*2,
                y=value,
                label=paste0("Δβ = ", round(eff_size,2))),
            col="black", size=5) +
    geom_text(data=filter(eff_sizes_vln_mono_m,!is.na(derivative_cs),syll_pos=='final'), 
            aes(x=phone_dur_uttr*2,
                y=value*1.4,
                label=paste0("β_s = ", round(derivative_cs,2))),
            col="black", size=5) +
    geom_text(data=filter(eff_sizes_vln_mono_m,!is.na(derivative_cl),syll_pos=='final'), 
            aes(x=phone_dur_uttr*2,
                y=value*1.4*1.4,
                label=paste0("β_l = ", round(derivative_cl,2))),
            col="black", size=5) +
    geom_text(data=filter(eff_sizes_vln_mono_m,!is.na(derivative_nc),syll_pos=='final'), 
            aes(x=phone_dur_uttr*2,
                y=value*1.4*1.4*1.4,
                label=paste0("β = ", round(derivative_nc,2))),
            col="black", size=5) +
  geom_text(data=eff_sizes_vln_mono_m,
            aes(x=phone_dur_uttr*1.4,
                y=value/1.4/1.4/1.4,label=paste0("med = ", round(phone_dur_uttr_median,2))),
            col="black", size=5) +
  scale_color_manual(values = c("no_contrast" = "magenta",
                                "contrast_long" = "blue",
                                "contrast_short" = "limegreen",
                                "nocontrast_dur_uttr" = "hotpink", 
                                "long_dur_uttr" = "cyan", 
                                "short_dur_uttr" = "green"),
                     guide = "none") +
  # scale_linetype_manual(values = c("high" = "dashed",
  #                                  "nonhigh" = "solid"
  #                                  )) +
  scale_linewidth_manual(values = c("final" = 1,
                                    "nonfinal" = 0.5)) +
  scale_x_log10(breaks = c(0.05, 0.07, 0.1, 0.14)) +
  scale_y_log10(breaks = c(0.05, 0.07, 0.1, 0.14),
                #limits = c(0.03, 0.2)
                ) +
  #scale_x_continuous(limits=c(0.035,0.145), 
  #                   breaks=seq(0.04,0.14,0.02),
  #                   labels=c(".04",".06",".08",".10",".12",".14")) +
  #scale_y_continuous(breaks=seq(0.04,0.20,0.02)) +
    # labels for axes
  xlab("Average phone duration by utterance") +
  labs(y = "Average long vs short or no-contrast monophthong duration in s") +
  #ylab("Average V vs. C duration in s") +
  # remove clutter
  theme_minimal() +
  theme(panel.grid = element_blank(),
        axis.line = element_line(size=0.8),
        axis.ticks = element_line(size=0.8),
        axis.text = element_text(size=16, colour="black"),
        axis.title = element_text(size=18, face="bold"),
        strip.background.y = element_blank(), 
        strip.placement = "outside",
        strip.text=element_text(size=18, face="bold"))
  ggsave(paste0("doreco_graphs_v1/vl/", m, ".png"), width=15, height=15, dpi=300)
}


```

## Comparing CV Model Effect Sizes at Medians

Comparing effect sizes / slopes and medians for VL models

```{r}
for (m in mods_vln_mono) {
  eff_sizes_vln_mono_m <- eff_sizes_vln_mono[[m]] %>%
    filter(vowel_length_contrast != "none") %>%
  ggplot(aes(x=phone_dur_uttr,y=eff_size)) +
    geom_point(aes(col=vowel_length_contrast,shape=syll_pos)) +
    scale_color_manual(values = c("unrestricted" = "black",
                                "restricted" = "red")) +
    xlab("Median average phone duration by utterance") +
    labs(y = "Δβ ") +
    theme_minimal() +
    theme(panel.grid = element_blank(),
        axis.line = element_line(size=0.8),
        axis.ticks = element_line(size=0.8),
        axis.text = element_text(size=16, colour="black"),
        axis.title = element_text(size=18, face="bold"),
        strip.background.y = element_blank(), 
        strip.placement = "outside",
        strip.text=element_text(size=18, face="bold"))
  ggsave(paste0("doreco_graphs_v1/vl/vl_eff_size_comp_contrast_", m, ".png"), width=8, height=5, dpi=300)
  
  
  
  eff_sizes_vln_mono_m <- eff_sizes_vln_mono[[m]] %>%
  ggplot(aes(x=phone_dur_uttr)) +
    geom_point(aes(y=derivative_nc,shape=syll_pos),col="magenta") +
    geom_point(aes(y=derivative_cl,shape=syll_pos),col="blue") +
    geom_point(aes(y=derivative_cs,shape=syll_pos),col="limegreen") +
    geom_segment(aes(y = derivative_cs, yend=derivative_cl,
                     col = vowel_length_contrast)) +
    geom_segment(aes(y = derivative_nc, yend=derivative_cs,
                     col = vowel_length_contrast),
                 linetype = "dashed") +
    geom_segment(aes(y = derivative_nc, yend=derivative_cl,
                     col = vowel_length_contrast),
                 linetype = "dashed") +
    scale_color_manual(values = c("unrestricted" = "black",
                                "restricted" = "red")) +
    xlab("Median average phone duration by utterance") +
    labs(y = "β") +
    theme_minimal() +
    theme(panel.grid = element_blank(),
        axis.line = element_line(size=0.8),
        axis.ticks = element_line(size=0.8),
        axis.text = element_text(size=16, colour="black"),
        axis.title = element_text(size=18, face="bold"),
        strip.background.y = element_blank(), 
        strip.placement = "outside",
        strip.text=element_text(size=18, face="bold"))
  ggsave(paste0("doreco_graphs_v1/vl/vl_deriv_comp_all_", m, ".png"), width=8, height=5, dpi=300)
}
```


Comparing predicted phone durations and slopes for VL models

```{r}
for (m in mods_vln_mono) {
  eff_sizes_vln_mono_m <- eff_sizes_vln_mono[[m]] %>%
  ggplot(aes(x=phone_dur_uttr)) +
    geom_point(aes(y=phone_dur_nc),col="magenta") +
    geom_point(aes(y=phone_dur_cl),col="blue") +
    geom_point(aes(y=phone_dur_cs),col="limegreen") +
    geom_segment(aes(y = phone_dur_cs, yend=phone_dur_cl,
                     col = vowel_length_contrast)) +
    geom_segment(aes(y = phone_dur_nc, yend=phone_dur_cs,
                     col = vowel_length_contrast),
                 linetype = "dashed") +
    geom_segment(aes(y = phone_dur_nc, yend=phone_dur_cl,
                     col = vowel_length_contrast),
                 linetype = "dashed") +
    scale_color_manual(values = c("unrestricted" = "black",
                                "restricted" = "red")) +
    labs(x = "Median average phone duration by utterance",
         y = "Predicted individual phone duration at median") +
    theme_minimal() +
    theme(panel.grid = element_blank(),
        axis.line = element_line(size=0.8),
        axis.ticks = element_line(size=0.8),
        axis.text = element_text(size=16, colour="black"),
        axis.title = element_text(size=18, face="bold"),
        strip.background.y = element_blank(), 
        strip.placement = "outside",
        strip.text=element_text(size=18, face="bold"))
  ggsave(paste0("doreco_graphs_v1/vl/vl_medians_rate_vs_dur_all_", m, ".png"), width=8, height=5, dpi=300)
  
  eff_sizes_vln_mono_m <- eff_sizes_vln_mono[[m]] %>%
  ggplot() +
    geom_point(aes(x=phone_dur_nc,y=derivative_nc),col="magenta") +
    geom_point(aes(x=phone_dur_cl,y=derivative_cl),col="blue") +
    geom_point(aes(x=phone_dur_cs,y=derivative_cs),col="limegreen") +
    geom_segment(aes(x = phone_dur_cs, xend=phone_dur_cl,
                     y = derivative_cs, yend=derivative_cl,
                     col = vowel_length_contrast)) +
    geom_segment(aes(x = phone_dur_nc, xend=phone_dur_cs,
                     y = derivative_nc, yend=derivative_cs,
                     col = vowel_length_contrast),
                 linetype = "dashed") +
    geom_segment(aes(x = phone_dur_nc, xend=phone_dur_cl,
                     y = derivative_nc, yend=derivative_cl,
                     col = vowel_length_contrast),
                 linetype = "dashed") +
    scale_color_manual(values = c("unrestricted" = "black",
                                "restricted" = "red")) +
    xlab("Predicted individual phone duration at median") +
    labs(y = "β") +
    theme_minimal() +
    theme(panel.grid = element_blank(),
        axis.line = element_line(size=0.8),
        axis.ticks = element_line(size=0.8),
        axis.text = element_text(size=16, colour="black"),
        axis.title = element_text(size=18, face="bold"),
        strip.background.y = element_blank(), 
        strip.placement = "outside",
        strip.text=element_text(size=18, face="bold"))
  ggsave(paste0("doreco_graphs_v1/vl/vl_deriv_preds_all_", m, ".png"), width=8, height=5, dpi=300)
}
```


Comparing effect sizes and medians for VL models

```{r}
for (m in mods_vln_mono) {
  eff_sizes_vln_mono_pos_m <- eff_sizes_vln_mono[[m]]
  
  eff_sizes_vln_mono_pos_m_wider <- eff_sizes_vln_mono_pos_m %>%
  pivot_wider(names_from = syll_pos,
              values_from = c(eff_size,
                              derivative_nc,
                              derivative_cs,
                              derivative_cl,
                              phone_dur_nc,
                              phone_dur_cs,
                              phone_dur_cl))
  
  eff_sizes_vln_mono_pos_m %>%
    filter(vowel_length_contrast != "none") %>%
  ggplot(aes(x=phone_dur_uttr)) +
    geom_point(aes(y=eff_size,col=vowel_length_contrast,shape=syll_pos)) +
    geom_segment(data = eff_sizes_vln_mono_pos_m_wider,
                 aes(y=eff_size_final,
                     yend=eff_size_nonfinal,
                     col = vowel_length_contrast),
                 linetype="dashed") +
    scale_color_manual(values = c("unrestricted" = "black",
                                "restricted" = "red")) +
    scale_x_log10(breaks=c(0.05,0.06,0.07,0.08,0.09,0.1,0.11,0.125,0.15)) +
    xlab("Median average phone duration by utterance") +
    labs(y = "Δβ ") +
    theme_minimal() +
    theme(panel.grid = element_blank(),
        axis.line = element_line(size=0.8),
        axis.ticks = element_line(size=0.8),
        axis.text = element_text(size=16, colour="black"),
        axis.title = element_text(size=18, face="bold"),
        strip.background.y = element_blank(), 
        strip.placement = "outside",
        strip.text=element_text(size=18, face="bold"))
  ggsave(paste0("doreco_graphs_v1/vl/vl_eff_size_comp_contrast_", m, ".png"), width=8, height=5, dpi=300)

  eff_sizes_vln_mono_pos_m %>%
  ggplot(aes(x=phone_dur_uttr)) +
    geom_point(aes(y=derivative_nc,shape=syll_pos),col="magenta") +
    geom_point(aes(y=derivative_cl,shape=syll_pos),col="blue") +
    geom_point(aes(y=derivative_cs,shape=syll_pos),col="limegreen") +
    geom_segment(aes(y = derivative_cs, yend=derivative_cl,
                     col = vowel_length_contrast)) +
    # geom_segment(aes(y = derivative_nc, yend=derivative_cs,
    #                  col = vowel_length_contrast),
    #              linetype = "dashed") +
    # geom_segment(aes(y = derivative_nc, yend=derivative_cl,
    #                  col = vowel_length_contrast),
    #              linetype = "dashed") +
    geom_segment(data = eff_sizes_vln_mono_pos_m_wider,
                 aes(y=derivative_nc_final,
                     yend=derivative_nc_nonfinal,
                     col = vowel_length_contrast),
                 linetype="dashed") +
    scale_color_manual(values = c("unrestricted" = "black",
                                "restricted" = "red")) +
    scale_x_log10(breaks=c(0.05,0.06,0.07,0.08,0.09,0.1,0.11,0.125,0.15)) +
    xlab("Median average phone duration by utterance") +
    labs(y = "β") +
    theme_minimal() +
    theme(panel.grid = element_blank(),
        axis.line = element_line(size=0.8),
        axis.ticks = element_line(size=0.8),
        axis.text = element_text(size=16, colour="black"),
        axis.title = element_text(size=18, face="bold"),
        strip.background.y = element_blank(), 
        strip.placement = "outside",
        strip.text=element_text(size=18, face="bold"))
  ggsave(paste0("doreco_graphs_v1/vl/vl_deriv_comp_all_", m, ".png"), width=8, height=5, dpi=300)
}
```

```{r}
for (m in mods_vln_mono) {
  eff_sizes_vln_mono_pos_m <- eff_sizes_vln_mono[[m]]
  
  eff_sizes_vln_mono_pos_m_wider <- eff_sizes_vln_mono_pos_m %>%
  pivot_wider(names_from = syll_pos,
              values_from = c(eff_size,
                              derivative_nc,
                              derivative_cs,
                              derivative_cl,
                              phone_dur_nc,
                              phone_dur_cs,
                              phone_dur_cl))
  
  eff_sizes_vln_mono_pos_m %>%
  ggplot(aes(x=phone_dur_uttr)) +
    geom_point(aes(y=phone_dur_nc,shape=syll_pos),col="magenta") +
    geom_point(aes(y=phone_dur_cl,shape=syll_pos),col="blue") +
    geom_point(aes(y=phone_dur_cs,shape=syll_pos),col="limegreen") +
    geom_segment(aes(y = phone_dur_cs, yend=phone_dur_cl,
                     col = vowel_length_contrast)) +
    geom_segment(aes(y = phone_dur_nc, yend=phone_dur_cs,
                     col = vowel_length_contrast),
                 linetype = "dashed") +
    geom_segment(aes(y = phone_dur_nc, yend=phone_dur_cl,
                     col = vowel_length_contrast),
                 linetype = "dashed") +
    scale_color_manual(values = c("unrestricted" = "black",
                                "restricted" = "red")) +
    scale_x_log10(breaks=c(0.067,0.1,0.15)) +
    scale_y_log10(breaks=c(0.067,0.1,0.15)) +
    labs(x = "Median average phone duration by utterance",
         y = "Predicted individual phone duration at median") +
    theme_minimal() +
    theme(panel.grid = element_blank(),
        axis.line = element_line(size=0.8),
        axis.ticks = element_line(size=0.8),
        axis.text = element_text(size=16, colour="black"),
        axis.title = element_text(size=18, face="bold"),
        strip.background.y = element_blank(), 
        strip.placement = "outside",
        strip.text=element_text(size=18, face="bold"))
  ggsave(paste0("doreco_graphs_v1/vl/vl_medians_rate_vs_dur_all_", m, ".png"), width=8, height=5, dpi=300)
  
  eff_sizes_vln_mono_pos_m %>%
  ggplot() +
    facet_wrap(~vowel_length_contrast) +
    geom_point(aes(x=phone_dur_nc,y=derivative_nc,shape=syll_pos),col="magenta") +
    geom_point(aes(x=phone_dur_cl,y=derivative_cl,shape=syll_pos),col="blue") +
    geom_point(aes(x=phone_dur_cs,y=derivative_cs,shape=syll_pos),col="limegreen") +
    geom_segment(aes(x = phone_dur_cs, xend=phone_dur_cl,
                     y = derivative_cs, yend=derivative_cl,
                     col = vowel_length_contrast)) +
    geom_segment(data = eff_sizes_vln_mono_pos_m_wider,
                   aes(x = phone_dur_nc_final, xend=phone_dur_cs_nonfinal,
                     y = derivative_nc_final, yend=derivative_cs_nonfinal,
                     col = vowel_length_contrast),
                 linetype="dashed") +
    geom_segment(data = eff_sizes_vln_mono_pos_m_wider,
                 aes(x = phone_dur_nc_final, xend=phone_dur_cl_nonfinal,
                     y = derivative_nc_final, yend=derivative_cl_nonfinal,
                     col = vowel_length_contrast),
                 linetype="dashed") +
    geom_segment(data = eff_sizes_vln_mono_pos_m_wider,
                   aes(x = phone_dur_nc_final, xend=phone_dur_nc_nonfinal,
                     y = derivative_nc_final, yend=derivative_nc_nonfinal),
                 linetype="dashed") +
    geom_segment(data = eff_sizes_vln_mono_pos_m_wider,
                   aes(x = phone_dur_cl_final, xend=phone_dur_cl_nonfinal,
                     y = derivative_cl_final, yend=derivative_cl_nonfinal),
                 linetype="dashed") +
    geom_segment(data = eff_sizes_vln_mono_pos_m_wider,
                   aes(x = phone_dur_cs_final, xend=phone_dur_cs_nonfinal,
                     y = derivative_cs_final, yend=derivative_cs_nonfinal),
                 linetype="dashed") +
    scale_color_manual(values = c("unrestricted" = "black",
                                "restricted" = "red")) +
    scale_x_log10(breaks=c(0.067,0.1,0.15)) +
    xlab("Predicted individual phone duration at median") +
    labs(y = "β") +
    theme_minimal() +
    theme(panel.grid = element_blank(),
        axis.line = element_line(size=0.8),
        axis.ticks = element_line(size=0.8),
        axis.text = element_text(size=16, colour="black"),
        axis.title = element_text(size=18, face="bold"),
        strip.background.y = element_blank(), 
        strip.placement = "outside",
        strip.text=element_text(size=18, face="bold"))
  ggsave(paste0("doreco_graphs_v1/vl/vl_deriv_preds_all_", m, ".png"), width=12, height=5, dpi=300)
}
```









# VL Model Density-Weighted Effect Sizes

## Computing VL Model Effect Curves and Density-Weighted Effect Sizes

```{r}
eff_curves_vln_mono_contrast <- list()
deriv_curves_vln_mono_nocontrast <- list()

eff_curves_vln_mono <- list()
eff_sizes_dw_vln_mono <- list()

newdat_vl_deriv_curve <- expand.grid(
  phone_dur_uttr = seq(
    min(d_vl$phone_dur_uttr),
    max(d_vl$phone_dur_uttr),
    length.out = 100
  ),
  vht = d_vl$segtype_vowel_height[1], #unique(d_vl$segtype_vowel_height)
  syll_pos = unique(d_vl$syll_pos),
  speaker = d_vl$speaker[1],
  language = unique(d_vl$language),
  vlcLn = unique(d_vl$vlcLn)
) %>%
  mutate(vowel_length_contrast = case_when(
           language %in% languages_no_contrast ~ "none",
           language %in% languages_restricted ~ "restricted",
           language %in% languages_unrestricted ~ "unrestricted")) %>%
  filter(vowel_length_contrast == "none" & vlcLn == "no_contrast" |
           (vowel_length_contrast %in% c("unrestricted") &
             vlcLn %in% c("contrast_long","contrast_short")) |
           ((vowel_length_contrast %in% c("restricted") &
             ((vlcLn %in% c("contrast_long","contrast_short") & syll_pos == "nonfinal") |
                (vlcLn %in% c("no_contrast") & syll_pos == "final"))))
         )

newdat_vl_deriv_curve$vlcLn_o <- as.ordered(newdat_vl_deriv_curve$vlcLn)
contrasts(newdat_vl_deriv_curve$vlcLn_o) <- "contr.treatment"

newdat_vl_deriv_curve$vht_o <- as.ordered(newdat_vl_deriv_curve$vht)
#contrasts(newdat_vl_deriv_curve$vht_o) <- "contr.treatment"

newdat_vl_deriv_curve$syll_pos_o <- as.ordered(newdat_vl_deriv_curve$syll_pos)
contrasts(newdat_vl_deriv_curve$syll_pos_o) <- "contr.treatment"

newdat_vl_deriv_curve$language_f <- as.factor(newdat_vl_deriv_curve$language)
newdat_vl_deriv_curve$speaker_f <- as.factor(newdat_vl_deriv_curve$speaker)


language_dist_stats <- d_vl %>%
  select(language,phone_dur_uttr) %>%
  group_by(language) %>%
  nest() %>%
  mutate(
    phone_dur_uttr_med = as.numeric(map(data,~median(.x$phone_dur_uttr))),
    dens_data = map(data,~broom::tidy(density(.x$phone_dur_uttr,
                                              n = 100,
                                              from = min(d_vl$phone_dur_uttr),
                                              to = max(d_vl$phone_dur_uttr))))
         ) %>%
  unnest(dens_data) %>%
  mutate(phone_dur_uttr = x,
         density = y/max(y)) %>%
  ungroup() %>%
  select(language,phone_dur_uttr_med,phone_dur_uttr,density)
  

newdat_vl_deriv_curve <- left_join(
  mutate(newdat_vl_deriv_curve,
         phone_dur_uttr = round(phone_dur_uttr,15)),
  mutate(language_dist_stats,
         phone_dur_uttr = round(phone_dur_uttr,15)),
  by = c("language","phone_dur_uttr")
)


eps <- 0.001

newdat_vl_deriv_curve_1 <- newdat_vl_deriv_curve %>%
  mutate(phone_dur_uttr = exp(log(phone_dur_uttr) - (eps/2)))
newdat_vl_deriv_curve_2 <- newdat_vl_deriv_curve %>%
  mutate(phone_dur_uttr = exp(log(phone_dur_uttr) + (eps/2)))

mods_cv <- c("cv_mod_baseline", "cv_mod_type"
             #, "cv_mod_type2"
             )

for (m in mods_vln_mono) {
X1_curve <- predict(get(m), 
              newdat_vl_deriv_curve_1,
              exclude=c(
                "s(phone_dur_uttr,speakerSegtype)",
                "s(phone_dur_uttr,speaker_f)",
                "s(phone_dur_uttr,speaker_f):segtype_ovowel_dur_uttr",
                "s(phone_dur_uttr,phone_type_f)"
                #"s(phone_dur_uttr,phone_type_new_f)"
              )
) %>% log()
X2_curve <- predict(get(m), 
              newdat_vl_deriv_curve_2,
              exclude=c(
                "s(phone_dur_uttr,speakerSegtype)",
                "s(phone_dur_uttr,speaker_f)",
                "s(phone_dur_uttr,speaker_f):segtype_ovowel_dur_uttr",
                "s(phone_dur_uttr,phone_type_f)"
                #"s(phone_dur_uttr,phone_type_new_f)"
              )
) %>% log()

X0_curve <- predict(get(m), 
              newdat_vl_deriv_curve,
              exclude=c(
                "s(phone_dur_uttr,speakerSegtype)",
                "s(phone_dur_uttr,speaker_f)",
                "s(phone_dur_uttr,speaker_f):segtype_ovowel_dur_uttr",
                "s(phone_dur_uttr,phone_type_f)"
                #"s(phone_dur_uttr,phone_type_new_f)"
              )
)

# finite difference approximation of first derivative
# the design matrix
first_deriv_curve <- (X2_curve - X1_curve) / eps
#first_deriv <- Xp %*% coef(get(m))



newdat_vl_deriv_curve$derivative <- first_deriv_curve
newdat_vl_deriv_curve$model <- m

newdat_vl_deriv_curve$phone_dur <- X0_curve






eff_sizes_vln_mono_contrast <- newdat_vln_mono_deriv %>%
  dplyr::select(vowel_length_contrast, language, vlcLn, syll_pos,
                phone_dur_uttr, model, phone_dur, derivative) %>%
  filter(vlcLn != "no_contrast") %>%
  group_by(vowel_length_contrast, language, phone_dur_uttr, model, syll_pos) %>%
  summarise(derivative_cs = derivative[vlcLn == "contrast_short"],
            derivative_cl = derivative[vlcLn == "contrast_long"],
            eff_size = derivative_cl - derivative_cs,
            phone_dur_cs = phone_dur[vlcLn == "contrast_short"],
            phone_dur_cl = phone_dur[vlcLn == "contrast_long"]) %>%
  ungroup() %>%
  arrange(desc(eff_size))

derivs_vln_mono_nocontrast <- newdat_vln_mono_deriv %>%
  dplyr::select(vowel_length_contrast, language, vlcLn, syll_pos,
                phone_dur_uttr, model, phone_dur, derivative) %>%
  filter(vowel_length_contrast != "unrestricted") %>%
  group_by(vowel_length_contrast, language, phone_dur_uttr, model, syll_pos) %>%
  summarise(derivative_nc = derivative[vlcLn == "no_contrast"],
            phone_dur_nc = phone_dur[vlcLn == "no_contrast"]) %>%
  ungroup() %>%
  arrange(desc(derivative_nc))

eff_sizes_vln_mono[[m]] <- full_join(eff_sizes_vln_mono_contrast,
                                     derivs_vln_mono_nocontrast,
                                     by=c("vowel_length_contrast",
                                          "language",
                                          "phone_dur_uttr",
                                          "model",
                                          "syll_pos"))






eff_curves_vln_mono_contrast[[m]] <- newdat_vl_deriv_curve %>%
  dplyr::select(vowel_length_contrast, language, vlcLn, syll_pos,
                phone_dur_uttr, model, phone_dur, derivative,
                phone_dur_uttr_med, density) %>%
  filter(vlcLn != "no_contrast") %>%
  group_by(vowel_length_contrast, language, phone_dur_uttr_med, phone_dur_uttr, density, model, syll_pos) %>%
  summarise(derivative_cs = derivative[vlcLn == "contrast_short"],
            derivative_cl = derivative[vlcLn == "contrast_long"],
            delta_deriv = derivative_cl - derivative_cs,
            phone_dur_cs = phone_dur[vlcLn == "contrast_short"],
            phone_dur_cl = phone_dur[vlcLn == "contrast_long"]) %>%
  group_by(language) %>%
  mutate(derivative_cs_w = derivative_cs*density,
         derivative_cl_w = derivative_cl*density,
         delta_deriv_w = delta_deriv*density,
         phone_dur_cs_w = phone_dur_cs*density,
         phone_dur_cl_w = phone_dur_cl*density) %>%
  ungroup()
deriv_curves_vln_mono_nocontrast[[m]] <- newdat_vl_deriv_curve %>%
  dplyr::select(vowel_length_contrast, language, vlcLn, syll_pos,
                phone_dur_uttr, model, phone_dur, derivative,
                phone_dur_uttr_med, density) %>%
  filter(vowel_length_contrast != "unrestricted") %>%
  group_by(vowel_length_contrast, language, phone_dur_uttr_med, phone_dur_uttr, density, model, syll_pos) %>%
  summarise(derivative_nc = derivative[vlcLn == "no_contrast"],
            phone_dur_nc = phone_dur[vlcLn == "no_contrast"]) %>%
  group_by(language) %>%
  mutate(derivative_nc_w = derivative_nc*density,
         phone_dur_nc_w = phone_dur_nc*density) %>%
  ungroup()
eff_curves_vln_mono[[m]] <- full_join(eff_curves_vln_mono_contrast[[m]],
                                deriv_curves_vln_mono_nocontrast[[m]],
                                by=c("vowel_length_contrast",
                                     "language",
                                     "phone_dur_uttr",
                                     "phone_dur_uttr_med",
                                     "density",
                                     "model",
                                     "syll_pos"))



eff_sizes_dw_vln_mono[[m]] <- eff_curves_vln_mono[[m]] %>%
  group_by(language,syll_pos) %>%
  summarize(density_area = sum(density),
            derivative_cs_w_area = sum(derivative_cs_w),
            derivative_cl_w_area = sum(derivative_cl_w),
            derivative_nc_w_area = sum(derivative_nc_w),
            delta_deriv_w_area = sum(delta_deriv_w),
            phone_dur_cl_w_area = sum(phone_dur_cl_w),
            phone_dur_cs_w_area = sum(phone_dur_cs_w),
            phone_dur_nc_w_area = sum(phone_dur_nc_w),
            eff_size_dw = delta_deriv_w_area/density_area,
            derivative_cs_dw = derivative_cs_w_area/density_area,
            derivative_cl_dw = derivative_cl_w_area/density_area,
            derivative_nc_dw = derivative_nc_w_area/density_area,
            phone_dur_cl_dw = phone_dur_cl_w_area/density_area,
            phone_dur_cs_dw = phone_dur_cs_w_area/density_area,
            phone_dur_nc_dw = phone_dur_nc_w_area/density_area
            ) %>%
  select(-c(density_area,
                   derivative_cs_w_area,
                   derivative_cl_w_area,
                   derivative_nc_w_area,
                   delta_deriv_w_area,
                   phone_dur_cl_w_area,
                   phone_dur_cs_w_area,
                   phone_dur_nc_w_area,
                   ))




eff_sizes_dw_vln_mono[[m]] <- left_join(
  eff_sizes_dw_vln_mono[[m]],
  rename(eff_sizes_vln_mono[[m]],eff_size_med = eff_size),
  by = c("language","syll_pos")
)

eff_sizes_dw_vln_mono[[m]] <- eff_sizes_dw_vln_mono[[m]] %>%
  arrange(desc(eff_size_dw))
}

eff_curves_vln_mono_df <- bind_rows(eff_curves_vln_mono)
eff_curves_vln_mono_df %>% write_csv("eff_curves_vln_mono_df.csv")
eff_curves_vln_mono_df <- read_csv("./eff_curves_vln_mono_df.csv")

eff_sizes_dw_vln_mono_df <- bind_rows(eff_sizes_dw_vln_mono)
eff_sizes_dw_vln_mono_df %>% write_csv("eff_sizes_dw_vln_mono_df.csv")
eff_sizes_dw_vln_mono_df <- read_csv("./eff_sizes_dw_vln_mono_df.csv")
```



## Making VL Prediction Plots Showing Density-Weighted Effect Sizes

Ordering by effect size at medians & showing this number in the plot of VL model predictions.

```{r}

for (m in mods_vln_mono) {
  eff_sizes_dw_vln_mono_m <- eff_sizes_dw_vln_mono[[m]]
  eff_sizes_dw_vln_mono_m$language <- factor(eff_sizes_dw_vln_mono_m$language,
                                          levels=unique(eff_sizes_dw_vln_mono_m$language))
  eff_sizes_dw_vln_mono_m$phone_dur_uttr_median <- eff_sizes_dw_vln_mono_m$phone_dur_uttr
  eff_sizes_dw_vln_mono_m$phone_dur_uttr <- 0.05
  eff_sizes_dw_vln_mono_m$value <- 0.13
  
  newdat_m <- filter(newdat_vln_mono_all, model==m)
  newdat_m$language <- factor(newdat_m$language, levels=unique(eff_sizes_dw_vln_mono_m$language))
  
  plot_dat <- newdat_m
  plot_dat$language <- factor(plot_dat$language, levels=unique(eff_sizes_dw_vln_mono_m$language))
  
  d_uttr_vln_mono$language <- factor(d_uttr_vln_mono$language, levels=unique(eff_sizes_dw_vln_mono_m$language))
  
  plot_dat %>%
  #filter(vowel_length_contrast != "none") %>%
  ggplot(data = ., aes(x = phone_dur_uttr, y=value)) +
  facet_wrap(. ~ language) +
  geom_point(data = d_uttr_vln_mono,
             aes(y = value, col=measure), alpha = 0.3, pch=16) +
  geom_line(aes(y = dur, col = vlcLn_o, lwd = syll_pos_o)) +
  # geom_ribbon(data = filter(newdat_cv_all,
  #                          model == "cv_mod_baseline",
  #                          segtype_o == "consonant"),
  #           aes(ymin = ll, ymax = ul), col = NA, fill = "skyblue",
  #           alpha = 0.5) +
  # geom_ribbon(data = filter(newdat_cv_all,
  #                          model == "cv_mod_baseline",
  #                          segtype_o == "syllabic"),
  #           aes(ymin = ll, ymax = ul), col = NA, fill = "yellow",
  #           alpha = 0.5) +
  geom_text(data=filter(eff_sizes_dw_vln_mono_m,!is.na(derivative_cs),syll_pos=='nonfinal'), 
            aes(y=value,label=paste0("Δβ = ", round(eff_size_dw,2))),
            col="black", size=5) +
    geom_text(data=filter(eff_sizes_dw_vln_mono_m,!is.na(derivative_cs),syll_pos=='nonfinal'), 
            aes(y=value*1.4,label=paste0("β_s = ", round(derivative_cs_dw,2))),
            col="black", size=5) +
    geom_text(data=filter(eff_sizes_dw_vln_mono_m,!is.na(derivative_cl),syll_pos=='nonfinal'), 
            aes(y=value*1.4*1.4,label=paste0("β_l = ", round(derivative_cl_dw,2))),
            col="black", size=5) +
    geom_text(data=filter(eff_sizes_dw_vln_mono_m,!is.na(derivative_nc),syll_pos=='nonfinal'), 
            aes(y=value*1.4*1.4*1.4,label=paste0("β = ", round(derivative_nc_dw,2))),
            col="black", size=5) +
  geom_text(data=filter(eff_sizes_dw_vln_mono_m,!is.na(derivative_cs_dw),syll_pos=='final'), 
            aes(x=phone_dur_uttr*2,
                y=value,
                label=paste0("Δβ = ", round(eff_size,2))),
            col="black", size=5) +
    geom_text(data=filter(eff_sizes_dw_vln_mono_m,!is.na(derivative_cs_dw),syll_pos=='final'), 
            aes(x=phone_dur_uttr*2,
                y=value*1.4,
                label=paste0("β_s = ", round(derivative_cs_dw,2))),
            col="black", size=5) +
    geom_text(data=filter(eff_sizes_dw_vln_mono_m,!is.na(derivative_cl_dw),syll_pos=='final'), 
            aes(x=phone_dur_uttr*2,
                y=value*1.4*1.4,
                label=paste0("β_l = ", round(derivative_cl_dw,2))),
            col="black", size=5) +
    geom_text(data=filter(eff_sizes_dw_vln_mono_m,!is.na(derivative_nc_dw),syll_pos=='final'), 
            aes(x=phone_dur_uttr*2,
                y=value*1.4*1.4*1.4,
                label=paste0("β = ", round(derivative_nc_dw,2))),
            col="black", size=5) +
  # geom_text(data=eff_sizes_dw_vln_mono_m,
  #           aes(x=phone_dur_uttr*1.4,
  #               y=value/1.4/1.4/1.4,label=paste0("med = ", round(phone_dur_uttr,2))),
  #           col="black", size=5) +
  scale_color_manual(values = c("no_contrast" = "magenta",
                                "contrast_long" = "blue",
                                "contrast_short" = "limegreen",
                                "nocontrast_dur_uttr" = "hotpink", 
                                "long_dur_uttr" = "cyan", 
                                "short_dur_uttr" = "green"),
                     guide = "none") +
  # scale_linetype_manual(values = c("high" = "dashed",
  #                                  "nonhigh" = "solid"
  #                                  )) +
  scale_linewidth_manual(values = c("final" = 1,
                                    "nonfinal" = 0.5)) +
  scale_x_log10(breaks = c(0.05, 0.07, 0.1, 0.14)) +
  scale_y_log10(breaks = c(0.05, 0.07, 0.1, 0.14),
                #limits = c(0.03, 0.2)
                ) +
  #scale_x_continuous(limits=c(0.035,0.145), 
  #                   breaks=seq(0.04,0.14,0.02),
  #                   labels=c(".04",".06",".08",".10",".12",".14")) +
  #scale_y_continuous(breaks=seq(0.04,0.20,0.02)) +
    # labels for axes
  xlab("Average phone duration by utterance") +
  labs(y = "Average long vs short or no-contrast monophthong duration in s") +
  #ylab("Average V vs. C duration in s") +
  # remove clutter
  theme_minimal() +
  theme(panel.grid = element_blank(),
        axis.line = element_line(size=0.8),
        axis.ticks = element_line(size=0.8),
        axis.text = element_text(size=16, colour="black"),
        axis.title = element_text(size=18, face="bold"),
        strip.background.y = element_blank(), 
        strip.placement = "outside",
        strip.text=element_text(size=18, face="bold"))
  ggsave(paste0("doreco_graphs_v1/vl/", m, ".png"), width=15, height=15, dpi=300)
}


```

## Plotting VL Model Effect Curves

```{r}
for (m in mods_vln_mono) {

  eff_curves_vln_mono_m <- eff_curves_vln_mono[[m]]
  eff_sizes_dw_vln_mono_m <- eff_sizes_dw_vln_mono[[m]]
  
  eff_sizes_dw_vln_mono_m$language <- factor(eff_sizes_dw_vln_mono_m$language,
                                             levels=unique(eff_sizes_dw_vln_mono_m$language))
  eff_curves_vln_mono_m$language <- factor(eff_curves_vln_mono_m$language,
                                           levels=unique(eff_curves_vln_mono_m$language))
  
  eff_curves_vln_mono_m %>%
    ggplot(data = ., aes(x = phone_dur_uttr)) +
    facet_wrap(. ~ language) +
    #geom_ribbon(data = filter(newdat_cv_all,
    #                          model == "dur",
    #                          segtype_o == "vowel_dur_uttr"),
    #            aes(ymin = ll, ymax = ul), col = NA, fill = "grey",
    #            alpha = 0.5) +
    #geom_ribbon(data = filter(newdat_cv,
    #                          measure_type == "dur",
    #                          segtype_o == "cons_dur_uttr"),
    #            aes(ymin = ll, ymax = ul), col = NA, fill = "grey",
    #            alpha = 0.5) +
    # geom_density(data = d_cv,
    #              aes(x = phone_dur_uttr, after_stat(scaled)),
    #              n = 100, trim = FALSE, #bounds = c(min(d_cv$phone_dur_uttr),max(d_cv$phone_dur_uttr)),
    #              fill = "forestgreen",lwd=0, alpha=0.3) +
    geom_area(aes(y = density),
              fill = "darkgreen",alpha=0.3) +
    # geom_line(aes(y = derivative_v),
    #           col = "orange", lwd = 1) +
    # geom_line(aes(y = derivative_c),
    #           col = "purple", lwd = 1) +
    geom_line(aes(y = delta_deriv,lwd = syll_pos),
              col = "darkblue") +
    # geom_line(aes(y = delta_deriv_w),
    #           col = "firebrick", lwd = 1) +
    geom_text(data = eff_sizes_dw_vln_mono_m,
              aes(label=paste0("Δβ[w] == ", round(eff_size_dw,2)),
                  x = 0.05, y = 0.9),
              col="darkblue", size=5, parse=TRUE) +
    geom_text(data = eff_sizes_dw_vln_mono_m,
              aes(label=paste0("Δβ[m] == ", round(eff_size_med,2)),
                  x = 0.05, y = 0.75),
              col="red", size=5, parse=TRUE) +
    geom_hline(data = eff_sizes_dw_vln_mono_m,
               aes(yintercept = eff_size_dw),
               linetype = 'dashed',
               col = "darkblue") +
    geom_hline(data = eff_sizes_dw_vln_mono_m,
               aes(yintercept = eff_size_med),
               linetype = 'dashed',
               col = "red") +
    geom_hline(yintercept = 0, linetype = 'dashed') +
    geom_vline(aes(xintercept = phone_dur_uttr_med),
               col = 'red') +
    scale_linewidth_manual(values = c("final" = 1,
                                      "nonfinal" = 0.5)) +
    scale_x_log10(breaks = c(0.05, 0.07, 0.1, 0.14)) +
    scale_y_continuous(
      limits=c(-0.2,1.1)
      ) +
    labs(title = "Effect Curves",
         x = "Average segment duration in s (inverse speech rate)",
         y = "Delta derivative") +
    theme_minimal() +
    theme(panel.grid = element_blank(),
        title = element_text(size=24, face="bold"),
        axis.line = element_line(size=0.8),
        axis.ticks = element_line(size=0.8),
        axis.text = element_text(size=16, colour="black"),
        axis.title = element_text(size=18, face="bold"),
        strip.background.y = element_blank(), 
        strip.placement = "outside",
        strip.text=element_text(size=18, face="bold"))
  ggsave(paste0("doreco_graphs_v1/vl/effect_curves_", m, ".png"), width=15, height=15, dpi=300)
}
```

Derivative curves

```{r}
for (m in mods_vln_mono) {

  eff_curves_vln_mono_m <- eff_curves_vln_mono[[m]]
  eff_sizes_dw_vln_mono_m <- eff_sizes_dw_vln_mono[[m]]
  
  eff_sizes_dw_vln_mono_m$language <- factor(eff_sizes_dw_vln_mono_m$language,
                                     levels=unique(eff_sizes_dw_vln_mono_m$language))
  eff_curves_vln_mono_m$language <- factor(eff_curves_vln_mono_m$language,
                                     levels=unique(eff_sizes_dw_vln_mono_m$language))
  
  eff_curves_vln_mono_m %>%
    ggplot(data = ., aes(x = phone_dur_uttr)) +
    facet_wrap(. ~ language) +
    #geom_ribbon(data = filter(newdat_cv_all,
    #                          model == "dur",
    #                          segtype_o == "vowel_dur_uttr"),
    #            aes(ymin = ll, ymax = ul), col = NA, fill = "grey",
    #            alpha = 0.5) +
    #geom_ribbon(data = filter(newdat_cv,
    #                          measure_type == "dur",
    #                          segtype_o == "cons_dur_uttr"),
    #            aes(ymin = ll, ymax = ul), col = NA, fill = "grey",
    #            alpha = 0.5) +
    # geom_density(data = d_cv,
    #              aes(x = phone_dur_uttr, after_stat(scaled)),
    #              n = 100, trim = FALSE, #bounds = c(min(d_cv$phone_dur_uttr),max(d_cv$phone_dur_uttr)),
    #              fill = "forestgreen",lwd=0, alpha=0.3) +
    # geom_line(aes(y = density),
    #           col = "darkgreen", lwd = 0.5) +
    geom_area(aes(y = density/2+1),
              fill = "darkgreen",alpha=0.3) +
    geom_line(aes(y = derivative_nc,lwd = syll_pos),
              col = "magenta") +
    geom_line(aes(y = derivative_cl,lwd = syll_pos),
              col = "blue") +
    geom_line(aes(y = derivative_cs,lwd = syll_pos),
              col = "limegreen") +
    scale_linewidth_manual(values = c("final" = 1,
                                      "nonfinal" = 0.5)) +
    geom_hline(yintercept = 1, linetype = 'dashed') +
    # geom_line(aes(y = delta_deriv),
    #           col = "darkblue", lwd = 1) +
    geom_vline(aes(xintercept = phone_dur_uttr_med),
               col = 'red') +
    scale_x_log10(breaks = c(0.05, 0.07, 0.1, 0.14)) +
    scale_y_log10(n.breaks = 8) +
    labs(title = "Derivative Curves",
         x = "Average segment duration in s (inverse speech rate)",
         y = "Predicted phone duration") +
    theme_minimal() +
    theme(panel.grid = element_blank(),
        title = element_text(size=24, face="bold"),
        axis.line = element_line(size=0.8),
        axis.ticks = element_line(size=0.8),
        axis.text = element_text(size=16, colour="black"),
        axis.title = element_text(size=18, face="bold"),
        strip.background.y = element_blank(), 
        strip.placement = "outside",
        strip.text=element_text(size=18, face="bold"))
  ggsave(paste0("doreco_graphs_v1/vl/derivative_curves_", m, ".png"), width=15, height=15, dpi=300)
}
```

## Comparing VL Model Density-Weighted Effect Sizes

Comparing predicted phone durations and slopes for VL models

```{r}
for (m in mods_vln_mono_pos) {
  eff_sizes_vln_mono_pos_m <- eff_sizes_dw_vln_mono[[m]]
  
  eff_sizes_vln_mono_pos_m_wider <- eff_sizes_vln_mono_pos_m %>%
  pivot_wider(names_from = syll_pos,
              values_from = c(eff_size_med,
                              derivative_nc,
                              derivative_cs,
                              derivative_cl,
                              phone_dur_nc,
                              phone_dur_cs,
                              phone_dur_cl,
                              eff_size_dw,
                              derivative_nc_dw,
                              derivative_cs_dw,
                              derivative_cl_dw,
                              phone_dur_nc_dw,
                              phone_dur_cs_dw,
                              phone_dur_cl_dw))
  
  eff_sizes_vln_mono_pos_m %>%
  ggplot() +
    facet_wrap(~vowel_length_contrast) +
    geom_point(aes(x=phone_dur_nc_dw,y=derivative_nc_dw,shape=syll_pos),col="magenta") +
    geom_point(aes(x=phone_dur_cl_dw,y=derivative_cl_dw,shape=syll_pos),col="blue") +
    geom_point(aes(x=phone_dur_cs_dw,y=derivative_cs_dw,shape=syll_pos),col="limegreen") +
    geom_segment(aes(x = phone_dur_cs_dw, xend=phone_dur_cl_dw,
                     y = derivative_cs_dw, yend=derivative_cl_dw,
                     col = vowel_length_contrast)) +
    geom_segment(data = eff_sizes_vln_mono_pos_m_wider,
                   aes(x = phone_dur_nc_dw_final, xend=phone_dur_cs_dw_nonfinal,
                     y = derivative_nc_dw_final, yend=derivative_cs_dw_nonfinal,
                     col = vowel_length_contrast),
                 linetype="dashed") +
    geom_segment(data = eff_sizes_vln_mono_pos_m_wider,
                 aes(x = phone_dur_nc_dw_final, xend=phone_dur_cl_dw_nonfinal,
                     y = derivative_nc_dw_final, yend=derivative_cl_dw_nonfinal,
                     col = vowel_length_contrast),
                 linetype="dashed") +
    geom_segment(data = eff_sizes_vln_mono_pos_m_wider,
                   aes(x = phone_dur_nc_dw_final, xend=phone_dur_nc_dw_nonfinal,
                     y = derivative_nc_dw_final, yend=derivative_nc_dw_nonfinal),
                 linetype="dashed") +
    geom_segment(data = eff_sizes_vln_mono_pos_m_wider,
                   aes(x = phone_dur_cl_dw_final, xend=phone_dur_cl_dw_nonfinal,
                     y = derivative_cl_dw_final, yend=derivative_cl_dw_nonfinal),
                 linetype="dashed") +
    geom_segment(data = eff_sizes_vln_mono_pos_m_wider,
                   aes(x = phone_dur_cs_dw_final, xend=phone_dur_cs_dw_nonfinal,
                     y = derivative_cs_dw_final, yend=derivative_cs_dw_nonfinal),
                 linetype="dashed") +
    scale_color_manual(values = c("unrestricted" = "black",
                                "restricted" = "red")) +
    scale_x_log10(breaks=c(0.067,0.1,0.15)) +
    xlab("Predicted individual phone duration (density weighted)") +
    labs(y = "β_dw") +
    theme_minimal() +
    theme(panel.grid = element_blank(),
        axis.line = element_line(size=0.8),
        axis.ticks = element_line(size=0.8),
        axis.text = element_text(size=16, colour="black"),
        axis.title = element_text(size=18, face="bold"),
        strip.background.y = element_blank(), 
        strip.placement = "outside",
        strip.text=element_text(size=18, face="bold"))
  ggsave(paste0("doreco_graphs_v1/vl/vl_deriv_preds_all_dw_", m, ".png"), width=12, height=5, dpi=300)
}
```

```{r warning=FALSE}
m<-"vln_mod_mono_full"

eff_sizes_vln_mono_pos_m <- eff_sizes_dw_vln_mono[[m]]
  
eff_sizes_vln_mono_pos_m_wider <- eff_sizes_vln_mono_pos_m %>%
pivot_wider(names_from = syll_pos,
            values_from = c(eff_size_med,
                            derivative_nc,
                            derivative_cs,
                            derivative_cl,
                            phone_dur_nc,
                            phone_dur_cs,
                            phone_dur_cl,
                            eff_size_dw,
                            derivative_nc_dw,
                            derivative_cs_dw,
                            derivative_cl_dw,
                            phone_dur_nc_dw,
                            phone_dur_cs_dw,
                            phone_dur_cl_dw))

eff_sizes_vln_mono_pos_m_wider %>%
  ggplot() +
  #facet_wrap(~vowel_length_contrast) +
  geom_vline(aes(xintercept=1), color="red") +
  geom_boxplot(aes(x=phone_dur_nc_dw_final/phone_dur_nc_dw_nonfinal,
                   y="no_contrast",
                   fill = vowel_length_contrast),alpha=0.25) +
  geom_point(aes(x=phone_dur_nc_dw_final/phone_dur_nc_dw_nonfinal,
                   y="no_contrast",
                   fill = vowel_length_contrast),shape=21) +
  geom_boxplot(aes(x=phone_dur_cl_dw_final/phone_dur_cl_dw_nonfinal,
                   y="contrast_long",
                   fill = vowel_length_contrast),alpha=0.25) +
  geom_point(aes(x=phone_dur_cl_dw_final/phone_dur_cl_dw_nonfinal,
                   y="contrast_long",
                   fill = vowel_length_contrast),shape=21) +
  geom_boxplot(aes(x=phone_dur_cs_dw_final/phone_dur_cs_dw_nonfinal,
                   y="contrast_short",
                   fill = vowel_length_contrast),alpha=0.25) +
  geom_point(aes(x=phone_dur_cs_dw_final/phone_dur_cs_dw_nonfinal,
                   y="contrast_short",
                   fill = vowel_length_contrast),shape=21) +
  geom_boxplot(aes(x=phone_dur_nc_dw_final/phone_dur_cs_dw_nonfinal,
                   y="no_contrast vs contrast_short",
                   fill = vowel_length_contrast),alpha=0.25) +
  geom_point(aes(x=phone_dur_nc_dw_final/phone_dur_cs_dw_nonfinal,
                   y="no_contrast vs contrast_short",
                   fill = vowel_length_contrast),shape=21) +
  scale_fill_manual(values = c("unrestricted" = "blue",
                               "restricted" = "red",
                               "none" = "magenta")) +
  labs(title = "Final Lengthening")

eff_sizes_vln_mono_pos_m_wider %>%
  ggplot() +
  #facet_wrap(~vowel_length_contrast) +
  geom_vline(aes(xintercept=0), color="red") +
  geom_boxplot(aes(x=derivative_nc_dw_final-derivative_nc_dw_nonfinal,
                   y="no_contrast",
                   fill = vowel_length_contrast),alpha=0.25) +
  geom_point(aes(x=derivative_nc_dw_final-derivative_nc_dw_nonfinal,
                   y="no_contrast",
                   fill = vowel_length_contrast),shape=21) +
  geom_boxplot(aes(x=derivative_cl_dw_final-derivative_cl_dw_nonfinal,
                   y="contrast_long",
                   fill = vowel_length_contrast),alpha=0.25) +
  geom_point(aes(x=derivative_cl_dw_final-derivative_cl_dw_nonfinal,
                   y="contrast_long",
                   fill = vowel_length_contrast),shape=21) +
  geom_boxplot(aes(x=derivative_cs_dw_final-derivative_cs_dw_nonfinal,
                   y="contrast_short",
                   fill = vowel_length_contrast),alpha=0.25) +
  geom_point(aes(x=derivative_cs_dw_final-derivative_cs_dw_nonfinal,
                   y="contrast_short",
                   fill = vowel_length_contrast),shape=21) +
  geom_boxplot(aes(x=derivative_nc_dw_final-derivative_cs_dw_nonfinal,
                   y="no_contrast vs contrast_short",
                   fill = vowel_length_contrast),alpha=0.25) +
  geom_point(aes(x=derivative_nc_dw_final-derivative_cs_dw_nonfinal,
                   y="no_contrast vs contrast_short",
                   fill = vowel_length_contrast),shape=21) +
  scale_fill_manual(values = c("unrestricted" = "blue",
                               "restricted" = "red",
                               "none" = "magenta")) +
  labs(title = "Final Stretchening")

eff_sizes_vln_mono_pos_m_wider %>%
  ggplot() +
  #facet_wrap(~vowel_length_contrast) +
  geom_vline(aes(xintercept=1), color="red") +
  geom_point(aes(x=phone_dur_cl_dw_final/phone_dur_cs_dw_final,
                   y="final",
                   fill = vowel_length_contrast),shape=21) +
  geom_boxplot(aes(x=phone_dur_cl_dw_final/phone_dur_cs_dw_final,
                   y="final",
                   fill = vowel_length_contrast),alpha=0.25) +
  geom_point(aes(x=phone_dur_cl_dw_nonfinal/phone_dur_cs_dw_nonfinal,
                   y="nonfinal",
                   fill = vowel_length_contrast),shape=21) +
  geom_boxplot(aes(x=phone_dur_cl_dw_nonfinal/phone_dur_cs_dw_nonfinal,
                   y="nonfinal",
                   fill = vowel_length_contrast),alpha=0.25) +
  scale_fill_manual(values = c("unrestricted" = "blue",
                               "restricted" = "red",
                               "none" = "magenta")) +
  labs(title = "Long Lengthening")

eff_sizes_vln_mono_pos_m_wider %>%
  ggplot() +
  #facet_wrap(~vowel_length_contrast) +
  geom_vline(aes(xintercept=0), color="red") +
  geom_point(aes(x=derivative_cl_dw_final-derivative_cs_dw_final,
                   y="final",
                   fill = vowel_length_contrast),shape=21) +
  geom_boxplot(aes(x=derivative_cl_dw_final-derivative_cs_dw_final,
                   y="final",
                   fill = vowel_length_contrast),alpha=0.25) +
  geom_point(aes(x=derivative_cl_dw_nonfinal-derivative_cs_dw_nonfinal,
                   y="nonfinal",
                   fill = vowel_length_contrast),shape=21) +
  geom_boxplot(aes(x=derivative_cl_dw_nonfinal-derivative_cs_dw_nonfinal,
                   y="nonfinal",
                   fill = vowel_length_contrast),alpha=0.25) +
  scale_fill_manual(values = c("unrestricted" = "blue",
                               "restricted" = "red",
                               "none" = "magenta")) +
  labs(title = "Long Stretchening")
```
```{r}
m<-"vln_mod_mono_full"

eff_sizes_vln_mono_pos_m <- eff_sizes_dw_vln_mono[[m]]
  
eff_sizes_vln_mono_pos_m_wider <- eff_sizes_vln_mono_pos_m %>%
pivot_wider(names_from = syll_pos,
            values_from = c(eff_size_med,
                            derivative_nc,
                            derivative_cs,
                            derivative_cl,
                            phone_dur_nc,
                            phone_dur_cs,
                            phone_dur_cl,
                            eff_size_dw,
                            derivative_nc_dw,
                            derivative_cs_dw,
                            derivative_cl_dw,
                            phone_dur_nc_dw,
                            phone_dur_cs_dw,
                            phone_dur_cl_dw))

eff_sizes_vln_mono_pos_m_wider %>%
  ggplot() +
  #facet_wrap(~vowel_length_contrast) +
  geom_vline(aes(xintercept=0), color="red") +
  geom_boxplot(aes(x=derivative_nc_dw_nonfinal,
                   y="no_contrast",
                   fill = vowel_length_contrast),alpha=0.25) +
  geom_point(aes(x=derivative_nc_dw_nonfinal,
                   y="no_contrast",
                   fill = vowel_length_contrast),shape=21) +
  geom_boxplot(aes(x=derivative_cl_dw_nonfinal,
                   y="contrast_long",
                   fill = vowel_length_contrast),alpha=0.25) +
  geom_point(aes(x=derivative_cl_dw_nonfinal,
                   y="contrast_long",
                   fill = vowel_length_contrast),shape=21) +
  geom_boxplot(aes(x=derivative_cs_dw_nonfinal,
                   y="contrast_short",
                   fill = vowel_length_contrast),alpha=0.25) +
  geom_point(aes(x=derivative_cs_dw_nonfinal,
                   y="contrast_short",
                   fill = vowel_length_contrast),shape=21) +
  scale_fill_manual(values = c("unrestricted" = "blue",
                               "restricted" = "red",
                               "none" = "magenta")) +
  labs(title = "Stretchiness by Vowel Length")
```

